### 2018-07-07 ms.Rmd
Initial setup

---
title: "Austraits – a curated plant trait database for the Australian flora"
bibliography: references.bib
csl: nature.csl  
output:
  word_document: default
  pdf_document:
    keep_tex: yes
  html_document: default
geometry: margin=0.75in
fontsize: 9pt
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmarkdown)
library(readr)
library(dplyr)

austraits <- readRDS("data/austraits.rds")
```

```{r authors, results='asis', echo=FALSE}
authors <- read_csv("data/authors.csv", col_types = cols(.default = "c"))

affils <- tibble(Affiliation=unique(authors$Affiliation), 
                 label = letters[seq_len(length(Affiliation)) ])
authors <- left_join(authors, affils, by = "Affiliation")

# print list of authors
sprintf("%s (%s)", authors$Author, authors$label) %>% paste0(collapse = ", ") %>% writeLines()

writeLines("\n\n")

# print list of affiliations
sprintf("**%s**- %s", affils$label, affils$Affiliation) %>% paste0(collapse = "; ") %>% writeLines()
```

# Abstract

Species traits have increasingly become an essential metric for comparing ecological strategies among plants across environmental space or evolutionary lineages. This is because traits are useful indicators of the the functions that species perform, regardless of their taxonomy, in communities and ecosystems. Whilst the volume of trait information has grown rapidly in the last two decades through targeted initiatives, the geographic coverage of trait observations across the globe is patchy, making comprehensive analyses of patterns of trait abundance and diversity impossible to achieve in some regions. One such region is Australia; a continent with a flora of over 20,000 higher-plant species, % of which are considered endemic. Here, we introduce a fully open database of XXX plant traits for the Australian Flora (named AusTraits). Austraits synthesises data from field campaigns, published literature, taxonomic monographs, and individual species descriptions. Traits vary in scope from physiological measures of performance (e.g. photosynthetic gas exchange, water-use efficiency, XX) to morphological parameters (e.g. leaf size, seed size, maximum height) which link to ecological strategy variation. AusTraits contains curated and harmonised species- and genus-level observations coupled to, where available, contextual information on site properties. This data descriptor provides information on Version 1.1 of AusTraits which houses data for XXX trait-by-species combinations across XX,XXX species. We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding the Australian flora.

# Background & Summary

- traits now widely agreed as important [@westoby2002]
-	Legacy data vs. newly-collected traits  considerations for these differ and Austraits can offer a path for how to sample comprehensively
-	Web portal, maybe like coral traits, for easily adding new data which can accommodate a range of user’s capabilities
-	Family level plotting for the paper, featuring two traits – continuous and categorical  height and growth form; two important traits which we have good coverage for
-	CURATED national collection of traits, similar to other biodiversity datasets 
-	Expose decisions which are being made in the processing of data into a harmonised resource with a view to nationalising the availability of trait information, similar to availability similar to the availability of species occurrence data
-	What can a national dataset facilitate? Alignment with conservation management across a federlaised system of government. Global platforms might be best built from national datasets (analogous to the GBIF initiative which aggregates occurrence data assembled by national initiatives). 
-	Current models of trait aggregation are geographically patchy – with some locations comprehensively covered e.g. Europe – but with little taxonomic or functional overlap with other regions (e.g. Africa, Australia). ~5% of the worlds plant species occur in these regions (300 trees in Europe 600 in America; in a single 50-ha plot in Australian tropical forests = same or more species). This actually matters when we increasingly want to apply traits as a proxy for understanding vegetation responses to disturbance and change. Whilst large-scale analyses are relatively well-served by global datasets which have representatives species from each biome or continent, the capacity to explore more nuanced questions about the function of regional vegetation is limited. 
-	Duplication of effort. Austraits will reduce the duplication of effort across government research agencies and university researchers seeking to create trait-based approaches to their questions. 
-	Better stewardship of data when collected by a localised group of interested parties.

# Methods

The data are housed within github repository traitecoevo with sequential releases being made available on this platform and Zenodo/Figshare. Access to Austraits will be  XXX. Austraits can also be accessed via the R package: `austraits`

## Processing pipeline:

1. what is the platform for curating and harmonising? 
2. which decisions do we think it is important to expose e.g. taxonomic alignment, trait mapping == standardisation of names and units, data cleaning processes (e.g. substitutions, reshaping of data, excluding values outside plausible bounds  we can put numbers on these as Daniel has the data-frame of excluded data).

In this way we offer more than an aggregation service, but seek to standardise data provided from a range of disparate sources and individuals.

## Schema

Austraits is designed to collate traits-based measurements with a range of relevant contextual information. Its schema broadly follows the principles of the established Observation and Measurement Ontology (Madin et al., 2007) in that trait data are connected to contextual information. Data flows from individual ‘studies’ which are any natural grouping for data collections, typically scientific publications, species descriptions, and field campaigns. Austraits currently contains X studies and each study has two associated files: `data.csv` and `metadata.yml`. `Metadata.yml` must conform to a standard set of fields to provide context and options for converting data into standardised format via information on trait-mapping.

## Traits

Includes data for Type

## Sources

## Measurement

# Taxonomic coverage

Austraits currently contains data for `r austraits$data$species_name %>% unique() %>% length()` species from `r austraits$species_list$family %>% unique() %>% length()` families; with `r austraits$data %>% nrow()` records for `r austraits$data$trait_name %>% unique() %>% length()` traits.

## References

\footnotesize

### 2018-07-07 ms.Rmd
Some relevant references

@@
 
-- traits now widely agreed as important [@westoby2002]
+- traits now widely agreed as important [@Westoby-2002]
 -	Legacy data vs. newly-collected traits  considerations for these differ and Austraits can offer a path for how to sample comprehensively
--	Web portal, maybe like coral traits, for easily adding new data which can accommodate a range of user’s capabilities
+-	Web portal, maybe like coral traits [@Madin-2016], for easily adding new data which can accommodate a range of user’s capabilities
 -	Family level plotting for the paper, featuring two traits – continuous and categorical  height and growth form; two important traits which we have good coverage for
@@
 Access to Austraits will be  XXX.
-Austraits can also be accessed via the R package: `austraits`
+Austraits can also be accessed via the R package: `austraits` Releases versioned, following [@Falster-2017]
 
@@
 Austraits is designed to collate traits-based measurements with a range of relevant contextual information.
-Its schema broadly follows the principles of the established Observation and Measurement Ontology (Madin et al., 2007) in that trait data are connected to contextual information.
+Its schema broadly follows the principles of the established Observation and Measurement Ontology [@Madin-2007] in that trait data are connected to contextual information.
 Data flows from individual ‘studies’ which are any natural grouping for data collections, typically scientific publications, species descriptions, and field campaigns.

### 2018-07-17 ms.Rmd
Add template style.

@@
 bibliography: references.bib
-csl: nature.csl  
+csl: assets/nature.csl  
 output:
-  word_document: default
+  word_document:
+    reference_docx: assets/manuscriptstyle.docx
+  html_document:
+    number_sections: yes
+    theme: flatly
   pdf_document:
-    keep_tex: yes
-  html_document: default
+    keep_tex: no
 geometry: margin=0.75in
@@
 # print list of authors
-sprintf("%s (%s)", authors$Author, authors$label) %>% paste0(collapse = ", ") %>% writeLines()
+sprintf("%s^%s^", authors$Author, authors$label) %>% paste0(collapse = ", ") %>% writeLines()
 
@@
 # print list of affiliations
-sprintf("**%s**- %s", affils$label, affils$Affiliation) %>% paste0(collapse = "; ") %>% writeLines()
+sprintf("^%s^ %s", affils$label, affils$Affiliation) %>% paste0(collapse = "; ") %>% writeLines()
 ```

### 2019-04-04 ms.Rmd
Adding content for paper

@@
     reference_docx: assets/manuscriptstyle.docx
-  html_document:
-    number_sections: yes
-    theme: flatly
-  pdf_document:
-    keep_tex: no
-geometry: margin=0.75in
-fontsize: 9pt
 editor_options:
@@
 
-```{r setup, include=FALSE, warning=FALSE, message=FALSE}
+```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
+# knitr defaults
+root.dir = rprojroot::find_root("austraits_ms.Rproj")
+knitr::opts_knit$set(root.dir = root.dir)
+knitr::opts_chunk$set(echo=FALSE, cache=FALSE, results='asis', message=FALSE, warning=FALSE)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
 library(knitr)
+library(tidyverse)
 library(rmarkdown)
-library(readr)
-library(dplyr)
 
+source("R/ms.R")
+
+## Assumes to items exist in global name space
+if(!exists("austraits")) {
   austraits <- readRDS("data/austraits.rds")
+  # stop("austraits must exist in global name space to knit a report")
+}
+
+definitions <- austraits$definitions
 ```
@@
 
-# Abstract
+# Overview
 
-Species traits have increasingly become an essential metric for comparing ecological strategies among plants across environmental space or evolutionary lineages.
-This is because traits are useful indicators of the the functions that species perform, regardless of their taxonomy, in communities and ecosystems.
-Whilst the volume of trait information has grown rapidly in the last two decades through targeted initiatives, the geographic coverage of trait observations across the globe is patchy, making comprehensive analyses of patterns of trait abundance and diversity impossible to achieve in some regions.
-One such region is Australia; a continent with a flora of over 20,000 higher-plant species, % of which are considered endemic.
-Here, we introduce a fully open database of XXX plant traits for the Australian Flora (named AusTraits).
-Austraits synthesises data from field campaigns, published literature, taxonomic monographs, and individual species descriptions.
-Traits vary in scope from physiological measures of performance (e.g.
-photosynthetic gas exchange, water-use efficiency, XX) to morphological parameters (e.g.
-leaf size, seed size, maximum height) which link to ecological strategy variation.
-AusTraits contains curated and harmonised species- and genus-level observations coupled to, where available, contextual information on site properties.
-This data descriptor provides information on Version 1.1 of AusTraits which houses data for XXX trait-by-species combinations across XX,XXX species.
-We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding the Australian flora.
+AusTraits harmonises data on `r austraits$traits$trait_name %>% unique() %>% length()` traits from `r austraits$methods$dataset_id %>% unique() %>% length()` different sources, including field campaigns, published literature, taxonomic monographs, and individual species descriptions.
 
-# Background & Summary
+This document provides information on the structure of AusTraits and corresponds to Version `r austraits$definitions$austraits$elements$build_info$elements$version$value` of the dataset.
 
-- traits now widely agreed as important [@Westoby-2002]
--	Legacy data vs. newly-collected traits  considerations for these differ and Austraits can offer a path for how to sample comprehensively
--	Web portal, maybe like coral traits [@Madin-2016], for easily adding new data which can accommodate a range of user’s capabilities
--	Family level plotting for the paper, featuring two traits – continuous and categorical  height and growth form; two important traits which we have good coverage for
--	CURATED national collection of traits, similar to other biodiversity datasets 
--	Expose decisions which are being made in the processing of data into a harmonised resource with a view to nationalising the availability of trait information, similar to availability similar to the availability of species occurrence data
--	What can a national dataset facilitate?
-Alignment with conservation management across a federlaised system of government.
-Global platforms might be best built from national datasets (analogous to the GBIF initiative which aggregates occurrence data assembled by national initiatives).
--	Current models of trait aggregation are geographically patchy – with some locations comprehensively covered e.g.
-Europe – but with little taxonomic or functional overlap with other regions (e.g.
-Africa, Australia).
-~5% of the worlds plant species occur in these regions (300 trees in Europe 600 in America; in a single 50-ha plot in Australian tropical forests = same or more species).
-This actually matters when we increasingly want to apply traits as a proxy for understanding vegetation responses to disturbance and change.
-Whilst large-scale analyses are relatively well-served by global datasets which have representatives species from each biome or continent, the capacity to explore more nuanced questions about the function of regional vegetation is limited.
--	Duplication of effort. Austraits will reduce the duplication of effort across government research agencies and university researchers seeking to create trait-based approaches to their questions. 
--	Better stewardship of data when collected by a localised group of interested parties.
+# Data structures
 
-# Methods
+# Tables
 
-The data are housed within github repository traitecoevo with sequential releases being made available on this platform and Zenodo/Figshare.
-Access to Austraits will be  XXX.
-Austraits can also be accessed via the R package: `austraits` Releases versioned, following [@Falster-2017]
+Table 1:  Main elements of Austraits
 
-## Processing pipeline:
+```{r}
+tibble(
+  Element = names(austraits), 
+  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description") 
+  ) %>% 
+  kable()
+```
 
-1. what is the platform for curating and harmonising? 
-2.
-which decisions do we think it is important to expose e.g.
-taxonomic alignment, trait mapping == standardisation of names and units, data cleaning processes (e.g.
-substitutions, reshaping of data, excluding values outside plausible bounds  we can put numbers on these as Daniel has the data-frame of excluded data).
+Table 2: Traits represented in database
+```{r}
+sites <- austraits$sites %>% 
+  filter(site_property %in%  c("longitude (deg)","latitude (deg)")) %>% 
+  spread(site_property, value)
 
-In this way we offer more than an aggregation service, but seek to standardise data provided from a range of disparate sources and individuals.
+data_geo <- 
+  left_join(by=c("dataset_id", "site_name"),
+      select(austraits$traits, -original_name), sites) %>%
+  left_join(by="species_name",
+      select(austraits$taxonomy, species_name, family))
 
-## Schema
+n_records <- 
+  data_geo %>% 
+  group_by(trait_name) %>%
+  summarise(
+            Description = "",
+            Type= "",
+            `Records (all)` = n(),
+            `Records (geo.)` = sum(!is.na(`longitude (deg)`) & !is.na(`latitude (deg)`)),
+            Studies = n_distinct(dataset_id),
+            Species = n_distinct(species_name),
+            Families = n_distinct(family)
+            )
 
-Austraits is designed to collate traits-based measurements with a range of relevant contextual information.
-Its schema broadly follows the principles of the established Observation and Measurement Ontology [@Madin-2007] in that trait data are connected to contextual information.
-Data flows from individual ‘studies’ which are any natural grouping for data collections, typically scientific publications, species descriptions, and field campaigns.
-Austraits currently contains X studies and each study has two associated files: `data.csv` and `metadata.yml`.
-`Metadata.yml` must conform to a standard set of fields to provide context and options for converting data into standardised format via information on trait-mapping.
+n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$trait_name]
+n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$trait_name]
 
-## Traits
+n_records %>%
+  arrange(desc(`Records (all)`)) %>% 
+  slice(1:40) %>%
+  kable()
 
-Includes data for Type
+```
 
-## Sources
+```{r, results="asis"}
+print_defintions_element <- function(elements) {
+  if(elements$type == "character") {
+    sprintf("**Content:** %s\n", elements$value) %>% 
+      writeLines()
+  }
 
-## Measurement
+  if(elements$type == "table") {
 
-# Taxonomic coverage
+    sprintf("**Content:** \n") %>% 
+      writeLines()
 
-Austraits currently contains data for `r austraits$data$species_name %>% unique() %>% length()` species from `r austraits$species_list$family %>% unique() %>% length()` families; with `r austraits$data %>% nrow()` records for `r austraits$data$trait_name %>% unique() %>% length()` traits.
+    elements$elements %>% 
+      list1_to_df() %>%
+      kable() %>% 
+      writeLines()
+  }
+}
+
+for(v in names(austraits)) {
+  elements <- austraits$definitions$austraits$elements[[v]]
+
+  sprintf("### %s {#%s}\n\n**Description:** %s\n", v, v, elements$description) %>% 
+    writeLines()
+
+  elements %>%
+    print_defintions_element()
+
+  writeLines(c(""))
+}
+```
+
+Value types
+
+```{r, results="asis"}
+austraits$definitions$value_type$values %>% 
+  list1_to_df()  %>% 
+  kable() %>% 
+  writeLines()
+```
+
+\footnotesize
+
+# Data sources:
+
+```{r, results='asis'}
+c(austraits$methods$source_primary_citation %>% unique(),
+austraits$methods$source_secondary_citation %>% unique()) %>% paste(collapse = "\n\n") %>% writeLines()
+```
 
 ## References
-
-\footnotesize

### 2019-04-04 R/ms.R

@@ -1,0 +1,11 @@
+# Convert a list with single entries to dataframe
+list1_to_df <- function(my_list) {
+  
+  for(f in names(my_list)) {
+    if(is.null(my_list[[f]])) 
+      my_list[[f]] <- NA
+  }
+  
+  tibble(key = names(my_list), value = unlist(my_list))
+}
+

### 2019-04-04 ms.Rmd
Updating table of traits with references

@@
 source("R/ms.R")
-
-## Assumes to items exist in global name space
-if(!exists("austraits")) {
-  austraits <- readRDS("data/austraits.rds")
-  # stop("austraits must exist in global name space to knit a report")
-}
-
-definitions <- austraits$definitions
 ```
@@
 sprintf("^%s^ %s", affils$label, affils$Affiliation) %>% paste0(collapse = "; ") %>% writeLines()
+```
+
+```{r austraits}
+
+## Assumes to items exist in global name space
+if(!exists("austraits")) {
+  austraits <- readRDS("data/austraits.rds")
+  # stop("austraits must exist in global name space to knit a report")
+}
+
+definitions <- austraits$definitions
+
+sources_all <- 
+  bind_rows(
+    austraits$methods %>% select(dataset_id, citation = source_primary_citation) %>% distinct() %>% mutate(type="primary"),
+    austraits$methods %>% select(dataset_id, citation = source_secondary_citation) %>% distinct() %>% mutate(type="secondary")
+  ) %>% 
+  na.omit() %>% 
+  arrange(dataset_id)
+
+sources_unique <- tibble(citation=unique(sources_all$citation)) %>% mutate(source_number=1:n())
+
+sources_all <- sources_all %>% left_join(sources_unique, by = "citation")
+
+refs <- sources_all %>% group_by(dataset_id) %>% 
+  summarise(refs = paste(sort(source_number), collapse = ", "))
+
 ```
@@
 
+trait_categories <- read_csv("data/trait_categories.csv")
+
 data_geo <- 
@@
   left_join(by="species_name",
-      select(austraits$taxonomy, species_name, family))
+      select(austraits$taxonomy, species_name, family)) %>% 
+  left_join(refs, by=c("dataset_id"))
 
@@
             Species = n_distinct(species_name),
-            Families = n_distinct(family)
-            )
+            Families = n_distinct(family),
+            `Data Citations` = unique(refs) %>% paste(collapse = ", ")
+            ) %>% 
+  left_join(trait_categories, by=c("trait_name")) %>% 
+  select(Tissue = tissue, Category = category, Trait = trait_name, everything()) %>% 
+  arrange(Tissue, Category, Trait)
 
-n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$trait_name]
-n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$trait_name]
+n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$Trait]
+n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$Trait]
 
 n_records %>%
-  arrange(desc(`Records (all)`)) %>% 
-  slice(1:40) %>%
   kable()
-
 ```
@@
 ```{r, results='asis'}
-c(austraits$methods$source_primary_citation %>% unique(),
-austraits$methods$source_secondary_citation %>% unique()) %>% paste(collapse = "\n\n") %>% writeLines()
+sprintf("%s. %s", sources_unique$source_number, sources_unique$citation) %>% paste(collapse = "\n\n") %>% writeLines()
 ```

### 2019-04-04 R/plot_pcp_vs_temp_representation.R
austraits climate space figure

@@ -1,0 +1,95 @@
+## Austraits: Precipitation–temperature space plot
+library(raster)
+library(tidyverse)
+library(hexbin)
+library(ggplot2)
+library(viridis)
+source("R/plotting_functions.R")
+
+#### 01 Get climate data for Australia ####
+# Download bioclim data using library (raster)
+bioclim<-getData("worldclim",var="bio",res=10)
+# Pick BIO1 (Mean Annual Temperature; T) and BIO12 (Annual Precipitation; P)
+bioclim<- bioclim[[c(1,12)]]
+names(bioclim) <- c("Temp","Prec")
+
+#### 02 Get the climate data for Australia ####
+# Load Australia landmass binary map
+au_map<-raster("data/au_binary.tif")
+# Clip bioclim data with the au map
+## crop and mask
+bioclim_au <- crop(bioclim, extent(au_map)) #%>% mask(.,au_map)
+new.bioclim <- projectRaster(bioclim_au, au_map) # harmonize the spatial extent and projection
+au_bioclim<-mask(new.bioclim, au_map)
+
+# Transform raster data into a tibble
+au_bioclim_table<-au_bioclim %>% 
+  as.data.frame() %>%
+  na.omit() %>%
+  as_tibble()
+
+#### 03 Get the point coordinates of Austraits data ####t
+
+# Get the coordinates of the sites
+austraits<-readRDS("data/austraits.rds")
+sites <- austraits$sites %>% 
+  filter(site_property %in%  c("longitude (deg)","latitude (deg)")) %>% 
+  spread(site_property, value)
+sites
+
+
+# get species number for each study
+austraits$traits %>%
+  group_by(dataset_id, site_name) %>%
+  distinct(species_name) %>% 
+  count(dataset_id, sort = TRUE) %>% 
+  ungroup()-> species_number_lookup
+
+# combine species number vs sites
+species_number_lookup %>% na.omit() %>%  
+left_join(sites, 
+          by = c("dataset_id" = "dataset_id", 
+                 "site_name" = "site_name"))->combined_sites
+
+# Get climate values for each site
+au_sites_clim<-combined_sites %>% rowid_to_column("ID") %>% 
+  rename(latitude=`latitude (deg)`, longitude=`longitude (deg)`) %>% 
+  combine_occurence_climate(au_bioclim)
+
+
+#### 03 Plot the climate data ####
+au_sites_clim
+au_bioclim_table
+
[…]

### 2019-04-04 R/plotting_functions.R
austraits climate space figure

@@ -1,0 +1,37 @@
+extract_climate_data<-function(df, climstack){
+  #df is a dataframe of trait data with lat and lon columns
+  #climstack is a stack of gridded climate data
+  
+  df%>%
+    # filter(plot_coord_avail==1)%>%
+    mutate(lat=as.numeric(lat), lon=as.numeric(lon)) %>% 
+    ungroup()%>%
+    dplyr::select( lat, lon)%>%
+    na.omit()-> coord
+  
+  coordinates(coord)=~lon+lat
+  proj4string(coord)=CRS("+proj=longlat +datum=WGS84")
+  raster::extract(climstack,coord,na.rm=F, df=T)->df_climate
+  
+  df_climate
+}
+
+
+#' Combine species occurence and climate data
+#'
+#' @param species_occurence_df 
+#' @param climate_raster_data 
+#'
+#' @return
+#' @export
+#'
+#' @examples
+combine_occurence_climate<-function(species_occurence_df, climate_raster_data){
+  species_occurence_df%>% 
+    group_by(ID)%>%
+    rename(lat=latitude, lon=longitude) %>%
+    extract_climate_data(climate_raster_data) %>%
+    left_join(species_occurence_df,., by="ID")->sp_clim_combined
+  
+  sp_clim_combined
+}

### 2019-04-05 R/fig_site_locations.R
figure for site locations

@@ -1,0 +1,27 @@
+## Austraits: Precipitation–temperature space plot
+library(raster)
+library(tidyverse)
+library(hexbin)
+library(ggplot2)
+library(viridis)
+source("R/plotting_functions.R")
+
+
+
+#### 01 Get climate data for Australia ####
+# Download bioclim data using library (raster)
+au_map<-raster("data/au_binary.tif") %>% 
+  aggregate(fact=6) %>% 
+  as.data.frame(xy=T)
+
+
+#### 02 Plot AU basemap ####
+ggplot()+
+  geom_raster(data=au_map, aes(x = x, y = y, fill = au_binary))+
+  scale_fill_identity() +theme_minimal() +xlab("")+ylab("")-> au_basemap
+
+#### 03 Overlay species occurences ####
+sites<-read_csv("data/austraits_sites.csv")
+au_basemap+ geom_point(data=sites, aes(y=`latitude (deg)`, x=`longitude (deg)`), color="grey", alpha=0.5)
+
+

### 2019-04-08 R/fig_climate_space_austraits.R
figure rearrangement

@@ -15,7 +15,7 @@ names(bioclim) <- c("Temp","Prec")
 
 #### 02 Get the climate data for Australia ####
 # Load Australia landmass binary map
-au_map<-raster("data/au_binary.tif")
+au_map<-raster("data/australia.tif")
 # Clip bioclim data with the au map
 ## crop and mask
 bioclim_au <- crop(bioclim, extent(au_map)) #%>% mask(.,au_map)
@@ -26,7 +26,7 @@ au_bioclim<-mask(new.bioclim, au_map)
 au_bioclim_table<-au_bioclim %>% 
   as.data.frame() %>%
   na.omit() %>%
-  as_tibble()
+  as_tibble %>% mutate(region="Australia")
 
 #### 03 Get the point coordinates of Austraits data ####t
 
@@ -51,6 +51,8 @@ left_join(sites,
           by = c("dataset_id" = "dataset_id", 
                  "site_name" = "site_name"))->combined_sites
 
+write_csv(combined_sites, "data/austraits_sites.csv")
+
 # Get climate values for each site
 au_sites_clim<-combined_sites %>% rowid_to_column("ID") %>% 
   rename(latitude=`latitude (deg)`, longitude=`longitude (deg)`) %>% 
@@ -61,35 +63,53 @@ au_sites_clim<-combined_sites %>% rowid_to_column("ID") %>%
 au_sites_clim
 au_bioclim_table
 
-# Option 1: hexagonal background + dot points (size ~ species number)
-austraits_climate_space<-ggplot(au_bioclim_table %>% sample_n(1e6), 
-  aes(x=Prec/10, y=Temp/10)) +
-  geom_hex(alpha=0.8) +
-  theme_minimal()+
-  xlab("Precipitation (cm)")+
-  ylab("Temperature (deg C)")+
-  scale_fill_viridis()+
-  geom_point(data = au_sites_clim, 
-             aes(x=Prec/10, y=Temp/10), 
-             alpha=0.5,shape=1,
-             size=2*log2(au_sites_clim$n),color="black",
-             inherit.aes = FALSE)+ 
-  theme(legend.position = "none") 
-ggsave(austraits_climate_space, "austraits_climate_space.png", width = 4, height = 4)
 
-austraits_climate_space
+# # Option 1: hexagonal background + dot points (size ~ species number)
+# austraits_climate_space<-ggplot(au_bioclim_table, 
[…]

### 2019-04-08 R/fig_site_locations.R

@@ -1,27 +1,53 @@
-## Austraits: Precipitation–temperature space plot
+## Austraits: site locations plot
+
 library(raster)
 library(tidyverse)
-library(hexbin)
 library(ggplot2)
-library(viridis)
+library(RColorBrewer)
 source("R/plotting_functions.R")
 
 
 
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)
-au_map<-raster("data/au_binary.tif") %>% 
-  aggregate(fact=6) %>% 
+au_map<-raster("data/australia.tif") %>% 
+ # aggregate(fact=6) %>% 
   as.data.frame(xy=T)
 
 
 #### 02 Plot AU basemap ####
 ggplot()+
-  geom_raster(data=au_map, aes(x = x, y = y, fill = au_binary))+
-  scale_fill_identity() +theme_minimal() +xlab("")+ylab("")-> au_basemap
+  geom_raster(data=au_map, aes(x = x, y = y, fill=factor(australia)))+
+  scale_fill_brewer(name="",type = "seq", palette = "#800000", guide=F)+theme_minimal() +xlab("")+ylab("")-> au_basemap
+au_basemap
 
 #### 03 Overlay species occurences ####
 sites<-read_csv("data/austraits_sites.csv")
-au_basemap+ geom_point(data=sites, aes(y=`latitude (deg)`, x=`longitude (deg)`), color="grey", alpha=0.5)
+austraits_site_locations<-au_basemap+
+  geom_point(data=sites, aes(y=`latitude (deg)`,
+                              x=`longitude (deg)`, size=n), 
+              alpha=0.3, inherit.aes = F)+
+  scale_size(name   = "N Species",
+             breaks = sqrt(fivenum(sites$n)),
+             labels = fivenum(sites$n))+
+  theme( legend.position = "bottom", legend.direction  = "horizontal") 
+
+austraits_site_locations
 
 
+ggsave("austraits_sites_distribution.png",
+       austraits_site_locations,  
+       height=4, width=4, units="in")
+
+
+austraits_site_locations<-au_basemap+
+  geom_jitter(data=sites, aes(y=`latitude (deg)`,
+                              x=`longitude (deg)`, color=n), 
+              alpha=0.3, inherit.aes = F, size=2)+
+  scale_color_gradient(name=" N Species", trans="log10", low='#F4A460', high='#800000') +
[…]

### 2019-04-08 R/fig_2.R
Figure 2 v1

@@ -1,0 +1,21 @@
+## Austraits: generate figure 2
+
+library(patchwork)
+
+source("R/fig_site_locations.R")
+source("R/fig_climate_space_austraits.R")
+source("R/fig_trait_composition.R")
+
+
+combined_plot<-(austraits_site_locations+
+                  austraits_climate_space)-
+                  {
+                    austraits_composition +
+                      plot_spacer()+
+                      plot_layout(widths = c(14, 1)) 
+                  }+
+  plot_layout(ncol = 1, heights = c(3, 1))
+
+ggsave("austraits_fig2.png",
+       combined_plot,  
+       height=6, width=12, units="in")

### 2019-04-08 R/fig_climate_space_austraits.R

@@ -59,9 +59,9 @@ au_sites_clim<-combined_sites %>% rowid_to_column("ID") %>%
   combine_occurence_climate(au_bioclim)
 
 
-#### 03 Plot the climate data ####
-au_sites_clim
-au_bioclim_table
+# #### 03 Plot the climate data ####
+# au_sites_clim
+# au_bioclim_table
 
 
 # # Option 1: hexagonal background + dot points (size ~ species number)
@@ -82,9 +82,9 @@ au_bioclim_table
 # 
 # austraits_climate_space
 
-# Option 2: Add whittaker plot
-au_bioclim_table %>%
-  sample_n(1e4) %>% group_by(region)->bioclim_small #sample data points to reduce computing load
+# # Option 2: Add whittaker plot
+# au_bioclim_table %>%
+#   sample_n(1e4) %>% group_by(region)->bioclim_small #sample data points to reduce computing load
 
 library(plotbiomes)
 library("RColorBrewer")
@@ -99,17 +99,17 @@ austraits_climate_space<-whittaker_base_plot() +
                alpha=0.5,
                size=log10(au_sites_clim$n),
                inherit.aes = FALSE)+
-  scale_colour_manual(name="",values = c("black", "lightgrey"))+
+  scale_colour_manual(name="",values = c("lightgrey","black"))+
   theme_bw() 
 
-austraits_climate_space
-
-ggsave("austraits_climate_space.png",
-       austraits_climate_space,  
-       height=4, width=8, units="in")
-
-
-# combine species number vs sites
-species_number_lookup %>% na.omit() %>%  
-  anti_join(sites, 
-            by = c("site_name" = "site_name"))->mismatch_sites
+#austraits_climate_space
+# 
+# ggsave("austraits_climate_space.png",
+#        austraits_climate_space,  category,
+#        height=4, width=8, units="in")
+# # 
+# 
+# # combine species number vs sites
+# species_number_lookup %>% na.omit() %>%  
+#   anti_join(sites, 
+#             by = c("site_name" = "site_name"))->mismatch_sites

### 2019-04-08 R/fig_site_locations.R

@@ -3,7 +3,7 @@
 library(raster)
 library(tidyverse)
 library(ggplot2)
-library(RColorBrewer)
+
 source("R/plotting_functions.R")
 
 
@@ -23,31 +23,32 @@ au_basemap
 
 #### 03 Overlay species occurences ####
 sites<-read_csv("data/austraits_sites.csv")
+# austraits_site_locations<-au_basemap+
+#   geom_point(data=sites, aes(y=`latitude (deg)`,
+#                               x=`longitude (deg)`, size=n), 
+#               alpha=0.3, inherit.aes = F)+
+#   scale_size(name   = "N Species",
+#              breaks = sqrt(fivenum(sites$n)),
+#              labels = fivenum(sites$n))+
+#   theme( legend.position = "bottom", legend.direction  = "horizontal") 
+# 
+# austraits_site_locations
+# 
+# 
+# ggsave("austraits_sites_distribution.png",
+#        austraits_site_locations,  
+#        height=4, width=4, units="in")
+
+
 austraits_site_locations<-au_basemap+
   geom_point(data=sites, aes(y=`latitude (deg)`,
-                              x=`longitude (deg)`, size=n), 
-              alpha=0.3, inherit.aes = F)+
-  scale_size(name   = "N Species",
-             breaks = sqrt(fivenum(sites$n)),
-             labels = fivenum(sites$n))+
-  theme( legend.position = "bottom", legend.direction  = "horizontal") 
-
-austraits_site_locations
-
-
-ggsave("austraits_sites_distribution.png",
-       austraits_site_locations,  
-       height=4, width=4, units="in")
-
-
-austraits_site_locations<-au_basemap+
-  geom_jitter(data=sites, aes(y=`latitude (deg)`,
                               x=`longitude (deg)`, color=n), 
               alpha=0.3, inherit.aes = F, size=2)+
   scale_color_gradient(name=" N Species", trans="log10", low='#F4A460', high='#800000') +
   theme(legend.justification = c(0, 0), legend.position = c(0, 0), legend.direction  = "horizontal") 
-austraits_site_locations
 
-ggsave("austraits_sites_distribution_v2.png",
-       austraits_site_locations,  
-       height=4, width=5, units="in")
+
+# ggsave("austraits_sites_distribution_v2.png",
+#        austraits_site_locations,  
+#        height=4, width=5, units="in")
+

### 2019-04-08 R/fig_trait_composition.R

@@ -2,24 +2,36 @@
 
 library(tidyverse)
 library(ggplot2)
+library(scales)
+library(RColorBrewer)
 source("R/plotting_functions.R")
 
 austraits<-readRDS("data/austraits.rds")
 trait_category_lookup<-read_csv("data/trait_categories.csv")
 
 austraits$traits %>% 
-  left_join(trait_category_lookup,by="trait_name") %>%
-  group_by(tissue, category) %>%
-  summarise(n = n()) -> summary_n_species
+  left_join(trait_category_lookup,by="trait_name")  %>% mutate(category = str_replace(category, "physiology", "physiological"))%>%
+  group_by(tissue, category) %>% 
+  summarise(n = n()) %>% rename(Category=category)-> summary_n_species
 
 
 
 ## Change the order of the 
-ggplot()+
-  geom_bar(data=summary_n_species, 
-           aes(x=tissue, y=n, fill=category), 
-           stat='identity', position = position_dodge(preserve = "total")) +
-  theme_classic()+ 
-  scale_y_log10()+
+positions <- c("leaf", "stem", "root", "reproductive", "whole_plant")
+cbPalette <- brewer.pal(4,"Pastel2")
 
+ggplot(data=summary_n_species, 
+       aes(x=tissue, y=n, fill=Category))+
+  geom_bar( stat='identity', position =position_dodge(preserve = "single")) +
+  theme_classic()+ 
+  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
+                labels = trans_format("log10", math_format(10^.x)))+
+  xlab("Tissue")+ ylab("N Species")+ scale_x_discrete(limits = positions)+ 
+  theme(legend.position = "right") +
+  scale_fill_manual(values=cbPalette)->austraits_composition
+
+# austraits_composition
+# ggsave("austraits_composition.png",
+#        austraits_composition,  
+#        height=5, width=8, units="in")
 

### 2019-04-09 R/fig_climate_space_austraits.R
Figure 2 v2

@@ -100,14 +100,14 @@ austraits_climate_space<-whittaker_base_plot() +
                size=log10(au_sites_clim$n),
                inherit.aes = FALSE)+
   scale_colour_manual(name="",values = c("lightgrey","black"))+
-  theme_bw() 
+  theme_classic()
 
 #austraits_climate_space
 # 
-# ggsave("austraits_climate_space.png",
-#        austraits_climate_space,  category,
-#        height=4, width=8, units="in")
-# # 
+ggsave("austraits_climate_space.png",
+       austraits_climate_space, 
+       height=4, width=8, units="in")
+#
 # 
 # # combine species number vs sites
 # species_number_lookup %>% na.omit() %>%  

### 2019-04-09 R/fig_site_locations.R

@@ -3,7 +3,7 @@
 library(raster)
 library(tidyverse)
 library(ggplot2)
-
+library(RColorBrewer)
 source("R/plotting_functions.R")
 
 
@@ -18,37 +18,37 @@ au_map<-raster("data/australia.tif") %>%
 #### 02 Plot AU basemap ####
 ggplot()+
   geom_raster(data=au_map, aes(x = x, y = y, fill=factor(australia)))+
-  scale_fill_brewer(name="",type = "seq", palette = "#800000", guide=F)+theme_minimal() +xlab("")+ylab("")-> au_basemap
-au_basemap
+  scale_fill_brewer(name="",type = "seq", palette = "#800000", guide=F)+xlab("")+ylab("")+ myTheme-> au_basemap
+#au_basemap
 
-#### 03 Overlay species occurences ####
+#### 03 Overlay site locations  ####
 sites<-read_csv("data/austraits_sites.csv")
-# austraits_site_locations<-au_basemap+
-#   geom_point(data=sites, aes(y=`latitude (deg)`,
-#                               x=`longitude (deg)`, size=n), 
-#               alpha=0.3, inherit.aes = F)+
-#   scale_size(name   = "N Species",
-#              breaks = sqrt(fivenum(sites$n)),
-#              labels = fivenum(sites$n))+
-#   theme( legend.position = "bottom", legend.direction  = "horizontal") 
-# 
+austraits_site_locations<-au_basemap+
+  geom_point(data=sites, aes(y=`latitude (deg)`,
+                              x=`longitude (deg)`, size=n),
+              alpha=0.3, inherit.aes = F)+
+  scale_size(name   = "N Species",
+             breaks = sqrt(fivenum(sites$n)),
+             labels = fivenum(sites$n))
+
 # austraits_site_locations
 # 
-# 
+# # #### 04 Save plots ####
 # ggsave("austraits_sites_distribution.png",
 #        austraits_site_locations,  
-#        height=4, width=4, units="in")
+#        height=4, width=6, units="in")
 
 
+
+#### 03 Overlay site locations  ####
 austraits_site_locations<-au_basemap+
   geom_point(data=sites, aes(y=`latitude (deg)`,
                               x=`longitude (deg)`, color=n), 
               alpha=0.3, inherit.aes = F, size=2)+
-  scale_color_gradient(name=" N Species", trans="log10", low='#F4A460', high='#800000') +
[…]

### 2019-04-09 R/fig_trait_composition.R

@@ -26,7 +26,7 @@ ggplot(data=summary_n_species,
   theme_classic()+ 
   scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                 labels = trans_format("log10", math_format(10^.x)))+
-  xlab("Tissue")+ ylab("N Species")+ scale_x_discrete(limits = positions)+ 
+  xlab("Major plant part")+ ylab("Species (n)")+ scale_x_discrete(limits = positions)+ 
   theme(legend.position = "right") +
   scale_fill_manual(values=cbPalette)->austraits_composition
 

### 2019-04-09 R/plotting_functions.R

@@ -35,3 +35,38 @@ combine_occurence_climate<-function(species_occurence_df, climate_raster_data){
   
   sp_clim_combined
 }
+
+
+myTheme <-
+  
+  theme(legend.position = "none",
+        
+        panel.grid.major = element_blank(),
+        
+        panel.grid.minor = element_blank(),
+        
+        panel.background = element_blank(),
+        
+        panel.border = element_rect(colour = "black", fill=NA, size=1),
+        
+        axis.ticks.length = unit(1, "mm"),
+        
+        axis.ticks = element_line(size = 1),
+        
+        axis.line.x = element_line(),
+        
+        axis.line.y = element_line(),
+        
+        axis.title.x = element_text( size = 12, margin = margin(10, 0, 0, 0)),
+        
+        axis.title.y = element_text( size = 12, margin = margin(0, 10, 0, 0)),
+        
+        axis.text.x = element_text( size = 9, colour = "#666666", margin=margin(15,0,0,0,"pt")),
+        
+        axis.text.y = element_text( size = 9, colour = "#666666", margin=margin(0,15,0,0,"pt")),
+        
+        panel.spacing.y = unit(1, "points"),
+        
+        legend.title=element_text(size=12),
+        
+        legend.text=element_text(size=10))

### 2019-04-09 R/fig_ggtree.R
ggtree trial

@@ -1,0 +1,66 @@
+##  The phylogenetic coverage of traits in the Austraits, 
+#for the subset of species in the current molecular phylogeny.
+library(ggtree)
+library(treeio)
+library(ggplot2)
+library(tidyverse)
+library(stringi)
+library(taxonlookup)
+source("R/taxonlookup.R")
+
+# read a BEAST formatted tree:  Maximum clade credibility tree from the BEAST analysis of the B series
+tr<-read.beast(file = url("https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5543309/bin/ncomms16047-s5.txt"))
+
+# the latest molecular phylogeny: Ecography, DOI: 10.1111/ecog.04434
+#load(url("https://github.com/jinyizju/V.PhyloMaker/blob/master/data/GBOTB.extended.rda"))
+
+# The get.fields method return all available features that can be used for annotation.
+#get.fields(tr)
+#get.fields(GBOTB.extended)
+
+#tip_label<-GBOTB.extended$tip.label
+tip_label<-get.tree(tr)$tip.label
+
+get.tree(tr)$tip.label
+## extract table column
+tip_label %>% 
+  tibble::enframe()->tr_tiplabel #74,531 species
+
+## replace "_" with a white-space
+tr_tiplabel$value<-stri_replace_all_fixed(tr_tiplabel$value, "_", " ")
+
+## add taxon lookup
+tr_tiplabel<-tr_tiplabel %>% rename(species_name=value) %>% add_taxon_lookup()
+
+## add austraits and create a lookup table
+austraits$traits %>% 
+  left_join(trait_category_lookup,by="trait_name")  %>% mutate(category = str_replace(category, "physiology", "physiological"))%>%
+  group_by(species_name,tissue) %>% 
+  summarise(n = n()) %>% spread(key=tissue, value=n)->austraits_summary # 20,593 species
+
+
+#How many species are matched? 6,552 species
+tr_tiplabel %>%  inner_join(austraits_summary, by="species_name")
+# reference lookup
+tr_tiplabel %>%  left_join(austraits_summary,by="species_name") %>% select(-name)-> austraits_tree_lookup
+
+## replace  white-space with a "_"
+austraits_tree_lookup$species_name<-stri_replace_all_fixed(austraits_tree_lookup$species_name, " ", "_")
+
+
[…]

### 2019-04-09 R/taxonlookup.R

@@ -1,0 +1,14 @@
+taxon_lookup_func<-function(df){
+  df[["species_name"]]%>%
+    unique()%>%
+    lookup_table(., by_species=TRUE)%>%
+    tibble::rownames_to_column(., var = "species_name")-> taxon_result
+  
+  taxon_result
+}
+#function to join main data with the taxon_lookup_func outcome
+add_taxon_lookup<-function(df){
+  taxon_lookup_func(df)%>%
+    left_join(df, ., by="species_name")
+  
+}

### 2019-04-10 ms.Rmd
Reworking formatting of citations
- now using csl file to generate citations in proper nature style
- at the same time, lots of cleaning in austraits data too (see https://github.com/traitecoevo/austraits/issues/176)

--- 2019-04-04 ms.Rmd
@@
 ---
-title: "Austraits – a curated plant trait database for the Australian flora"
-bibliography: references.bib
+title: "Austraits -- a curated plant trait database for the Australian flora"
 csl: assets/nature.csl
+editor_options:
+  chunk_output_type: console
 output:
@@
     reference_docx: assets/manuscriptstyle.docx
-editor_options:
-  chunk_output_type: console
+bibliography: references.bib
 ---
@@
 
+```{r austraits}
+
+## Assumes to items exist in global name space
+austraits <- readRDS("data/austraits.rds")
+
+definitions <- austraits$definitions
+
+tmp <- RefManageR::WriteBib(austraits$sources, "references.bib")
+
+sources_all <- 
+  bind_rows(
+    austraits$methods %>% select(dataset_id, citation = source_primary_key) %>% 
+            distinct() %>% mutate(type="primary"),
+    austraits$methods %>% select(dataset_id, citation = source_secondary_key) %>% 
+            distinct() %>% mutate(type="secondary")
+  ) %>% 
+  na.omit() %>% 
+  arrange(dataset_id) %>% 
+  group_by(dataset_id) %>% 
+  summarise(refs = sprintf("@%s", citation) %>% 
+  paste(collapse = "; "))
+
+```
+
 # Overview
@@
 # Data structures
+
+# Figures
+
+![**Figure 1.
+The data curation pathway used to assemble the AusTraits database.** Trait observations are accessed from original data sources, including published floras and field campaigns.
+Factors such as units and taxonomy are harmonised to a common standard.
+Versioned releases are distributed to users, allowing the dataset to be used and re-used in a reproducible way.
+Errors are corrected and new datasets added via the blue path for sequential releases of the database.](figures/Pipeline.png)
+
+![**Figure 2.** (a.) Locations where the trait data were collected and the corresponding number of species.
+(b.) AusTraits’ sites (grey) within Australia’s precipitation-temperature space (black) superimposed upon Whittaker’s major biomes diagram.
+Data from WorldClim (worldclim.org).
+(c.) Overview of the number of species with which traits are measured within each major plant organ and trait category.](figures/austraits_fig2_blue.png)
 
@@
 
+trait_categories <- read_csv("data/trait_categories.csv", col_types = cols(.default = "c"))
+
 data_geo <- 
@@
   left_join(by="species_name",
-      select(austraits$taxonomy, species_name, family))
+      select(austraits$taxonomy, species_name, family)) %>% 
+  left_join(sources_all, by=c("dataset_id"))
 
@@
             Species = n_distinct(species_name),
-            Families = n_distinct(family)
-            )
+            Families = n_distinct(family),
+            `Data Citations` = refs %>% unique() %>% paste(collapse = "; ")
+            ) %>% 
+  left_join(trait_categories, by=c("trait_name")) %>% 
+  select(Tissue = tissue, Category = category, Trait = trait_name, everything()) %>% 
+  arrange(Tissue, Category, Trait)
 
-n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$trait_name]
-n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$trait_name]
+n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$Trait]
+n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$Trait]
 
 n_records %>%
-  arrange(desc(`Records (all)`)) %>% 
-  slice(1:40) %>%
+  select(-Tissue, Description) %>%
   kable()
-
 ```
@@
 
-```{r, results='asis'}
-c(austraits$methods$source_primary_citation %>% unique(),
-austraits$methods$source_secondary_citation %>% unique()) %>% paste(collapse = "\n\n") %>% writeLines()
+## Version 1 - via bib.print
+
+```{r, results='asis', eval=TRUE}
+sources_unique <-
+  bind_rows(
+    austraits$methods %>% select(citation = source_primary_citation)  %>% mutate(type="primary"),
+    austraits$methods %>% select(dataset_id, citation = source_secondary_citation) %>% mutate(type="secondary")
+  ) %>%
+  select(citation) %>%
+            distinct() %>% 
+            mutate(source_number=1:n())
+
+sprintf("%s. %s", sources_unique$source_number, sources_unique$citation) %>% paste(collapse = "\n\n") %>% writeLines()
+
+```
+
+## Version 3 - via csl
+
+```{r, results='asis', eval=FALSE}
+
+austraits$sources %>% sapply(function(i) i$key) -> keys
+
+sprintf("@%s",keys) %>% paste(collapse = "; ") %>% writeLines()
 ```

### 2019-04-10 R/fig_ggtree.R
Some suggested changes on code

@@ -1,3 +1,11 @@
+# Need to Install taxonlookup from XXX
+
+# Need to Install ggtree from https://guangchuangyu.github.io/software/ggtree/
+# this might not be so straight forward, as lots of dependencies.
+# source("https://bioconductor.org/biocLite.R")
+# biocLite("ggtree")
+
+
 ##  The phylogenetic coverage of traits in the Austraits, 
 #for the subset of species in the current molecular phylogeny.
 library(ggtree)
@@ -7,6 +15,11 @@ library(tidyverse)
 library(stringi)
 library(taxonlookup)
 source("R/taxonlookup.R")
+
+
+# load austraits 
+trait_category_lookup<-read_csv("data/trait_categories.csv")
+austraits <- readRDS("data/austraits.rds")
 
 # read a BEAST formatted tree:  Maximum clade credibility tree from the BEAST analysis of the B series
 tr<-read.beast(file = url("https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5543309/bin/ncomms16047-s5.txt"))
@@ -24,13 +37,13 @@ tip_label<-get.tree(tr)$tip.label
 get.tree(tr)$tip.label
 ## extract table column
 tip_label %>% 
-  tibble::enframe()->tr_tiplabel #74,531 species
+  tibble::enframe() -> tr_tiplabel #74,531 species
 
 ## replace "_" with a white-space
-tr_tiplabel$value<-stri_replace_all_fixed(tr_tiplabel$value, "_", " ")
+tr_tiplabel$value <- stri_replace_all_fixed(tr_tiplabel$value, "_", " ")
 
 ## add taxon lookup
-tr_tiplabel<-tr_tiplabel %>% rename(species_name=value) %>% add_taxon_lookup()
+tr_tiplabel <- tr_tiplabel %>% rename(species_name=value) %>% add_taxon_lookup()
 
 ## add austraits and create a lookup table
 austraits$traits %>% 
@@ -45,7 +58,7 @@ tr_tiplabel %>%  inner_join(austraits_summary, by="species_name")
 tr_tiplabel %>%  left_join(austraits_summary,by="species_name") %>% select(-name)-> austraits_tree_lookup
 
 ## replace  white-space with a "_"
-austraits_tree_lookup$species_name<-stri_replace_all_fixed(austraits_tree_lookup$species_name, " ", "_")
[…]

### 2019-04-10 R/fig_site_locations.R

@@ -10,7 +10,14 @@ source("R/plotting_functions.R")
 
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)
-au_map<-raster("data/australia.tif") %>% 
+
+# one strategy for retrieving data available on web
+filename <- "data/australia.tif"
+if(!file.exists(filename))
+  download.file("htpp:....", filename)
+
+# Read bioclim data using raster package
+au_map<-raster(filename) %>% 
  # aggregate(fact=6) %>% 
   as.data.frame(xy=T)
 

### 2019-04-10 R/fig_2.R
re-arranged r scripts for generating figures
- create a new folder named 'scripts' for generating the subfigures
- R folder remains for the functions
- built_figures.R is the wrapper script

@@ -1,21 +1,0 @@
-## Austraits: generate figure 2
-
-library(patchwork)
-
-source("R/fig_site_locations.R")
-source("R/fig_climate_space_austraits.R")
-source("R/fig_trait_composition.R")
-
-
-combined_plot<-(austraits_site_locations+
-                  austraits_climate_space)-
-                  {
-                    austraits_composition +
-                      plot_spacer()+
-                      plot_layout(widths = c(14, 1)) 
-                  }+
-  plot_layout(ncol = 1, heights = c(3, 1))
-
-ggsave("austraits_fig2.png",
-       combined_plot,  
-       height=6, width=12, units="in")

### 2019-04-10 built_figures.R

@@ -1,0 +1,26 @@
+
+## Austraits: Figure 2
+
+library(patchwork)
+
+source("scripts/fig_site_locations.R") # A
+source("scripts/fig_climate_space_austraits.R") # B
+source("scripts/fig_trait_composition.R") # C
+
+
+combined_plot<-(austraits_site_locations+
+                  austraits_climate_space)-
+                  {
+                    austraits_composition +
+                      plot_spacer()+
+                      plot_layout(widths = c(14, 1)) 
+                  }+
+  plot_layout(ncol = 1, heights = c(3, 1))
+
+
+pdf("figures/austraits_fig2.pdf",height=6, width=12)
+combined_plot
+dev.off()
+
+# ggsave("figures/austraits_fig2.png",
+#        combined_plot,  height=6, width=12, units="in")

### 2019-04-10 scripts/fig_climate_space_austraits.R

@@ -91,11 +91,11 @@ library("RColorBrewer")
 austraits_climate_space<-whittaker_base_plot() +
     # add the temperature - precipitation data points
     geom_point(data =au_bioclim_table, 
-             aes(x=Temp/10, y=Prec/10, color="Australia"), alpha=0.1,
+             aes(x=Temp/10, y=Prec/10, color="Australia"),shape = ".", alpha=0.1,
              inherit.aes = FALSE) + 
   
     geom_point(data = au_sites_clim, 
-               aes(x=Temp/10, y=Prec/10, color="Austraits sites"), 
+               aes(x=Temp/10, y=Prec/10, color="AusTraits sites"), 
                alpha=0.5,
                size=log10(au_sites_clim$n),
                inherit.aes = FALSE)+
@@ -103,11 +103,11 @@ austraits_climate_space<-whittaker_base_plot() +
   theme_classic()
 
 #austraits_climate_space
-# 
-ggsave("austraits_climate_space.png",
-       austraits_climate_space, 
-       height=4, width=8, units="in")
-#
+# # 
+# ggsave("austraits_climate_space.png",
+#        austraits_climate_space, 
+#        height=4, width=8, units="in")
+# #
 # 
 # # combine species number vs sites
 # species_number_lookup %>% na.omit() %>%  

### 2019-04-10 scripts/fig_site_locations.R

@@ -1,0 +1,54 @@
+## Austraits: site locations plot
+
+library(raster)
+library(tidyverse)
+library(ggplot2)
+library(RColorBrewer)
+source("R/plotting_functions.R")
+
+
+
+#### 01 Get climate data for Australia ####
+# Download bioclim data using library (raster)
+au_map<-raster("data/australia.tif") %>% 
+ # aggregate(fact=6) %>% 
+  as.data.frame(xy=T)
+
+
+#### 02 Plot AU basemap ####
+ggplot()+
+  geom_raster(data=au_map, aes(x = x, y = y, fill=factor(australia)))+
+  scale_fill_brewer(name="",type = "seq", palette = "#800000", guide=F)+xlab("")+ylab("")+ myTheme-> au_basemap
+#au_basemap
+
+#### 03 Overlay site locations  ####
+sites<-read_csv("data/austraits_sites.csv")
+austraits_site_locations<-au_basemap+
+  geom_point(data=sites, aes(y=`latitude (deg)`,
+                              x=`longitude (deg)`, size=n),
+              alpha=0.3, inherit.aes = F)+
+  scale_size(name   = "N Species",
+             breaks = sqrt(fivenum(sites$n)),
+             labels = fivenum(sites$n))
+
+# austraits_site_locations
+# 
+# # #### 04 Save plots ####
+# ggsave("austraits_sites_distribution.png",
+#        austraits_site_locations,  
+#        height=4, width=6, units="in")
+
+
+
+#### 03 Overlay site locations  ####
+austraits_site_locations<-au_basemap+
+  geom_point(data=sites, aes(y=`latitude (deg)`,
+                              x=`longitude (deg)`, color=n), 
+              alpha=0.3, inherit.aes = F, size=2)+
+  scale_color_gradient(name="Species (n)", trans="log10", low='#F4A460', high='#800000') +
+  theme(legend.justification = c(0, 0), legend.position = c(0.05, 0.05), legend.direction  = "horizontal")
+austraits_site_locations
+
+# ggsave("austraits_sites_distribution_v2.png",
+#        austraits_site_locations,  
+#        height=4, width=5, units="in")

### 2019-04-29 R/plotting_functions.R
swap the MRCA function from ggtree::MRCA to ape::getMRCA

@@ -85,15 +85,17 @@ get_node_tree<-function(x, t=tree$scenario.2$run.1){
   x[2]%>%
     unlist %>%
     rlang::parse_expr()%>%
-    rlang::eval_tidy()->y
+    rlang::eval_tidy()->y # is a vector of tip (species)
   
-  if(length(y)>10){
-    MRCA(t, y)->z
+  if(length(y)>10){ # select family with >10 species
+    #browser()
+    ape::getMRCA(t, y)->z # z is a node of the most recent common ancestor 
     mutate(x, node=z) %>% as.data.frame()
   } else
   {
     # parent(as_tibble(t),y) %>% select(node) %>% unlist()->z
-    mutate(x, node=NA) %>% as.data.frame() # z is the parent node, set into NA if you want to omit clade with only one species,
+    mutate(x, node=NA) %>%
+      as.data.frame() # z is the parent node, set into NA if you want to omit clade with only one species,
   }
   
 }

### 2019-04-29 scripts/fig_get_family_node.R
escape the quotation marks with backslashes

@@ -11,7 +11,7 @@ austraits_sp_list_sampled %>%
   filter(species_name %in% tree$scenario.2$run.1$tip.label) %>% 
   group_by(family) %>%
   summarise(species_name = paste(
-    paste0("'",species_name, "'"),collapse=", ")) %>% 
+    paste0("\"",species_name, "\""),collapse=", ")) %>% 
   mutate(species_name=paste0('c(',species_name, ')'))->family_list
 
 #### 2 Calculate the most recent common ancestor to the node for each family ####

### 2019-04-29 scripts/fig_trait_composition.R
fig 2 revision
-fixed the wrong species number on fig2c

@@ -12,13 +12,13 @@ trait_category_lookup<-read_csv("data/trait_categories.csv")
 austraits$traits %>% 
   left_join(trait_category_lookup,by="trait_name")  %>% mutate(category = str_replace(category, "physiology", "physiological"))%>%
   group_by(tissue, category) %>% 
-  summarise(n = n()) %>% rename(Category=category)-> summary_n_species
+  summarise(n = n_distinct(species_name)) %>% rename(Category=category) %>% drop_na()-> summary_n_species
 
 
 
 ## Change the order of the 
 positions <- c("leaf", "stem", "root", "reproductive", "whole_plant")
-cbPalette <- brewer.pal(4,"Pastel2")
+cbPalette <- brewer.pal(5,"Blues")[2:5]
 
 ggplot(data=summary_n_species, 
        aes(x=tissue, y=n, fill=Category))+

### 2019-04-29 scripts/fig_generate_phylogenies.R
Minor fixes

@@ -14,6 +14,8 @@ library(tidyverse)
 library(ggtree)
 library(treeio)
 library(ggplot2)
+library(tidytree)
+
 
 
 

### 2019-04-29 ms.Rmd
Figure 3 caption

@@
 (c.) Overview of the number of species with which traits are measured within each major plant organ and trait category.](figures/austraits_fig2_blue.png)
+
+![**Figure 3.** The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled species.
+The heatmap colour intensities denote the number of traits measured within each major plant organ.
+The most widespread family names (with more than ten species) are labelled on the edge of the tree.](figures/austraits_phlogenetic_coverage.png)
+
+<!-- Cite: -->
+  <!-- G Yu, DK Smith, H Zhu, Y Guan, TTY Lam*.
+ggtree: an R package for visualization and annotation of phylogenetic trees with their covariates and other associated data.
+Methods in Ecology and Evolution.
+2017, 8(1):28-36.
+doi: 10.1111/2041-210X.12628.
+-->
+  <!-- G Yu*, TTY Lam, H Zhu, Y Guan*.
+Two methods for mapping and visualizing associated data on phylogeny using ggtree.
+Molecular Biology and Evolution, 2018, 35(2):3041-3043.
+doi: 10.1093/molbev/msy194.
+-->
 

### 2019-04-29 build_figures.R
rename file

@@ -24,8 +24,6 @@ dev.off()
 
 
 
-
-
 #### Austraits: Phylogenetic coverage of traits ####
 source("scripts/fig_generate_phylogenies.R") # A
 source("scripts/fig_get_family_node.R") # B
@@ -65,8 +63,6 @@ for(j in 1:length(nodes_list)){
 }
 
 
-
-
 austraits_tree <-  gheatmap(base_tree, austraits_summary, 
                 offset=4, 
                 width=0.2, 
@@ -80,6 +76,6 @@ austraits_tree <-  gheatmap(base_tree, austraits_summary,
                 font.size=3,
                 hjust=1)
 
-pdf("figures/austraits_phlogenetic_coverage.pdf", height=10, width=10)
+pdf("figures/austraits_phlogenetic_coverage2.pdf", height=10, width=10)
 austraits_tree
 dev.off()

### 2019-12-03 ms.Rmd
transferred manuscript from msword

@@
 output:
+  html_document:
+    df_print: paged
+  pdf_document:
+    latex_engine: xelatex
   word_document:
     reference_docx: assets/manuscriptstyle.docx
+header-includes:
+- \usepackage{inputenc}
+- \usepackage[T1]{fontenc}
 bibliography: references.bib
@@
 
+### Abstract
+
+We introduce the AusTraits database - a compilation of measurements of plant traits for species in the Australian Flora (hereafter AusTraits).
+AusTraits synthesises data on 86 traits across 20,321 species from field campaigns, published literature, taxonomic monographs, and individual species descriptions.
+Traits
+vary in scope from physiological measures of performance (e.g. photosynthetic gas exchange, water-use efficiency) to morphological
+parameters (e.g. leaf area, seed mass, plant height) which link to aspects of ecological variation. AusTraits contains curated and
+harmonised individual-, species- and genus-level observations coupled to, where available, contextual information on site
+properties. This data descriptor provides information on Version 0.0.9 of AusTraits which contains data for 325,819 trait-by-species combinations across 20,321 species. We envision AusTraits as an on-
+going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
+
+## Background and Summary
+
+Species traits are essential metrics for comparing ecological strategies in plants arrayed across environmental space or evolutionary lineages1-4.
+Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function5-7.
+Traits thereby provide useful indicators of species behaviours in communities and ecosystems, regardless of their taxonomy8,9.
+Through targeted global initiatives the volume of available trait information for plants has grown rapidly in the last two decades.
+However, the geographic coverage of trait observations across the globe is patchy, limiting detailed analyses of trait variation and diversity in some regions.
+One such region is Australia; a continent with a flora of c.
+24,000 native higher-plant species10.
+While significant investment has been made in curating and digitising herbarium collections in Australia over the last two decades (e.g.
+The Australian Virtual Herbarium houses ~7 million specimen occurrence records; https://avh.ala.org.au), no complementary resource yet exists for consolidating information on plant traits.
+
+Here we introduce the AusTraits database (hereafter simply AusTraits) - a compilation of plant traits for the Australian flora.
+Currently, AusTraits draws together 148 primary sources – including published scientific papers, floras, and taxonomic species descriptions– and contains 325,819 measurements spread across 86 different traits for 20,321 species (Table 1; available online only).
+By providing a harmonised and curated dataset on 86 plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
+Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives (but see 11).
+Global initiatives, such as the TRY database12 and Open Traits Network, have been working towards sharing of data about plant traits though large numbers of Australian plant species remain unrepresented.
+
+To assemble AusTraits from diverse primary sources and make it available for reuse, we needed to overcome three main types of challenges (Figure 1).
+These included: 1) **Accessing** data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) **Harmonising** these diverse datasets into a federated resource, with common units, trait names, and data formats; and 3) **Distributing** versions of the data under suitable license, while also ensuring appropriate recognition for contributors.
+Successive versions AusTraits will iterate through steps 1 and 2 to incorporate new data and correct identified errors.
+We developed a workflow for harmonising and distributing data, drawing from emerging community standards and our collective experience building traits databases.
+Combined, the tools applied ensure current best-practices in the construction and distribution of data resources, by achieving the following outcomes:
+
+  - A reproducible and extensible compilation process from raw data to assembled dataset;
+  - Transparent mapping of original values (species names, trait names, units) into the harmonised dataset;
+  - High quality data for end users; 
+  - Easy addition of new data and correction of community-reported errors; 
+  - Satisfying the FAIR principles13, i.e. standards needed to make data resources Findable, Accessible, Interoperable, and Reusable; and
+  - Appropriate recognition of contributors.
+
+We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to accelerate our collective understanding the Australian flora.
+It is anticipated that open access to a comprehensive resource like AusTraits will generate significant new knowledge about the Australian flora across multiple scales of interest.
+AusTraits will likely reduce duplication of effort in the compilation of plant trait data, particularly for research students and government agencies seeking to create trait-based answers to their questions.
+
+### Methods
+
+#### Primary sources
+
+AusTraits was assembled from 146 distinct sources, including published literature, field campaigns, floras, and taxonomic treatments.
+We identified a list of candidate traits of interest, including growth form, plant height, leaf size, specific leaf area and seed mass, though several other traits were also collected opportunistically.
+A full list of traits and their sources appears in Table 1 (available online only).
+We then identified primary sources containing measurements for these traits and sought access to this data, including both data and metadata.
+Each original source was designated a `dataset_id`.
+Ordinarily the id is the surname of first author and year of publication, for the source’s primary citation (e.g.
+Falster_2005, for Dataset citation 40).
+Details from each source were saved with minimal modification into two plain text files:
+
+1.	`data.csv`: a table containing the actual trait data in comma-separated values format;
+2.	`metadata.yml`: contains relevant metadata for the study, as well as options for mapping trait names and units onto standard types, and any substitutions applied to the data in processing.
+These two files provide all the information needed to compile each study into a standardised format.
+
+#### Schema
+
+The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology14 in that, where available, trait data are connected to contextual information about the collection site (e.g.
+location co-ordinates, soil conditions, number of replicates) and information about the methods used to derive measurements (e.g.
+number of replicates, equipment used where available).
+The target format contains eight core objects, as described in Table 2.
+The contents of several of these objects are described further in Tables 3-6.
+This format was developed to include sufficient information about the data and methods used to collect trait measurements in each source.
+Where possible, trait measurement was recorded as being one of several possible `value_type` (Table 7), reflecting the type of measurement recorded.
+
+#### Harmonisation
+
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R15 and using custom code and drawing on functions from the packages `tidyverse`16, `stringr`17, `yaml`18, `rprojroot`, `remake`19, `gitr2`, `knitr`20 and `rmarkdown`21.
+In this workflow, we performed a series of operations, including reformatting data into standardising format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorial variables, encoding suitable metadata, and flagging data that did not pass quality checks.
+
+To enhance data quality, we produced detailed reports for each variable and source.
+These reports were reviewed both by AusTraits compilers as well as data contributors.
+Whilst all care is taken to curate datasets in AusTraits this is not intended to absolve individual researchers from ensuring the appropriate use of the data provided.
+
+#### Code Availability
+
+All code, raw and compiled data are hosted within a GitHub repository under the Trait Ecology and Evolution organisation (https://github.com/traitecoevo/austraits).
+Sequential versions of both source and compiled dataset are also being archived within Figshare (https://figshare.com/).
+The archived material includes all data sources and code for rebuilding the compiled dataset.
+
+#### Data Records
+
+A static version of the data contained in this descriptor has been released as version 0.0.9 of AusTraits, available on the project website (Data Citation 1) and Figshare (Data Citation 2).
+Details and references for all traits are summarised in Table 1 (available online only).
+The latest data can be downloaded directly from the project website https://github.com/traitecoevo/austraits.
+However, as validation (see Technical Validation, below) and data entry is ongoing, users are recommended to pull data from the static releases, to ensure results in their downstream analyses remain consistent as the database is updated.
+
+We envision AusTraits as a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
+For instance, we welcome the reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
+
+#### Technical Validation
+
+AusTraits is curated on a voluntary basis. Quality control of data and editorial procedures include:
+
+•	**Automated tests** which confirm that the species names provided can be mapped onto accepted names on The Plant List, that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
+Data that does not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use if desired.
+
+•	**Detailed review of each source based on bespoke reports**: A report is generated for each source, showing all data and metadata, and is checked by both an AusTraits curator and, where possible, the original contributor for accuracy.
+In these reports, observations for each trait are plotted against all other values for the trait in AusTraits.
+This allows AusTraits curators and contributors to quickly visually check the data to ensure it plots against other trait values as expected.
+Corrections suggested by contributors are combined back into AusTraits and made available with the next static release.
+
+•	**User feedback**: Data issues can be reported for any observation on the project website.
+
+#### Usage Notes
+
+Each data release is a compressed folder containing separate files for each of the major components of AusTraits (Table 2). Detailed examples for analysis are available on the project website.
+
+#### Acknowledgement
+
+RVG was supported by Australian Research Council DECRA Fellowship (DE170100208).
+
+#### Author contributions
+
+RVG, IJW conceived the original idea; RVG, EHW collated data from primary sources; DSF; EHW, DI, SCA, JL contributed to coding; AD error-checked trait observations; RVG, DSF, EHW wrote the first draft of the paper.
+[all other authors will be data contributors]
+
+#### Competing interests
+
+The authors have no conflicts of interest to declare.
+
 ```{r austraits}
-
 ## Assumes to items exist in global name space
@@
 
-tmp <- RefManageR::WriteBib(austraits$sources, "references.bib")
+tmp <- RefManageR::WriteBib(austraits$sources, "references.bib", biblatex = F)
+
+# check non ASCII file
+#tools::showNonASCIIfile("references.bib")
 
@@
 
-# Data sources:
+### Data sources:
 
-## Version 1 - via bib.print
+#### Version 1 - via bib.print
 
@@
 
-## Version 3 - via csl
+#### Version 3 - via csl
 
@@
 
-## References
+### References

### 2019-12-04 ms.Rmd
added manuscript references
added ms referenced

@@
 output:
+  pdf_document:
+    latex_engine: xelatex
   html_document:
     df_print: paged
-  pdf_document:
-    latex_engine: xelatex
   word_document:
@@
 - \usepackage[T1]{fontenc}
-bibliography: references.bib
+- \usepackage{float}
+bibliography: [ms_ref.bib]
 ---
+
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.pos = 'H')
+```
 
@@
 
-Species traits are essential metrics for comparing ecological strategies in plants arrayed across environmental space or evolutionary lineages1-4.
-Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function5-7.
-Traits thereby provide useful indicators of species behaviours in communities and ecosystems, regardless of their taxonomy8,9.
+Species traits are essential metrics for comparing ecological strategies in plants arrayed across environmental space or evolutionary lineages [@diaz2016global; @kunstler2016plant; @zanne2014three; @cornwell2014functional].
+Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function [@chapin1993evolution; @adler2014functional; @diaz1998plant].
+Traits thereby provide useful indicators of species behaviours in communities and ecosystems, regardless of their taxonomy [@violle2007let; @funk2017revisiting].
 Through targeted global initiatives the volume of available trait information for plants has grown rapidly in the last two decades.
@@
 One such region is Australia; a continent with a flora of c.
-24,000 native higher-plant species10.
+24,000 native higher-plant species [@council2010australian].
 While significant investment has been made in curating and digitising herbarium collections in Australia over the last two decades (e.g.
@@
 By providing a harmonised and curated dataset on 86 plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
-Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives (but see 11).
-Global initiatives, such as the TRY database12 and Open Traits Network, have been working towards sharing of data about plant traits though large numbers of Australian plant species remain unrepresented.
+Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives (but see @guerin2017opportunities).
+Global initiatives, such as the TRY database [@kattge2011try] and Open Traits Network, have been working towards sharing of data about plant traits though large numbers of Australian plant species remain unrepresented.
 
@@
   - Easy addition of new data and correction of community-reported errors; 
-  - Satisfying the FAIR principles13, i.e. standards needed to make data resources Findable, Accessible, Interoperable, and Reusable; and
+  - Satisfying the FAIR principles [@wilkinson2016fair], i.e. standards needed to make data resources Findable, Accessible, Interoperable, and Reusable; and
   - Appropriate recognition of contributors.
@@
 
-### Methods
+### Methodology
 
@@
 
-The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology14 in that, where available, trait data are connected to contextual information about the collection site (e.g.
+The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology [@madin2007ontology] in that, where available, trait data are connected to contextual information about the collection site (e.g.
 location co-ordinates, soil conditions, number of replicates) and information about the methods used to derive measurements (e.g.
@@
 
-To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R15 and using custom code and drawing on functions from the packages `tidyverse`16, `stringr`17, `yaml`18, `rprojroot`, `remake`19, `gitr2`, `knitr`20 and `rmarkdown`21.
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R [@team2018r] and using custom code and drawing on functions from the packages `tidyverse` [@wickham2016tidyverse], `stringr` [@wickham2017stringr], `yaml` [@stephens2014yaml], `rprojroot`, `remake` [@fitzjohn2016remake], `gitr2`, `knitr` [@xie1general] and `rmarkdown` [@allaire2015rmarkdown].
 In this workflow, we performed a series of operations, including reformatting data into standardising format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorial variables, encoding suitable metadata, and flagging data that did not pass quality checks.
@@
 
-•	**Automated tests** which confirm that the species names provided can be mapped onto accepted names on The Plant List, that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
+  - **Automated tests** which confirm that the species names provided can be mapped onto accepted names on The Plant List, that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
 Data that does not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use if desired.
 
-•	**Detailed review of each source based on bespoke reports**: A report is generated for each source, showing all data and metadata, and is checked by both an AusTraits curator and, where possible, the original contributor for accuracy.
+  - **Detailed review of each source based on bespoke reports**: A report is generated for each source, showing all data and metadata, and is checked by both an AusTraits curator and, where possible, the original contributor for accuracy.
 In these reports, observations for each trait are plotted against all other values for the trait in AusTraits.
@@
 
-•	**User feedback**: Data issues can be reported for any observation on the project website.
+  - **User feedback**: Data issues can be reported for any observation on the project website.
 
@@
 
-tmp <- RefManageR::WriteBib(austraits$sources, "references.bib", biblatex = F)
+tmp <- RefManageR::WriteBib(austraits$sources, "references.bib")
 
@@
 
-# Overview
+### Overview
 
@@
 
-# Data structures
+### Data structures
 
-# Figures
+\clearpage
+
+## Figures
 
@@
 
+\clearpage
+
 <!-- Cite: -->
@@
 
-# Tables
+## Tables
 

### 2020-02-05 ms.Rmd
Update MS details from details from data

@@
 - \usepackage{float}
-bibliography: [ms_ref.bib]
+bibliography: [data/references_ms.bib]
 ---
@@
 source("R/ms.R")
+
+austraits <- readRDS("data/austraits.rds")
 ```
@@
 ```{r authors, results='asis', echo=FALSE}
-authors <- read_csv("data/authors.csv", col_types = cols(.default = "c"))
+authors_build <- read_csv("data/lead_authors.csv", col_types = cols(.default = "c"))
 
-affils <- tibble(Affiliation=unique(authors$Affiliation), 
-                 label = letters[seq_len(length(Affiliation)) ])
-authors <- left_join(authors, affils, by = "Affiliation")
+to_exclude <- c("Barlow_1981","BRAIN_2007","CPRR_2002","Kew_2019_1","Kew_2019_2","Kew_2019_3",
+"Kew_2019_4","Kew_2019_5","Kew_2019_6","NHNSW_2016","NTH_2014","RBGK_2014","RBGSYD_0000","RBGSYD_2014","SAH_2014", "Wheeler_2002", "Hyland_2003", "SAH_2014")
+
+data_contributors <- austraits$contributors %>% select(-role) %>% 
+    # remove authors 
+    filter(!name %in% c("Peter Clarke", "Barbara Rice", "Westoby Lab")) %>% 
+    # intstituional collections
+    filter(!dataset_id %in% to_exclude) %>%
+    select(-dataset_id) %>% 
+    distinct() %>% 
+    mutate(surname = str_split(name, " ") %>% map_chr(last)) %>% 
+    arrange(surname) %>% rename(affiliation = institution)
+
+# Need to check how these institutions and individuals are to be recognised
+# austraits$contributors %>% select(-role) %>% 
+#     filter(dataset_id %in% to_exclude) %>% 
+#     filter(name!="Rachael Gallagher") %>% 
+#   write_csv("authors_check.csv")
+
+authors <- bind_rows(authors_build, 
+                     data_contributors %>% filter(! (name %in% authors_build$name)))
+
+affils <- tibble(affiliation=unique(authors$affiliation), 
+                 label = seq_len(length(affiliation)))
+authors <- left_join(authors, affils, by = "affiliation")
+
+authors$label[1:2] <- sprintf("%s,\\*", authors$label[1:2])
 
 # print list of authors
-sprintf("%s^%s^", authors$Author, authors$label) %>% paste0(collapse = ", ") %>% writeLines()
+
+sprintf("%s^%s^", authors$name, authors$label) %>% paste0(collapse = ", ") %>% sprintf("**Authors:** %s", .) %>% writeLines()
 
@@
 
+writeLines("^\\*^ contributed equally\n\n")
+
 # print list of affiliations
-sprintf("^%s^ %s", affils$label, affils$Affiliation) %>% paste0(collapse = "; ") %>% writeLines()
+sprintf("^%s^ %s", affils$label, affils$affiliation) %>% paste0(collapse = "; ")  %>% sprintf("**Affiliations:** %s", .) %>% writeLines()
+
 ```
@@
 We introduce the AusTraits database - a compilation of measurements of plant traits for species in the Australian Flora (hereafter AusTraits).
-AusTraits synthesises data on 86 traits across 20,321 species from field campaigns, published literature, taxonomic monographs, and individual species descriptions.
+AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$species_name %>% n_distinct()` species from field campaigns, published literature, taxonomic monographs, and individual species descriptions.
 Traits
@@
 harmonised individual-, species- and genus-level observations coupled to, where available, contextual information on site
-properties. This data descriptor provides information on Version 0.0.9 of AusTraits which contains data for 325,819 trait-by-species combinations across 20,321 species. We envision AusTraits as an on-
+properties.
+This data descriptor provides information on Version `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-species combinations.
+We envision AusTraits as an on-
 going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
@@
 Here we introduce the AusTraits database (hereafter simply AusTraits) - a compilation of plant traits for the Australian flora.
-Currently, AusTraits draws together 148 primary sources – including published scientific papers, floras, and taxonomic species descriptions– and contains 325,819 measurements spread across 86 different traits for 20,321 species (Table 1; available online only).
-By providing a harmonised and curated dataset on 86 plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
+Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources – including published scientific papers, floras, and taxonomic species descriptions– and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$species_name %>% n_distinct()` species (Table 1; available online only).
+By providing a harmonised and curated dataset on `r austraits$traits$trait_name %>% n_distinct()` plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
 Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives (but see @guerin2017opportunities).
-Global initiatives, such as the TRY database [@kattge2011try] and Open Traits Network, have been working towards sharing of data about plant traits though large numbers of Australian plant species remain unrepresented.
+Global initiatives, such as the TRY database [@Kattge-2020] and Open Traits Network [@Gallagher-2019], have been working towards sharing of data about plant traits though large numbers of Australian plant species remain unrepresented.
 
 To assemble AusTraits from diverse primary sources and make it available for reuse, we needed to overcome three main types of challenges (Figure 1).
-These included: 1) **Accessing** data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) **Harmonising** these diverse datasets into a federated resource, with common units, trait names, and data formats; and 3) **Distributing** versions of the data under suitable license, while also ensuring appropriate recognition for contributors.
-Successive versions AusTraits will iterate through steps 1 and 2 to incorporate new data and correct identified errors.
+These included: 1) **Accessing** data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) **Harmonising** these diverse datasets into a federated resource, with common units, trait names, and data formats; and 3) **Distributing** versions of the data under suitable license.
+Successive versions AusTraits will iterate through these to incorporate new data and correct identified errors.
 We developed a workflow for harmonising and distributing data, drawing from emerging community standards and our collective experience building traits databases.
@@
   - Satisfying the FAIR principles [@wilkinson2016fair], i.e. standards needed to make data resources Findable, Accessible, Interoperable, and Reusable; and
-  - Appropriate recognition of contributors.
+  - Recognition of contributors through citation of this resource and primary sources.
 
@@
 
-AusTraits was assembled from 146 distinct sources, including published literature, field campaigns, floras, and taxonomic treatments.
+AusTraits was assembled from `r austraits$sources %>% length()` distinct sources, including published literature, field campaigns, floras, and taxonomic treatments.
 We identified a list of candidate traits of interest, including growth form, plant height, leaf size, specific leaf area and seed mass, though several other traits were also collected opportunistically.
@@
 2.	`metadata.yml`: contains relevant metadata for the study, as well as options for mapping trait names and units onto standard types, and any substitutions applied to the data in processing.
+
 These two files provide all the information needed to compile each study into a standardised format.
@@
 
-To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R [@team2018r] and using custom code and drawing on functions from the packages `tidyverse` [@wickham2016tidyverse], `stringr` [@wickham2017stringr], `yaml` [@stephens2014yaml], `rprojroot`, `remake` [@fitzjohn2016remake], `gitr2`, `knitr` [@xie1general] and `rmarkdown` [@allaire2015rmarkdown].
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R [@R-2020] and using custom code and drawing on functions from the packages `tidyverse` [@wickham2016tidyverse], `stringr` [@wickham2017stringr], `yaml` [@stephens2014yaml], `remake` [@fitzjohn2016remake],  `knitr` [@xie1general] and `rmarkdown` [@allaire2015rmarkdown].
 In this workflow, we performed a series of operations, including reformatting data into standardising format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorial variables, encoding suitable metadata, and flagging data that did not pass quality checks.
@@
 
-All code, raw and compiled data are hosted within a GitHub repository under the Trait Ecology and Evolution organisation (https://github.com/traitecoevo/austraits).
-Sequential versions of both source and compiled dataset are also being archived within Figshare (https://figshare.com/).
+All code, raw and compiled data are hosted within a GitHub repository under the Trait Ecology and Evolution organisation (see https://github.com/traitecoevo/austraits.build, https://github.com/traitecoevo/austraits).
+Sequential versions of both source and compiled dataset are also being archived within Zenodo [@falster_2019].
 The archived material includes all data sources and code for rebuilding the compiled dataset.
@@
 
-A static version of the data contained in this descriptor has been released as version 0.0.9 of AusTraits, available on the project website (Data Citation 1) and Figshare (Data Citation 2).
+A static version of the data contained in this descriptor has been released as version`r austraits$build_info$version` of AusTraits, available on the project website (Data Citation 1) and Zenodo (Data Citation 2).
 Details and references for all traits are summarised in Table 1 (available online only).
@@
 
-RVG was supported by Australian Research Council DECRA Fellowship (DE170100208).
+DSF was supported by Australian Research Council Future Fellowship (FT160100113), RVG was supported by Australian Research Council DECRA Fellowship (DE170100208).
 
@@
 
-RVG, IJW conceived the original idea; RVG, EHW collated data from primary sources; DSF; EHW, DI, SCA, JL contributed to coding; AD error-checked trait observations; RVG, DSF, EHW wrote the first draft of the paper.
-[all other authors will be data contributors]
+RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, CB, JL, SA error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
+All other authors contributed the raw data and metadata undeprinning the resource and reviewed the harmonised data for errors.
 
@@
 
-tmp <- RefManageR::WriteBib(austraits$sources, "references.bib")
+tmp <- RefManageR::WriteBib(austraits$sources, "data/references_data.bib")
 

### 2020-02-05 scripts/fig_site_locations.R
- update fig2a with the latest austraits version - apply  ggpointdensity and viridis color scheme

@@ -1,54 +1,68 @@
 ## Austraits: site locations plot
-
 library(raster)
 library(tidyverse)
 library(ggplot2)
 library(RColorBrewer)
+library(viridis)
+library(ggpointdensity)
 source("R/plotting_functions.R")
 
 
 
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)
-au_map<-raster("data/australia.tif") %>% 
- # aggregate(fact=6) %>% 
-  as.data.frame(xy=T)
+
+# one strategy for retrieving data available on web
+filename <- "data/australia.tif"
+if (!file.exists(filename))
+  download.file("htpp:....", filename)
+
+# Read bioclim data using raster package
+au_map <- raster(filename) %>%
+  # aggregate(fact=6) %>%
+  as.data.frame(xy = T)
 
 
 #### 02 Plot AU basemap ####
-ggplot()+
-  geom_raster(data=au_map, aes(x = x, y = y, fill=factor(australia)))+
-  scale_fill_brewer(name="",type = "seq", palette = "#800000", guide=F)+xlab("")+ylab("")+ myTheme-> au_basemap
-#au_basemap
+ggplot() +
+  geom_raster(data = au_map, aes(
+    x = x,
+    y = y,
+    fill = factor(australia),
+    show.legend = FALSE)) +
+  myTheme -> au_basemap
+
+au_basemap
 
 #### 03 Overlay site locations  ####
-sites<-read_csv("data/austraits_sites.csv")
-austraits_site_locations<-au_basemap+
-  geom_point(data=sites, aes(y=`latitude (deg)`,
-                              x=`longitude (deg)`, size=n),
-              alpha=0.3, inherit.aes = F)+
-  scale_size(name   = "N Species",
-             breaks = sqrt(fivenum(sites$n)),
-             labels = fivenum(sites$n))
-
-# austraits_site_locations
-# 
-# # #### 04 Save plots ####
-# ggsave("austraits_sites_distribution.png",
-#        austraits_site_locations,  
-#        height=4, width=6, units="in")
+#sites <- read_csv("data/austraits_sites.csv")
+source("R/get_sites.R")
+## Filter sites
+sites %>% 
+  filter(`latitude (deg)` > (-45),
+         `latitude (deg)` < (-9.5),
+         `longitude (deg)` < (153),
+         `longitude (deg)` > (110)) %>% 
[…]

### 2020-02-05 scripts/fig_trait_composition.R
apply viridis color scheme

@@ -10,9 +10,12 @@ austraits<-readRDS("data/austraits.rds")
 trait_category_lookup<-read_csv("data/trait_categories.csv")
 
 austraits$traits %>% 
-  left_join(trait_category_lookup,by="trait_name")  %>% mutate(category = str_replace(category, "physiology", "physiological"))%>%
+  left_join(trait_category_lookup,by="trait_name")  %>%
+  mutate(category = str_replace(category, "physiology", "physiological"))%>%
   group_by(tissue, category) %>% 
-  summarise(n = n_distinct(species_name)) %>% rename(Category=category) %>% drop_na()-> summary_n_species
+  summarise(n = n_distinct(species_name)) %>%
+  rename(Category=category) %>%
+  drop_na()-> summary_n_species
 
 
 
@@ -28,7 +31,7 @@ ggplot(data=summary_n_species,
                 labels = trans_format("log10", math_format(10^.x)))+
   xlab("Major plant part")+ ylab("Species (n)")+ scale_x_discrete(limits = positions)+ 
   theme(legend.position = "right") +
-  scale_fill_manual(values=cbPalette)->austraits_composition
+  scale_fill_viridis_d()->austraits_composition
 
 # austraits_composition
 # ggsave("austraits_composition.png",

### 2020-02-05 scripts/fig_climate_space_austraits.R
update with the latest austraits version

@@ -3,7 +3,6 @@ library(raster)
 library(tidyverse)
 library(hexbin)
 library(ggplot2)
-library(viridis)
 source("R/plotting_functions.R")
 
 #### 01 Get climate data for Australia ####
@@ -28,33 +27,11 @@ au_bioclim_table<-au_bioclim %>%
   na.omit() %>%
   as_tibble %>% mutate(region="Australia")
 
-#### 03 Get the point coordinates of Austraits data ####t
-
-# Get the coordinates of the sites
-austraits<-readRDS("data/austraits.rds")
-sites <- austraits$sites %>% 
-  filter(site_property %in%  c("longitude (deg)","latitude (deg)")) %>% 
-  spread(site_property, value)
-sites
-
-
-# get species number for each study
-austraits$traits %>%
-  group_by(dataset_id, site_name) %>%
-  distinct(species_name) %>% 
-  count(dataset_id, sort = TRUE) %>% 
-  ungroup()-> species_number_lookup
-
-# combine species number vs sites
-species_number_lookup %>% na.omit() %>%  
-left_join(sites, 
-          by = c("dataset_id" = "dataset_id", 
-                 "site_name" = "site_name"))->combined_sites
-
-write_csv(combined_sites, "data/austraits_sites.csv")
+#### 03 Get the point coordinates of Austraits data ####
+source("R/get_sites.R")
 
 # Get climate values for each site
-au_sites_clim<-combined_sites %>% rowid_to_column("ID") %>% 
+au_sites_clim<- sites %>% rowid_to_column("ID") %>% 
   rename(latitude=`latitude (deg)`, longitude=`longitude (deg)`) %>% 
   combine_occurence_climate(au_bioclim)
 
@@ -96,20 +73,13 @@ austraits_climate_space<-whittaker_base_plot() +
   
     geom_point(data = au_sites_clim, 
                aes(x=Temp/10, y=Prec/10, color="AusTraits sites"), 
-               alpha=0.5,
-               size=log10(au_sites_clim$n),
-               inherit.aes = FALSE)+
-  scale_colour_manual(name="",values = c("lightgrey","black"))+
+               alpha=0.2,
+              # size=log10(au_sites_clim$n),
+               size=1,
+               stroke = 0,
+               inherit.aes = FALSE,
+               position = "jitter")+
+  scale_colour_manual(name="",values = c("red","black"))+
   theme_classic()
 
-#austraits_climate_space
-# # 
[…]

### 2020-02-05 R/get_sites.R
a new script for collating austraits sites

@@ -1,0 +1,33 @@
+austraits <- readRDS("data/austraits.rds")
+library(tidyverse)
+
+# collect sites
+austraits$sites %>% 
+  as_tibble() %>% 
+  filter(site_property %in% c("latitude (deg)", "longitude (deg)")) %>% 
+  pivot_wider( names_from =site_property, values_from=value) %>% 
+  mutate(`latitude (deg)` = as.numeric(`latitude (deg)`),
+         `longitude (deg)` = as.numeric(`longitude (deg)`)) -> sites
+
+# calculate species number
+austraits$traits %>%
+  as_tibble() %>% 
+  dplyr::select(dataset_id, species_name, site_name) %>% 
+  dplyr::group_by(dataset_id, site_name) %>% 
+  distinct(species_name) %>% 
+  count(dataset_id, sort = TRUE) %>% 
+  ungroup() -> sp_number
+
+sites<- sites%>%
+  left_join(sp_number, by=c("dataset_id", "site_name"))
+rm(sp_number)
+# 
+# library(leaflet)
+# library(htmltools)
+# sites2 %>%
+#   mutate(label=paste(dataset_id,site_name)) %>%
+# leaflet() %>%
+#    addTiles() %>%
+#   addMarkers(~`longitude (deg)`, ~`latitude (deg)`,
+#              popup = ~htmlEscape(label),
+#              clusterOptions = markerClusterOptions())

### 2020-02-06 build_figures.R
- tidy up code

@@ -1,4 +1,3 @@
-
 #### Austraits: Figure 2 ####
 
 library(patchwork)
@@ -18,7 +17,7 @@ combined_plot<-(austraits_site_locations+
   plot_layout(ncol = 1, heights = c(3, 1))
 
 
-pdf("figures/austraits_fig2.pdf",height=6, width=12)
+png("figures/austraits_fig2.png",height=6, width=12, units = "in", res = 300)
 combined_plot
 dev.off()
 

### 2020-02-06 scripts/fig_climate_space_austraits.R

@@ -3,83 +3,75 @@ library(raster)
 library(tidyverse)
 library(hexbin)
 library(ggplot2)
+library(RColorBrewer)
+library(plotbiomes)
+
 source("R/plotting_functions.R")
 
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)
-bioclim<-getData("worldclim",var="bio",res=10)
+bioclim <- getData("worldclim", var = "bio", res = 10)
 # Pick BIO1 (Mean Annual Temperature; T) and BIO12 (Annual Precipitation; P)
-bioclim<- bioclim[[c(1,12)]]
-names(bioclim) <- c("Temp","Prec")
+bioclim <- bioclim[[c(1, 12)]]
+names(bioclim) <- c("Temp", "Prec")
 
 #### 02 Get the climate data for Australia ####
 # Load Australia landmass binary map
-au_map<-raster("data/australia.tif")
+au_map <- raster("data/australia.tif")
 # Clip bioclim data with the au map
 ## crop and mask
 bioclim_au <- crop(bioclim, extent(au_map)) #%>% mask(.,au_map)
-new.bioclim <- projectRaster(bioclim_au, au_map) # harmonize the spatial extent and projection
-au_bioclim<-mask(new.bioclim, au_map)
+new.bioclim <-
+  projectRaster(bioclim_au, au_map) # harmonize the spatial extent and projection
+au_bioclim <- raster::mask(new.bioclim, au_map)
 
 # Transform raster data into a tibble
-au_bioclim_table<-au_bioclim %>% 
+au_bioclim_table <- au_bioclim %>%
   as.data.frame() %>%
   na.omit() %>%
-  as_tibble %>% mutate(region="Australia")
+  as_tibble %>%
+  mutate(region = as.factor("Australia"))
 
 #### 03 Get the point coordinates of Austraits data ####
 source("R/get_sites.R")
 
 # Get climate values for each site
-au_sites_clim<- sites %>% rowid_to_column("ID") %>% 
-  rename(latitude=`latitude (deg)`, longitude=`longitude (deg)`) %>% 
+au_sites_clim <- sites %>% rowid_to_column("ID") %>%
+  rename(latitude = `latitude (deg)`, longitude = `longitude (deg)`) %>%
   combine_occurence_climate(au_bioclim)
 
 
-# #### 03 Plot the climate data ####
-# au_sites_clim
-# au_bioclim_table
-
-
-# # Option 1: hexagonal background + dot points (size ~ species number)
[…]

### 2020-02-06 scripts/fig_site_locations.R

@@ -28,8 +28,8 @@ ggplot() +
   geom_raster(data = au_map, aes(
     x = x,
     y = y,
-    fill = factor(australia),
-    show.legend = FALSE)) +
+    fill = factor(australia)
+  )) +
   myTheme -> au_basemap
 
 au_basemap
@@ -38,31 +38,43 @@ au_basemap
 #sites <- read_csv("data/austraits_sites.csv")
 source("R/get_sites.R")
 ## Filter sites
-sites %>% 
-  filter(`latitude (deg)` > (-45),
-         `latitude (deg)` < (-9.5),
-         `longitude (deg)` < (153),
-         `longitude (deg)` > (110)) %>% 
-  filter(!site_name %in%c("site_at_-17.4167_degS_and_151.8833_degE",
-                          "site_at_-16.5833_degS_and_150.3167_degE",
-                          "site_at_-16.9333_degS_and_149.1833_degE",
-                          "site_at_-16.9833_degS_and_149.8833_degE")) -> sites2
+sites %>%
+  filter(
+    `latitude (deg)` > (-45),
+    `latitude (deg)` < (-9.5),
+    `longitude (deg)` < (153),
+    `longitude (deg)` > (110)
+  ) %>%
+  filter(
+    !site_name %in% c(
+      "site_at_-17.4167_degS_and_151.8833_degE",
+      "site_at_-16.5833_degS_and_150.3167_degE",
+      "site_at_-16.9333_degS_and_149.1833_degE",
+      "site_at_-16.9833_degS_and_149.8833_degE"
+    )
+  ) -> sites2
 
 
 
 #### 03 Overlay site locations  ####
 austraits_site_locations <- au_basemap +
-  geom_pointdensity(data = sites2,
-                    aes(y = `latitude (deg)`,
-                        x = `longitude (deg)`
-                    ),
-                    inherit.aes = FALSE,
-                    show.legend = TRUE,
-                    adjust=1) +
-  scale_color_viridis(option="plasma")+
-  theme(legend.justification = c(-0.1, 0), 
-        legend.position = c(0.05, 0.05), 
-        legend.direction  = "horizontal")+
-  scale_fill_grey(name="", start = 0.8,guide=FALSE, na.value = "white")
-
-
+  geom_pointdensity(
+    data = sites2,
+    aes(y = `latitude (deg)`,
+        x = `longitude (deg)`),
+    inherit.aes = FALSE,
[…]

### 2020-02-06 scripts/fig_trait_composition.R

@@ -31,10 +31,5 @@ ggplot(data=summary_n_species,
                 labels = trans_format("log10", math_format(10^.x)))+
   xlab("Major plant part")+ ylab("Species (n)")+ scale_x_discrete(limits = positions)+ 
   theme(legend.position = "right") +
-  scale_fill_viridis_d()->austraits_composition
-
-# austraits_composition
-# ggsave("austraits_composition.png",
-#        austraits_composition,  
-#        height=5, width=8, units="in")
+  scale_fill_viridis_d()-> austraits_composition
 

### 2020-02-06 scripts/fig_climate_space_austraits.R
- make ploygon borders smaller

@@ -42,7 +42,13 @@ au_sites_clim <- sites %>% rowid_to_column("ID") %>%
 
 
 #### 04 Plot the climate data ####
-whittaker_base_plot() +
+ggplot() +
+  geom_polygon(data = Whittaker_biomes,
+               aes(x    = temp_c,
+                   y    = precp_cm,
+                   fill = biome),
+               colour = "gray98", # colour of polygon border
+               size   = 0.1)  +
   # add the temperature - precipitation data points
   geom_point(
     data = au_bioclim_table,
@@ -71,6 +77,7 @@ whittaker_base_plot() +
     breaks = names(Ricklefs_colors),
     labels = names(Ricklefs_colors),
     values =  alpha(Ricklefs_colors, 0.5)
+    
   ) +
   theme_classic() +
   guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) -> austraits_climate_space

### 2020-02-06 scripts/fig_climate_space_austraits.R
adjust x and y axis labels

@@ -80,5 +80,8 @@ ggplot() +
     
   ) +
   theme_classic() +
-  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) -> austraits_climate_space
+  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
+  xlab(expression(Temperature (degree * C))) +
+  ylab(" Precipitation (cm)")+
+  theme(text = element_text(size = 12)) -> austraits_climate_space
 

### 2020-07-09 ms.Rmd
Update data source to v 2.0

@@
 
-austraits <- readRDS("data/austraits.rds")
+austraits <- readRDS("data/austraits_2.0.0.rds")
 ```
@@
 
-to_exclude <- c("Barlow_1981","BRAIN_2007","CPRR_2002","Kew_2019_1","Kew_2019_2","Kew_2019_3",
-"Kew_2019_4","Kew_2019_5","Kew_2019_6","NHNSW_2016","NTH_2014","RBGK_2014","RBGSYD_0000","RBGSYD_2014","SAH_2014", "Wheeler_2002", "Hyland_2003", "SAH_2014")
-
-data_contributors <- austraits$contributors %>% select(-role) %>% 
-    # remove authors 
-    filter(!name %in% c("Peter Clarke", "Barbara Rice", "Westoby Lab")) %>% 
-    # intstituional collections
-    filter(!dataset_id %in% to_exclude) %>%
-    select(-dataset_id) %>% 
+data_contributors <- 
+  austraits$contributors %>% 
+    filter(
+          # old studies
+          !dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981"),
+           !role %in% c("assistant"),
+           # deceased
+           !name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper")
+           ) %>% 
+    mutate(year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
+    arrange(year) %>%
+    # take most recent address for each author
+    group_by(name) %>% 
+    summarise(institution = last(institution),
+              dataset_ids = paste(dataset_id, collapse = ", ")) %>% 
+    ungroup() %>% 
     distinct() %>% 
     mutate(surname = str_split(name, " ") %>% map_chr(last)) %>% 
-    arrange(surname) %>% rename(affiliation = institution)
-
-# Need to check how these institutions and individuals are to be recognised
-# austraits$contributors %>% select(-role) %>% 
-#     filter(dataset_id %in% to_exclude) %>% 
-#     filter(name!="Rachael Gallagher") %>% 
-#   write_csv("authors_check.csv")
+    rename(affiliation = institution) %>%
+#    bind_rows(
+#      read_csv("data/data_pending.csv")) %>%
+    arrange(surname) %>% 
+    select(-surname)
 
 authors <- bind_rows(authors_build, 
-                     data_contributors %>% filter(! (name %in% authors_build$name)))
+                     data_contributors %>% filter(! (name %in% authors_build$name))
+                    )
+
+#  authors %>% select(dataset_ids, everything()) %>% mutate_all(~ replace_na(.x, "")) %>% write_csv("data/author_list.csv")
+
+# Update author info from google sheets
 
@@
 
-We introduce the AusTraits database - a compilation of measurements of plant traits for species in the Australian Flora (hereafter AusTraits).
-AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$species_name %>% n_distinct()` species from field campaigns, published literature, taxonomic monographs, and individual species descriptions.
+We introduce the AusTraits database - a compilation of measurements of plant traits for taxa in the Australian Flora (hereafter AusTraits).
+AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$taxon_name %>% n_distinct()` taxa from field campaigns, published literature, taxonomic monographs, and individual taxa descriptions.
 Traits
@@
 properties.
-This data descriptor provides information on Version `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-species combinations.
+This data descriptor provides information on Version `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxa combinations.
 We envision AusTraits as an on-
 going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
+
+## Todo
+
+Refs
+
+* Australian Plant Census, Centre of Australian National Biodiversity Research, Council of Heads of Australasian Herbaria, (14/05/2020) https://id.biodiversity.org.au/tree/51354547
+* Australian Plant Name Index (continuously updated), Centre of Australian National Biodiversity Research, www.biodiversity.org.au/nsl/services/apni (14/05/2020)
 
@@
 Here we introduce the AusTraits database (hereafter simply AusTraits) - a compilation of plant traits for the Australian flora.
-Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources – including published scientific papers, floras, and taxonomic species descriptions– and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$species_name %>% n_distinct()` species (Table 1; available online only).
+Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources – including published scientific papers, floras, and taxonomic descriptions– and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa (Table 1; available online only).
 By providing a harmonised and curated dataset on `r austraits$traits$trait_name %>% n_distinct()` plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
 Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives (but see @guerin2017opportunities).
-Global initiatives, such as the TRY database [@Kattge-2020] and Open Traits Network [@Gallagher-2019], have been working towards sharing of data about plant traits though large numbers of Australian plant species remain unrepresented.
+Global initiatives, such as the TRY database [@Kattge-2020] and Open Traits Network [@Gallagher-2019], have been working towards sharing of data about plant traits though large numbers of Australian plant taxa remain unrepresented.
 
@@
 These included: 1) **Accessing** data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) **Harmonising** these diverse datasets into a federated resource, with common units, trait names, and data formats; and 3) **Distributing** versions of the data under suitable license.
-Successive versions AusTraits will iterate through these to incorporate new data and correct identified errors.
+Successive versions of AusTraits will iterate through these to incorporate new data and correct identified errors.
 We developed a workflow for harmonising and distributing data, drawing from emerging community standards and our collective experience building traits databases.
@@
   - A reproducible and extensible compilation process from raw data to assembled dataset;
-  - Transparent mapping of original values (species names, trait names, units) into the harmonised dataset;
+  - Transparent mapping of original values (taxon names, trait names, units) into the harmonised dataset;
   - High quality data for end users; 
@@
 
-  - **Automated tests** which confirm that the species names provided can be mapped onto accepted names on The Plant List, that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
+  - **Automated tests** which confirm that the taxa names provided can be mapped onto accepted names on The Plant List, that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
 Data that does not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use if desired.
@@
 
+We achknowelde XXXXXXXXXXx institutions, who
+
 #### Author contributions
@@
 ## Assumes to items exist in global name space
-austraits <- readRDS("data/austraits.rds")
+austraits <- readRDS("data/austraits_2.0.0.rds")
 
@@
 
-AusTraits harmonises data on `r austraits$traits$trait_name %>% unique() %>% length()` traits from `r austraits$methods$dataset_id %>% unique() %>% length()` different sources, including field campaigns, published literature, taxonomic monographs, and individual species descriptions.
+AusTraits harmonises data on `r austraits$traits$trait_name %>% unique() %>% length()` traits from `r austraits$methods$dataset_id %>% unique() %>% length()` different sources, including field campaigns, published literature, taxonomic monographs, and individual taxa descriptions.
 
@@
 
-![**Figure 2.** (a.) Locations where the trait data were collected and the corresponding number of species.
+![**Figure 2.** (a.) Locations where the trait data were collected and the corresponding number of taxa.
 (b.) AusTraits’ sites (grey) within Australia’s precipitation-temperature space (black) superimposed upon Whittaker’s major biomes diagram.
 Data from WorldClim (worldclim.org).
-(c.) Overview of the number of species with which traits are measured within each major plant organ and trait category.](figures/austraits_fig2_blue.png)
+(c.) Overview of the number of taxa with which traits are measured within each major plant organ and trait category.](figures/austraits_fig2_blue.png)
 
-![**Figure 3.** The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled species.
+![**Figure 3.** The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
 The heatmap colour intensities denote the number of traits measured within each major plant organ.
-The most widespread family names (with more than ten species) are labelled on the edge of the tree.](figures/austraits_phlogenetic_coverage.png)
+The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_phlogenetic_coverage.png)
 
@@
       select(austraits$traits, -original_name), sites) %>%
-  left_join(by="species_name",
-      select(austraits$taxonomy, species_name, family)) %>% 
+  left_join(by="taxon_name",
+      select(austraits$taxa, taxon_name, family)) %>% 
   left_join(sources_all, by=c("dataset_id"))
@@
             Studies = n_distinct(dataset_id),
-            Species = n_distinct(species_name),
+            Taxon = n_distinct(taxon_name),
             Families = n_distinct(family),
@@
 
-```{r, results="asis"}
+```{r, results="asis", eval=FALSE}
 print_defintions_element <- function(elements) {

### 2020-07-10 R/get_sites.R
splitted fig 3c into: - georeferenced - non-georeferenced

@@ -12,18 +12,27 @@ austraits$sites %>%
 # calculate species number
 austraits$traits %>%
   as_tibble() %>% 
-  dplyr::select(dataset_id, species_name, site_name) %>% 
+  dplyr::select(dataset_id, taxon_name, site_name) %>% 
   dplyr::group_by(dataset_id, site_name) %>% 
-  distinct(species_name) %>% 
+  distinct(taxon_name) %>% 
   count(dataset_id, sort = TRUE) %>% 
   ungroup() -> sp_number
 
 # get trait categories
-trait_category_lookup <- read_csv("data/trait_categories.csv")
+trait_category_lookup <- read_csv("data/traits_filled_in.csv") %>% 
+  dplyr::select(trait_name, tissue, category)
+
+# get trait list
+trait_list<- austraits$traits %>%
+  as_tibble() %>% 
+  dplyr::select(dataset_id, taxon_name, site_name, trait_name) 
+
+trait_composition_list <- trait_list %>% 
+  full_join(trait_category_lookup, by = "trait_name") %>% 
+  full_join(sites, by = c("dataset_id", "site_name"))
 
 # uncategorised/ unclassified traits
 unclassified_traits <- austraits$traits %>%
-  as_tibble()  %>% 
   dplyr::select(dataset_id, site_name, trait_name) %>% 
   left_join(trait_category_lookup, by = "trait_name") %>% 
   dplyr::filter(is.na(category)) %>% 

### 2020-07-10 scripts/fig_trait_composition.R

@@ -4,32 +4,54 @@ library(tidyverse)
 library(ggplot2)
 library(scales)
 library(RColorBrewer)
-source("R/plotting_functions.R")
-
-austraits<-readRDS("data/austraits.rds")
-trait_category_lookup<-read_csv("data/trait_categories.csv")
-
-austraits$traits %>% 
-  left_join(trait_category_lookup,by="trait_name")  %>%
-  mutate(category = str_replace(category, "physiology", "physiological"))%>%
-  group_by(tissue, category) %>% 
-  summarise(n = n_distinct(species_name)) %>%
-  rename(Category=category) %>%
-  drop_na()-> summary_n_species
 
 
+## Change the order of the
+positions <-
+  c("leaf", "stem", "bark", "root", "reproductive", "whole_plant")
+cbPalette <- brewer.pal(5, "Blues")[2:5]
 
-## Change the order of the 
-positions <- c("leaf", "stem", "root", "reproductive", "whole_plant")
-cbPalette <- brewer.pal(5,"Blues")[2:5]
 
-ggplot(data=summary_n_species, 
-       aes(x=tissue, y=n, fill=Category))+
-  geom_bar( stat='identity', position =position_dodge(preserve = "single")) +
-  theme_classic()+ 
-  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
-                labels = trans_format("log10", math_format(10^.x)))+
-  xlab("Major plant part")+ ylab("Species (n)")+ scale_x_discrete(limits = positions)+ 
+# Georeferenced trait data
+trait_composition_list %>% 
+  filter(!is.na(`latitude (deg)`), !is.na(`longitude (deg)`)) %>% 
+  group_by(tissue, category) %>%
+  summarise(n = n_distinct(taxon_name)) %>%
+  rename(Category = category) %>%
+   drop_na() %>% 
+  ggplot(
+    aes(x = tissue, y = n, fill = Category)) +
+  geom_bar(stat = 'identity', position = position_dodge(preserve = "single")) +
+  theme_minimal() +
+  scale_y_log10(
+    breaks = trans_breaks("log10", function(x)
+      10 ^ x),
+    labels = trans_format("log10", math_format(10 ^ .x))
+  ) +
+  xlab("Major plant part") + ylab("Species (n)") + 
+  scale_x_discrete(limits = positions) +
   theme(legend.position = "right") +
-  scale_fill_viridis_d()-> austraits_composition
[…]

### 2020-07-15 scripts/fig_trait_composition.R
split austraits composition barplot into two parts:  - georeferenced sites  - non-georeferenced sites

@@ -32,7 +32,8 @@ trait_composition_list %>%
   scale_x_discrete(limits = positions) +
   theme(legend.position = "right") +
   scale_fill_viridis_d() + 
-  annotation_logticks(sides="l") -> austraits_composition_georef
+  annotation_logticks(sides="l") +
+  ggtitle("Georeferenced") -> austraits_composition_georef
 
 # Non-georeferenced trait data
 trait_composition_list %>% 
@@ -54,4 +55,5 @@ trait_composition_list %>%
   scale_x_discrete(limits = positions) +
   theme(legend.position = "right") +
   scale_fill_viridis_d(option = "plasma") + 
-  annotation_logticks(sides="l") -> austraits_composition_non_georef
+  annotation_logticks(sides="l")+
+  ggtitle("Non-Georeferenced") -> austraits_composition_non_georef

### 2020-07-15 scripts/fig_climate_space_austraits.R
moved biome legend into the plotting area

@@ -61,9 +61,9 @@ ggplot() +
   
   # add the AusTraits Sites
   geom_point(
-    data = au_sites_clim,
+    data = au_sites_clim ,
     aes(x = Temp / 10, y = Prec / 10, color = "AusTraits sites"),
-    alpha = 0.2,
+    alpha = 0.3,
     stroke = 0,
     size = 1,
     inherit.aes = FALSE,
@@ -71,7 +71,7 @@ ggplot() +
   ) +
   
   # set color for  the temperature - precipitation data points and the the AusTraits Sites
-  scale_colour_manual(name = "", values = c("#d3d3d3", "#233D4D")) +
+  scale_colour_manual(name = "Australian climate space", values = c("#FF7F50", "#233D4D")) +
   scale_fill_manual(
     name   = "Whittaker biomes",
     breaks = names(Ricklefs_colors),
@@ -83,5 +83,13 @@ ggplot() +
   guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
   xlab(expression(Temperature (degree * C))) +
   ylab(" Precipitation (cm)")+
-  theme(text = element_text(size = 12)) -> austraits_climate_space
+  theme(text = element_text(size = 12))  +
+  theme(
+    legend.justification = c(-0.1, 0),
+    legend.position = c(0.005, 0.25),
+    legend.text=element_text(size=8),
+    legend.title=element_text(size=10)
+    #legend.key.size = unit(1, "cm")
+  ) -> austraits_climate_space
 
+#austraits_climate_space

### 2020-07-15 scripts/fig_site_locations.R
figure 2A Austraits site locations by organs/tissue

@@ -57,6 +57,23 @@ sites %>%
 
 
 #### 03 Overlay site locations  ####
-austraits_site_locations <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = NA )
-austraits_site_locations_by_tissue <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "tissue" )
-austraits_site_locations_by_category <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "category" )
+# austraits_site_locations <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = NA )
+# austraits_site_locations_by_tissue <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "tissue" )
+# austraits_site_locations_by_category <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "category" )
+
+austraits_site_locations_by_tissue_fig_2 <- sites2 %>% 
+  drop_na() %>%
+  produce_site_map( "latitude (deg)", "longitude (deg)", feature = "tissue" )+
+  theme(
+    #legend.justification = c(-.05, -.05),
+    legend.position ="bottom",
+   # legend.direction  = "horizontal",
+    strip.background = element_blank(),
+    strip.text.x = element_text(
+      size = 12),
+     legend.key.height = unit(0.5, "cm"),
+    legend.key.width = unit(2, "cm"),
+     legend.box = "horizontal" #,
+    # legend.spacing  = unit(0.1, "cm")
+    # legend.text = element_text(size=2)
+  )

### 2020-07-15 build_figures.R
export fig 2 and fig 3

@@ -7,21 +7,24 @@ source("scripts/fig_climate_space_austraits.R") # B
 source("scripts/fig_trait_composition.R") # C
 
 
-combined_plot<-(austraits_site_locations+
-                  austraits_climate_space)-
-                  {
-                    austraits_composition +
-                      plot_spacer()+
-                      plot_layout(widths = c(14, 1)) 
-                  }+
-  plot_layout(ncol = 1, heights = c(3, 1))
+# combined_plot <- (austraits_site_locations+
+#                   austraits_climate_space)-
+#                   {
+#                     austraits_composition +
+#                       plot_spacer()+
+#                       plot_layout(widths = c(14, 1)) 
+#                   }+
+#   plot_layout(ncol = 1, heights = c(3, 1))
 
 
-png("figures/austraits_fig2.png",height=6, width=12, units = "in", res = 300)
-combined_plot
+
+pdf("figures/austraits_fig2.pdf",height=12, width=8)
+austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A') 
 dev.off()
 
-
+pdf("figures/austraits_fig3.pdf",height=6, width=8)
+austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'A') 
+dev.off()
 
 #### Austraits: Phylogenetic coverage of traits ####
 source("scripts/fig_generate_phylogenies.R") # A
@@ -43,7 +46,7 @@ base_tree<-ggtree(base_tree, # turn base_tree back to ggtree format
   guides(color=FALSE)
 #(values = cbp2)
 # rotate the base_tree
-base_tree<-rotate_tree(base_tree, 300)
+base_tree <- rotate_tree(base_tree, 300)
 
 # attach clade label
 for(j in 1:length(nodes_list)){
@@ -75,6 +78,6 @@ austraits_tree <-  gheatmap(base_tree, austraits_summary,
                 font.size=3,
                 hjust=1)
 
-pdf("figures/austraits_phlogenetic_coverage2.pdf", height=10, width=10)
+pdf("figures/fig_4_austraits_phlogenetic_coverage.pdf", height=10, width=10)
 austraits_tree
 dev.off()

### 2020-07-15 R/plotting_functions.R
changed point size and transparency for fig 2b

@@ -120,7 +120,9 @@ produce_site_map <- function(df, lat, lon, feature=NA){
           x = !!as.name(lon)),
       inherit.aes = FALSE,
       show.legend = TRUE,
-      adjust = 1
+      adjust = 1,
+      size = 0.5,
+      alpha=0.8
     ) +
     scale_color_viridis(option = "plasma") +
     theme(

### 2020-07-15 build_figures.R
fig_4 updated taxonomy w/ national color of AU color pallete

@@ -31,22 +31,20 @@ source("scripts/fig_generate_phylogenies.R") # A
 source("scripts/fig_get_family_node.R") # B
 
 #### 1 Plot a base phylogenetic tree ####
-base_tree <- ggtree(tree$scenario.2$run.1) #base_tree
+base_tree <- ggtree(tree$scenario.2$run.1, layout = "circular",  size=0.01,)+ 
+  guides(color=FALSE) #base_tree
+
+# rotate the base_tree
+base_tree <- rotate_tree(base_tree, 300)
 
 # make a list of nodes
 node_lookup_outcome %>% 
   select(node) %>% 
-  unlist()->nodes_list
+  unlist() -> nodes_list
 
 
-base_tree <- groupClade(base_tree, .node=nodes_list) # group by nodes
-base_tree<-ggtree(base_tree, # turn base_tree back to ggtree format
-                  size=0.01, # change line thickness
-                  layout = "circular") + 
-  guides(color=FALSE)
-#(values = cbp2)
-# rotate the base_tree
-base_tree <- rotate_tree(base_tree, 300)
+base_tree <- ggtree::groupClade(base_tree, .node=nodes_list) # group by nodes
+
 
 # attach clade label
 for(j in 1:length(nodes_list)){
@@ -64,20 +62,32 @@ for(j in 1:length(nodes_list)){
   }
 }
 
-
+tree_breaks = c(1, 3, 10, 30, 100)
 austraits_tree <-  gheatmap(base_tree, austraits_summary, 
                 offset=4, 
-                width=0.2, 
-                low="#77dd77", 
-                high="red", 
+                width=0.2,
                 color = NULL,
                 colnames_angle=30,
                 colnames_position = "top", 
                 colnames_offset_x =0,
                 colnames_offset_y =50,
                 font.size=3,
-                hjust=1)
+                hjust=1) +
+  scale_fill_gradientn( trans="log10", 
+                       # low="#014127",
+                       # high="#ff4500",
+                       colours=c("#ffcd00", "#94c338", "#2bac66", "#00843D"), # National colours of Australia
+                       labels = tree_breaks,
+                       breaks = tree_breaks, 
+                       na.value = 'white',
+                       name="Trait count") +
+  theme(legend.position="bottom", legend.justification = "right")+
+  guides(fill = guide_colourbar(title.position="top",
+                                label.position = "bottom", 
+                                barwidth = 10, 
[…]

### 2020-07-15 scripts/fig_generate_phylogenies.R
replace species_name to taxon_name

@@ -23,23 +23,21 @@ library(tidytree)
 ## add austraits and read the lookup table
 austraits <- readRDS("data/austraits.rds")
 #tree<-readRDS("data/AusTraits_phylogenies.rds") 
-trait_category_lookup<-read_csv("data/trait_categories.csv")
+trait_category_lookup<-read_csv("data/traits_filled_in.csv") %>% 
+  dplyr::select(trait_name, tissue, category)
 
 
 ##### 2 Build a phylogenetic tree ####
 #build a tree with n number of species
-n_species_tree<-2000
+n_species_tree <- 2000
 
 ## generate list of species from austraits (or sample 2000 species)
 set.seed(10)
-austraits$taxonomy %>% 
-  filter(species_name!="Stenocarpus top end'") %>%
-  #sample_n(40) %>% 
-  select(species_name, genus, family) %>% 
+austraits$taxa %>% 
+  select(taxon_name, genus, family) %>% 
   filter(!is.na(family))%>% 
-  mutate(species_name = gsub(" ", "_", species_name)) %>% 
-  sample_n(n_species_tree)->austraits_sp_list_sampled
-#20,641 species in austraits$taxonomy vs 20,593 species in austraits$traits
+  mutate(taxon_name = gsub(" ", "_", taxon_name)) %>% 
+  sample_n(n_species_tree) -> austraits_sp_list_sampled
 
 
 # Build a tree using V.PhyloMaker (probably in a newick format)
@@ -54,23 +52,23 @@ tree <- phylo.maker(sp.list = austraits_sp_list_sampled,
 austraits$traits %>% 
   left_join(trait_category_lookup,by="trait_name")  %>% 
   mutate(category = str_replace(category, "physiology", "physiological"))%>%
-  group_by(species_name,tissue) %>%
-  distinct(species_name, trait_name) %>% 
+  group_by(taxon_name,tissue) %>%
+  distinct(taxon_name, trait_name) %>% 
   summarise(n = n()) %>%
   drop_na() %>%
   arrange(-n) %>% 
   ungroup() %>%
-  mutate(species_name = gsub(" ", "_", species_name)) %>% 
-  filter(species_name %in% tree$species.list$species) %>% 
+  mutate(taxon_name = gsub(" ", "_", taxon_name)) %>% 
+  filter(taxon_name %in% tree$species.list$species) %>% 
   spread(key=tissue,value=n) %>%
-  select(species_name, root, whole_plant, stem, reproductive, leaf) %>% 
[…]

### 2020-07-15 scripts/fig_get_family_node.R

@@ -7,12 +7,12 @@ source("R/plotting_functions.R")
 rm(myTheme)
 
 #### 1 Group species by family ####
-austraits_sp_list_sampled %>%
-  filter(species_name %in% tree$scenario.2$run.1$tip.label) %>% 
+family_list <- austraits_sp_list_sampled %>%
+  filter(taxon_name %in% tree$scenario.2$run.1$tip.label) %>% 
   group_by(family) %>%
-  summarise(species_name = paste(
-    paste0("\"",species_name, "\""),collapse=", ")) %>% 
-  mutate(species_name=paste0('c(',species_name, ')'))->family_list
+  summarise(taxon_name = paste(
+    paste0("\"",taxon_name, "\""),collapse=", ")) %>% 
+  mutate(taxon_name=paste0('c(',taxon_name, ')')) 
 
 #### 2 Calculate the most recent common ancestor to the node for each family ####
 
@@ -25,7 +25,7 @@ for(i in 1:nrow(family_list)){
 
 # convert list to data frame
 node_lookup %>%
-  dplyr::bind_rows() ->node_lookup_outcome 
+  dplyr::bind_rows() -> node_lookup_outcome 
 
 rm(node_lookup)
 

### 2020-07-15 build_figures.R
minor colorbar scale adjustment

@@ -62,7 +62,7 @@ for(j in 1:length(nodes_list)){
   }
 }
 
-tree_breaks = c(1, 3, 10, 30, 100)
+tree_breaks = c(1, 3, 10, 30, 80)
 austraits_tree <-  gheatmap(base_tree, austraits_summary, 
                 offset=4, 
                 width=0.2,

### 2020-07-15 scripts/fig_link_tree_and_trait_count_matrix.R
separate fig 4 as a single R script

@@ -1,0 +1,70 @@
+#### 1 Plot a base phylogenetic tree ####
+base_tree <-
+  ggtree(tree$scenario.2$run.1, layout = "circular",  size = 0.01, ) +
+  guides(color = FALSE) #base_tree
+
+# rotate the base_tree
+base_tree <- rotate_tree(base_tree, 300)
+
+# make a list of nodes
+node_lookup_outcome %>%
+  select(node) %>%
+  unlist() -> nodes_list
+
+
+base_tree <-
+  ggtree::groupClade(base_tree, .node = nodes_list) # group by nodes
+
+
+# attach clade label
+for (j in 1:length(nodes_list)) {
+  #Then add each clade label
+  if (is.na(node_lookup_outcome$node[j])) {
+    print("skip labelling clade with one sp")
+  } else{
+    base_tree <-
+      base_tree + geom_cladelabel(
+        node = node_lookup_outcome$node[j],
+        label = node_lookup_outcome$family[j],
+        offset = 100,
+        offset.text = 2,
+        barsize = 1,
+        fontsize = 3,
+        angle = "auto"
+      )
+  }
+}
+
+#### 2 Plot tree with multiple associated trait count matrix####
+tree_breaks = c(1, 3, 10, 30, 80)
+austraits_tree <-  gheatmap(
+  base_tree,
+  austraits_summary,
+  offset = 4,
+  width = 0.2,
+  color = NULL,
+  colnames_angle = 30,
+  colnames_position = "top",
+  colnames_offset_x = 0,
+  colnames_offset_y = 50,
+  font.size = 3,
+  hjust = 1
+) +
+  scale_fill_gradientn(
+    trans = "log10",
+    # low="#014127",
+    # high="#ff4500",
+    colours = c("#ffcd00", "#94c338", "#2bac66", "#00843D"),
+    # National colours of Australia
+    labels = tree_breaks,
+    breaks = tree_breaks,
+    na.value = 'white',
+    name = "Trait count"
+  ) +
+  theme(legend.position = "bottom", legend.justification = "right") +
+  guides(fill = guide_colourbar(
+    title.position = "top",
+    label.position = "bottom",
+    barwidth = 10,
+    barheight = 1
+  ))

### 2020-07-15 R/get_sites.R
reformat code to improve readibility

@@ -2,65 +2,57 @@ austraits <- readRDS("data/austraits.rds")
 library(tidyverse)
 
 # collect sites
-austraits$sites %>% 
-  as_tibble() %>% 
-  filter(site_property %in% c("latitude (deg)", "longitude (deg)")) %>% 
-  pivot_wider( names_from =site_property, values_from=value) %>% 
-  mutate(`latitude (deg)` = as.numeric(`latitude (deg)`),
-         `longitude (deg)` = as.numeric(`longitude (deg)`)) -> sites
+austraits$sites %>%
+  as_tibble() %>%
+  filter(site_property %in% c("latitude (deg)", "longitude (deg)")) %>%
+  pivot_wider(names_from = site_property, values_from = value) %>%
+  mutate(
+    `latitude (deg)` = as.numeric(`latitude (deg)`),
+    `longitude (deg)` = as.numeric(`longitude (deg)`)
+  ) -> sites
 
 # calculate species number
 austraits$traits %>%
-  as_tibble() %>% 
-  dplyr::select(dataset_id, taxon_name, site_name) %>% 
-  dplyr::group_by(dataset_id, site_name) %>% 
-  distinct(taxon_name) %>% 
-  count(dataset_id, sort = TRUE) %>% 
+  as_tibble() %>%
+  dplyr::select(dataset_id, taxon_name, site_name) %>%
+  dplyr::group_by(dataset_id, site_name) %>%
+  distinct(taxon_name) %>%
+  count(dataset_id, sort = TRUE) %>%
   ungroup() -> sp_number
 
 # get trait categories
-trait_category_lookup <- read_csv("data/traits_filled_in.csv") %>% 
+trait_category_lookup <- read_csv("data/traits_filled_in.csv") %>%
   dplyr::select(trait_name, tissue, category)
 
 # get trait list
-trait_list<- austraits$traits %>%
-  as_tibble() %>% 
-  dplyr::select(dataset_id, taxon_name, site_name, trait_name) 
+trait_list <- austraits$traits %>%
+  as_tibble() %>%
+  dplyr::select(dataset_id, taxon_name, site_name, trait_name)
 
-trait_composition_list <- trait_list %>% 
-  full_join(trait_category_lookup, by = "trait_name") %>% 
+trait_composition_list <- trait_list %>%
+  full_join(trait_category_lookup, by = "trait_name") %>%
   full_join(sites, by = c("dataset_id", "site_name"))
 
 # uncategorised/ unclassified traits
[…]

### 2020-07-15 build_figures.R

@@ -1,93 +1,36 @@
-#### Austraits: Figure 2 ####
+#### Figures for the Austraits' manuscript ####
 
 library(patchwork)
 
+#### Figure 2: geographical and biome map####
 source("scripts/fig_site_locations.R") # A
 source("scripts/fig_climate_space_austraits.R") # B
+
+
+pdf("figures/austraits_fig_2_biome_map.pdf",
+    height = 12,
+    width = 8)
+austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A')
+dev.off()
+
+#### Figure 3: trait composition count ####
 source("scripts/fig_trait_composition.R") # C
-
-
-# combined_plot <- (austraits_site_locations+
-#                   austraits_climate_space)-
-#                   {
-#                     austraits_composition +
-#                       plot_spacer()+
-#                       plot_layout(widths = c(14, 1)) 
-#                   }+
-#   plot_layout(ncol = 1, heights = c(3, 1))
-
-
-
-pdf("figures/austraits_fig2.pdf",height=12, width=8)
-austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A') 
+pdf("figures/austraits_fig_3_composition.pdf",
+    height = 6,
+    width = 8)
+austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'A')
 dev.off()
 
-pdf("figures/austraits_fig3.pdf",height=6, width=8)
-austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'A') 
-dev.off()
-
-#### Austraits: Phylogenetic coverage of traits ####
+#### Figure 4: Phylogenetic representation of AusTraits ####
 source("scripts/fig_generate_phylogenies.R") # A
 source("scripts/fig_get_family_node.R") # B
-
-#### 1 Plot a base phylogenetic tree ####
-base_tree <- ggtree(tree$scenario.2$run.1, layout = "circular",  size=0.01,)+ 
-  guides(color=FALSE) #base_tree
-
-# rotate the base_tree
-base_tree <- rotate_tree(base_tree, 300)
-
-# make a list of nodes
-node_lookup_outcome %>% 
-  select(node) %>% 
-  unlist() -> nodes_list
+source("scripts/fig_link_tree_and_trait_count_matrix.R") # B
 
 
[…]

### 2020-07-15 scripts/fig_climate_space_austraits.R

@@ -32,7 +32,7 @@ au_bioclim_table <- au_bioclim %>%
   as_tibble %>%
   mutate(region = as.factor("Australia"))
 
-#### 03 Get the point coordinates of Austraits data ####
+#### 03 Extract geographic coordinates of AusTraits' sites#
 source("R/get_sites.R")
 
 # Get climate values for each site
@@ -41,14 +41,17 @@ au_sites_clim <- sites %>% rowid_to_column("ID") %>%
   combine_occurence_climate(au_bioclim)
 
 
-#### 04 Plot the climate data ####
-ggplot() +
-  geom_polygon(data = Whittaker_biomes,
-               aes(x    = temp_c,
-                   y    = precp_cm,
-                   fill = biome),
-               colour = "gray98", # colour of polygon border
-               size   = 0.1)  +
+#### 04 Plot austraits sites over whittaker biome plot ####
+austraits_climate_space <- ggplot() +
+  geom_polygon(
+    data = Whittaker_biomes,
+    aes(x    = temp_c,
+        y    = precp_cm,
+        fill = biome),
+    colour = "gray98",
+    # colour of polygon border
+    size   = 0.1
+  )  +
   # add the temperature - precipitation data points
   geom_point(
     data = au_bioclim_table,
@@ -82,14 +85,14 @@ ggplot() +
   theme_classic() +
   guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
   xlab(expression(Temperature (degree * C))) +
-  ylab(" Precipitation (cm)")+
+  ylab(" Precipitation (cm)") +
   theme(text = element_text(size = 12))  +
   theme(
     legend.justification = c(-0.1, 0),
     legend.position = c(0.005, 0.25),
-    legend.text=element_text(size=8),
-    legend.title=element_text(size=10)
+    legend.text = element_text(size = 8),
+    legend.title = element_text(size = 10)
     #legend.key.size = unit(1, "cm")
-  ) -> austraits_climate_space
+  ) 
 
 #austraits_climate_space

### 2020-07-15 scripts/fig_site_locations.R

@@ -38,7 +38,7 @@ ggplot() +
 #sites <- read_csv("data/austraits_sites.csv")
 source("R/get_sites.R")
 ## Filter sites
-sites %>%
+sites2 <- sites %>%
   filter(
     `latitude (deg)` > (-45),
     `latitude (deg)` < (-9.5),
@@ -52,7 +52,7 @@ sites %>%
       "site_at_-16.9333_degS_and_149.1833_degE",
       "site_at_-16.9833_degS_and_149.8833_degE"
     )
-  ) -> sites2
+  ) 
 
 
 
@@ -61,19 +61,18 @@ sites %>%
 # austraits_site_locations_by_tissue <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "tissue" )
 # austraits_site_locations_by_category <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "category" )
 
-austraits_site_locations_by_tissue_fig_2 <- sites2 %>% 
+austraits_site_locations_by_tissue_fig_2 <- sites2 %>%
   drop_na() %>%
-  produce_site_map( "latitude (deg)", "longitude (deg)", feature = "tissue" )+
+  produce_site_map("latitude (deg)", "longitude (deg)", feature = "tissue") +
   theme(
     #legend.justification = c(-.05, -.05),
-    legend.position ="bottom",
-   # legend.direction  = "horizontal",
+    legend.position = "bottom",
+    # legend.direction  = "horizontal",
     strip.background = element_blank(),
-    strip.text.x = element_text(
-      size = 12),
-     legend.key.height = unit(0.5, "cm"),
+    strip.text.x = element_text(size = 12),
+    legend.key.height = unit(0.5, "cm"),
     legend.key.width = unit(2, "cm"),
-     legend.box = "horizontal" #,
+    legend.box = "horizontal" #,
     # legend.spacing  = unit(0.1, "cm")
     # legend.text = element_text(size=2)
   )

### 2020-07-15 scripts/fig_trait_composition.R

@@ -6,21 +6,18 @@ library(scales)
 library(RColorBrewer)
 
 
-## Change the order of the
+## Rearrange order of tissue
 positions <-
   c("leaf", "stem", "bark", "root", "reproductive", "whole_plant")
-cbPalette <- brewer.pal(5, "Blues")[2:5]
-
 
 # Georeferenced trait data
-trait_composition_list %>% 
-  filter(!is.na(`latitude (deg)`), !is.na(`longitude (deg)`)) %>% 
+austraits_composition_georef <- trait_composition_list %>%
+  filter(!is.na(`latitude (deg)`),!is.na(`longitude (deg)`)) %>%
   group_by(tissue, category) %>%
   summarise(n = n_distinct(taxon_name)) %>%
   rename(Category = category) %>%
-   drop_na() %>% 
-  ggplot(
-    aes(x = tissue, y = n, fill = Category)) +
+  drop_na() %>%
+  ggplot(aes(x = tissue, y = n, fill = Category)) +
   geom_bar(stat = 'identity', position = position_dodge(preserve = "single")) +
   theme_minimal() +
   scale_y_log10(
@@ -28,22 +25,21 @@ trait_composition_list %>%
       10 ^ x),
     labels = trans_format("log10", math_format(10 ^ .x))
   ) +
-  xlab("Major plant part") + ylab("Species (n)") + 
+  xlab("Major plant part") + ylab("Species (n)") +
   scale_x_discrete(limits = positions) +
   theme(legend.position = "right") +
-  scale_fill_viridis_d() + 
-  annotation_logticks(sides="l") +
-  ggtitle("Georeferenced") -> austraits_composition_georef
+  scale_fill_viridis_d() +
+  annotation_logticks(sides = "l") +
+  ggtitle("Georeferenced")
 
 # Non-georeferenced trait data
-trait_composition_list %>% 
-  filter(is.na(`latitude (deg)`), is.na(`longitude (deg)`)) %>% 
+austraits_composition_non_georef <- trait_composition_list %>%
+  filter(is.na(`latitude (deg)`), is.na(`longitude (deg)`)) %>%
   group_by(tissue, category) %>%
   summarise(n = n_distinct(taxon_name)) %>%
   rename(Category = category) %>%
-   drop_na() %>% 
-  ggplot(
-    aes(x = tissue, y = n, fill = Category)) +
+  drop_na() %>%
+  ggplot(aes(x = tissue, y = n, fill = Category)) +
[…]

### 2020-07-15 R/plotting_functions.R
added check for missing package and  automatic installation, load libraries within plotting_functions.R

@@ -1,3 +1,65 @@
+#### check for missing packages and load libraries####
+list_of_packages <-
+  c(
+    "rmarkdown",
+    "knitr",
+    "rprojroot",
+    "tidyverse",
+    "RefManageR",
+    "ape",
+    "shiny",
+    "RLumShiny",
+    "ggplot2",
+    "raster",
+    "RColorBrewer",
+    "dplyr",
+    "rlang",
+    "tibble",
+    "hexbin",
+    "phytools",
+    "tidytree",
+    "ggpointdensity",
+    "viridis",
+    "scales",
+    "ggstance",
+    "stringi",
+    "cowplot"
+  )
+new_packages <-
+  list_of_packages[!(list_of_packages %in% installed.packages()[, "Package"])]
+
+if (length(new_packages))
+  install.packages(new_packages)
+
+# load libraries
+lapply(list_of_packages, library, character.only = TRUE)
+
+if(!require("V.PhyloMaker")){
+  devtools::install_github("jinyizju/V.PhyloMaker")
+  library( "V.PhyloMaker")
+}
+
+if(!require(patchwork)){
+  devtools::install_github("thomasp85/patchwork")
+  library( patchwork)
+}
+
+if(!require(plotbiomes)){
+  devtools::install_github("valentinitnelav/plotbiomes")
+  library(plotbiomes)
+}
+
+if(!require(ggtree)){
+  devtools::install_github("GuangchuangYu/ggtree")
+  library(ggtree)
+}
+if(!require(treeio)){
+  devtools::install_github("GuangchuangYu/treeio")
+  library(treeio)
+}
+
+
+##### Functions ####
 extract_climate_data<-function(df, climstack){
   #df is a dataframe of trait data with lat and lon columns
   #climstack is a stack of gridded climate data

### 2020-07-15 scripts/fig_climate_space_austraits.R

@@ -1,13 +1,4 @@
 ## Austraits: Precipitation–temperature space plot
-library(raster)
-library(tidyverse)
-library(hexbin)
-library(ggplot2)
-library(RColorBrewer)
-library(plotbiomes)
-
-source("R/plotting_functions.R")
-
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)
 bioclim <- getData("worldclim", var = "bio", res = 10)

### 2020-07-15 scripts/fig_site_locations.R

@@ -1,13 +1,4 @@
 ## Austraits: site locations plot
-library(raster)
-library(tidyverse)
-library(ggplot2)
-library(RColorBrewer)
-library(viridis)
-library(ggpointdensity)
-source("R/plotting_functions.R")
-
-
 
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)

### 2020-07-15 ms.Rmd
Update the figures and their captions

@@
 - \usepackage{float}
+- \usepackage{fontspec}
+- \usepackage{unicode-math}
+# - \setmainfont{Arial}
+# - \setmathfont{Cambria Math}
 bibliography: [data/references_ms.bib]
@@
 
-We achknowelde XXXXXXXXXXx institutions, who
+We acknowledge XXXXXXXXXXx institutions, who
 
@@
 
-![**Figure 1.
-The data curation pathway used to assemble the AusTraits database.** Trait observations are accessed from original data sources, including published floras and field campaigns.
+![The data curation pathway used to assemble the AusTraits database.** Trait observations are accessed from original data sources, including published floras and field campaigns.
 Factors such as units and taxonomy are harmonised to a common standard.
@@
 
-![**Figure 2.** (a.) Locations where the trait data were collected and the corresponding number of taxa.
-(b.) AusTraits’ sites (grey) within Australia’s precipitation-temperature space (black) superimposed upon Whittaker’s major biomes diagram.
-Data from WorldClim (worldclim.org).
-(c.) Overview of the number of taxa with which traits are measured within each major plant organ and trait category.](figures/austraits_fig2_blue.png)
+![(a.) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s major biomes diagram.
+Climate data were extracted from Worldclim (www.worldclim.org).(B.) Locations of georeferenced-sites grouped by plant tissues ](figures/austraits_fig_2_biome_map.pdf)
 
-![**Figure 3.** The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
-The heatmap colour intensities denote the number of traits measured within each major plant organ.
-The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_phlogenetic_coverage.png)
+![Overview of the number of taxa with which traits are measured within each plant tissue and trait category; (A.) accounts for the georeferenced trait data (correspond with Fig 2B) whereas (B.) for the non-georeferenced trait data.
+](figures/austraits_fig_3_composition.pdf)
+
+![The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
+The heatmap colour intensity denotes the number of traits measured within each plant tissue.
+The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_fig_4_austraits_phlogenetic_coverage.pdf)
 

### 2020-07-15 build_figures.R

@@ -1,6 +1,6 @@
-#### Figures for the Austraits' manuscript ####
+#### Figures for the Austraits manuscript ####
 
-library(patchwork)
+source("R/plotting_functions.R") # Load required packages and functions
 
 #### Figure 2: geographical and biome map####
 source("scripts/fig_site_locations.R") # A
@@ -14,17 +14,17 @@ austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annot
 dev.off()
 
 #### Figure 3: trait composition count ####
-source("scripts/fig_trait_composition.R") # C
+source("scripts/fig_trait_composition.R")
 pdf("figures/austraits_fig_3_composition.pdf",
     height = 6,
     width = 8)
 austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'A')
 dev.off()
 
-#### Figure 4: Phylogenetic representation of AusTraits ####
-source("scripts/fig_generate_phylogenies.R") # A
+#### Figure 4: Phylogenetic representation of AusTraits  ####
+source("scripts/fig_generate_phylogenies.R") # A Based on 2000 randomly-sampled species
 source("scripts/fig_get_family_node.R") # B
-source("scripts/fig_link_tree_and_trait_count_matrix.R") # B
+source("scripts/fig_link_tree_and_trait_count_matrix.R") # C
 
 
 pdf(

### 2020-07-15 scripts/fig_generate_phylogenies.R

@@ -1,24 +1,5 @@
 ## 1 Build a phylogenetic tree from the list of species in AusTraits
 
-
-##### 0 Load libraries ####
-# check for missing packages
-list.of.packages <- c("phytools", "V.PhyloMaker", "tidyverse")
-new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-if(length(new.packages)) install.packages(new.packages)
-
-# load libraries
-library(phytools)
-library("V.PhyloMaker"); #devtools::install_github(jinyizju/V.PhyloMaker)
-library(tidyverse)
-library(ggtree)
-library(treeio)
-library(ggplot2)
-library(tidytree)
-
-
-
-
 ##### 1 Read data #####
 ## add austraits and read the lookup table
 austraits <- readRDS("data/austraits.rds")

### 2020-07-15 scripts/fig_trait_composition.R

@@ -1,11 +1,4 @@
 ## Austraits: trait composition plot
-
-library(tidyverse)
-library(ggplot2)
-library(scales)
-library(RColorBrewer)
-
-
 ## Rearrange order of tissue
 positions <-
   c("leaf", "stem", "bark", "root", "reproductive", "whole_plant")

### 2020-07-21 R/get_sites.R
changed austraits.rds to austraits_2.0.0.rds
uploaded data/traits_filled_in.csv

@@ -1,4 +1,4 @@
-austraits <- readRDS("data/austraits.rds")
+austraits <- readRDS("data/austraits_2.0.0.rds")
 library(tidyverse)
 
 # collect sites

### 2020-07-21 scripts/fig_generate_phylogenies.R

@@ -2,7 +2,7 @@
 
 ##### 1 Read data #####
 ## add austraits and read the lookup table
-austraits <- readRDS("data/austraits.rds")
+austraits <- readRDS("data/austraits_2.0.0.rds")
 #tree<-readRDS("data/AusTraits_phylogenies.rds") 
 trait_category_lookup<-read_csv("data/traits_filled_in.csv") %>% 
   dplyr::select(trait_name, tissue, category)

### 2020-07-21 data/2020.05.28-TRY/AusTraits.Rmd

TRY

```{r}
library(tidyverse)

bind_rows(
read_tsv("2020.05.28-TRY/tde202052811159.txt", skip=3),
read_tsv("2020.05.28-TRY/tde2020528113232.txt", skip=3)
) %>% distinct() -> TRY
```

```{r}
TRY %>% select(AccSpeciesName, Trait) %>% distinct()
TRY %>% select(AccSpeciesName) %>% distinct()
TRY %>% select(Lat, Lon) %>% distinct()
```

Note, tonnes of repeated values, especially for categorical traits so total number trait values is misleading

```{r}
TRY %>% group_by(AccSpeciesName, Trait) %>% summarise(n=n()) %>% arrange(desc(n))
```

TRY - data for 3820 species, 46,328 trait x species combinations, 1448 locations

AUSTRAITS

```{r}
readRDS("austraits.rds") -> Austraits
Austraits$traits %>% select(species_name, trait_name) %>% distinct()
Austraits$traits$species_name %>% unique()
Austraits$sites
```

Austraits -- data over 650000 records, for 339 traits, for 23,480 species. 282000 trait x species combinations, 22,000 locations

### 2020-07-21 R/taxonlookup.R
removed unused functions and the previous trait classification table

@@ -1,14 +1,0 @@
-taxon_lookup_func<-function(df){
-  df[["species_name"]]%>%
-    unique()%>%
-    lookup_table(., by_species=TRUE)%>%
-    tibble::rownames_to_column(., var = "species_name")-> taxon_result
-  
-  taxon_result
-}
-#function to join main data with the taxon_lookup_func outcome
-add_taxon_lookup<-function(df){
-  taxon_lookup_func(df)%>%
-    left_join(df, ., by="species_name")
-  
-}

### 2020-10-28 ms.Rmd
Rework: Adding Description file, combine sources, misc changes

@@
 ---
-title: "Austraits -- a curated plant trait database for the Australian flora"
+title: "AusTraits -- a curated plant trait database for the Australian flora"
 csl: assets/nature.csl
@@
 # - \setmathfont{Cambria Math}
-bibliography: [data/references_ms.bib]
+bibliography: [data/references_all.bib]
 ---
@@
 austraits <- readRDS("data/austraits_2.0.0.rds")
+
+sources_all <- 
+  bibtex::read.bib("data/references_ms.bib")
+
+RefManageR::WriteBib(austraits$sources, "data/references_data.bib")
+
+c(readLines("data/references_ms.bib"), "\n", readLines("data/references_data.bib")) %>% writeLines("data/references_all.bib")
+
 ```
@@
 
+# check list of acknowledgements and assistants
+
 data_contributors <- 
@@
     filter(
-          # old studies
+          # old studies or excluded
           !dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981"),
@@
            # deceased
-           !name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper")
+           !name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper"),
+            # no response or can't locate
+           !name %in% c("Huw Morgan", "Wendy Cooper", "Muhammad Islam", "Michael Lawes")
            ) %>% 
@@
 
-#  authors %>% select(dataset_ids, everything()) %>% mutate_all(~ replace_na(.x, "")) %>% write_csv("data/author_list.csv")
+authors %>% select(dataset_ids, everything()) %>% mutate_all(~ replace_na(.x, "")) %>% write_csv("data/author_list.csv")
 
@@
 going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
-
-## Todo
-
-Refs
-
-* Australian Plant Census, Centre of Australian National Biodiversity Research, Council of Heads of Australasian Herbaria, (14/05/2020) https://id.biodiversity.org.au/tree/51354547
-* Australian Plant Name Index (continuously updated), Centre of Australian National Biodiversity Research, www.biodiversity.org.au/nsl/services/apni (14/05/2020)
 
@@
 
+## Comparison to TRY
+
+numbers of taxa / figures
+
+* AusTraits includes more seed and fruit nutrient data; possibly reflecting the interest in Australia, specifically WA, in understanding how fruit and seeds are provisioned in nutrient depauperate environments
+* AusTraits includes more categorical values, especially variables documenting different components of species’ fire response strategies, reflecting the importance of fire in shaping Australian communities and the research to document different strategies species have evolved to succeed in fire-prone environments
+
+- On the point about many Australian species being unrepresented in global trait databases (end of the second paragraph in Background & Summary) -- I was wondering if this could be exemplified by reporting the number of species in AusTraits compared to the number of Australian species in the TRY Plant Database (in order to show that AusTraits fills this gap substantially).
+- I was also wondering if the number of studies that have used global plant trait databases could be reported as an example to show how valuable AusTraits will be (e.g.
+TRY Plant Trait Database having 1613 citations according to Google Scholar) (relevant to the point that is made in the second-last sentence of Background & Summary).
+
 ### Methodology
@@
 
-AusTraits was assembled from `r austraits$sources %>% length()` distinct sources, including published literature, field campaigns, floras, and taxonomic treatments.
-We identified a list of candidate traits of interest, including growth form, plant height, leaf size, specific leaf area and seed mass, though several other traits were also collected opportunistically.
+AusTraits was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments.
+Initially we identified a list of candidate traits of interest -- including growth form, plant height, leaf size, specific leaf area and seed mass -- then identified primary sources containing measurements for these trait, before contacting authors for access.
+As the compilation grew, we expanded the list of traits considered to include any measureable quantity that had been quantified for a moderate number of taxa ( n > 20).
 A full list of traits and their sources appears in Table 1 (available online only).
-We then identified primary sources containing measurements for these traits and sought access to this data, including both data and metadata.
-Each original source was designated a `dataset_id`.
-Ordinarily the id is the surname of first author and year of publication, for the source’s primary citation (e.g.
-Falster_2005, for Dataset citation 40).
 Details from each source were saved with minimal modification into two plain text files:
@@
 
+```{r, results="asis"}
+austraits$sources %>% sapply(function(i) i$key) -> keys
+
+sprintf("@%s",keys) %>% paste(collapse = "; ") %>% paste0("[", ., "]") %>% writeLines()
+```
+
+### Trait definitions
+
+TOP: [Garnier-2017]
+
+### Taxonomy
+
+Taxonomy updated via Australian Plant Census (APC) [@apc-2020], Australian Plant Names Index (APNI) [@apni-2020]
+
 #### Schema
+
+Each original source was designated a `dataset_id`.
+Ordinarily the id is the surname of first author and year of publication, for the source’s primary citation (e.g.
+Falster_2005, for Dataset citation 40).
 
@@
 To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R [@R-2020] and using custom code and drawing on functions from the packages `tidyverse` [@wickham2016tidyverse], `stringr` [@wickham2017stringr], `yaml` [@stephens2014yaml], `remake` [@fitzjohn2016remake],  `knitr` [@xie1general] and `rmarkdown` [@allaire2015rmarkdown].
-In this workflow, we performed a series of operations, including reformatting data into standardising format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorial variables, encoding suitable metadata, and flagging data that did not pass quality checks.
+In this workflow, we performed a series of operations, including reformatting data into standardising format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
 
@@
 Whilst all care is taken to curate datasets in AusTraits this is not intended to absolve individual researchers from ensuring the appropriate use of the data provided.
+
+#### Analysis
+
+Phylogenetic tree constructed by pruning  [@Smith-2018] using package `V.PhyloMaker` [@Jin-2020] and visualising using the package `ggtree` [@Yu-2017]
 
@@
 DSF was supported by Australian Research Council Future Fellowship (FT160100113), RVG was supported by Australian Research Council DECRA Fellowship (DE170100208).
+This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright, and funding from the Australian Research Data Commons under their "Transformative data collections" program.
 
-We acknowledge XXXXXXXXXXx institutions, who
+We acknowledge XXXXXXXXXXx institutions, who XXXXX
 
@@
 RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, CB, JL, SA error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
-All other authors contributed the raw data and metadata undeprinning the resource and reviewed the harmonised data for errors.
+All other authors contributed the raw data and metadata underpinning the resource and reviewed the harmonised data for errors.
 
@@
 ## Assumes to items exist in global name space
-austraits <- readRDS("data/austraits_2.0.0.rds")
-
-definitions <- austraits$definitions
-
-tmp <- RefManageR::WriteBib(austraits$sources, "data/references_data.bib")
 
@@
   bind_rows(
-    austraits$methods %>% select(dataset_id, citation = source_primary_key) %>% 
-            distinct() %>% mutate(type="primary"),
-    austraits$methods %>% select(dataset_id, citation = source_secondary_key) %>% 
-            distinct() %>% mutate(type="secondary")
+    austraits$methods %>% select(dataset_id, citation = source_primary_key),
+    austraits$methods %>% select(dataset_id, citation = source_secondary_key)
   ) %>% 
+  na_if("") %>%
   na.omit() %>% 
+  distinct() %>% 
   arrange(dataset_id) %>% 
   group_by(dataset_id) %>% 
-  summarise(refs = sprintf("@%s", citation) %>% 
-  paste(collapse = "; "))
+  summarise(refs = sprintf("@%s", citation) %>% str_replace_all("; ", "; @") %>% paste(collapse = "; ") )
 
@@
 
-<!-- Cite: -->
-  <!-- G Yu, DK Smith, H Zhu, Y Guan, TTY Lam*.
-ggtree: an R package for visualization and annotation of phylogenetic trees with their covariates and other associated data.
-Methods in Ecology and Evolution.
-2017, 8(1):28-36.
-doi: 10.1111/2041-210X.12628.
--->
-  <!-- G Yu*, TTY Lam, H Zhu, Y Guan*.
-Two methods for mapping and visualizing associated data on phylogeny using ggtree.
-Molecular Biology and Evolution, 2018, 35(2):3041-3043.
-doi: 10.1093/molbev/msy194.
--->
-
 ## Tables
@@
 ```
+
+\newpage
 
@@
             Families = n_distinct(family),
-            `Data Citations` = refs %>% unique() %>% paste(collapse = "; ")
+            `Data Citations` = refs %>% unique() %>% paste(collapse = "; ") %>% paste0("[", ., "]")
             ) %>% 
@@
 n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$Trait]
-n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$Trait]
+n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$Trait] %>% unlist()
 
 n_records %>%
-  select(-Tissue, Description) %>%
-  kable()
+  select(-Tissue, -Description) %>%
+  knitr::kable()
 ```
@@
 
-### Data sources:
-
-#### Version 1 - via bib.print
-
-```{r, results='asis', eval=TRUE}
-sources_unique <-
-  bind_rows(
-    austraits$methods %>% select(citation = source_primary_citation)  %>% mutate(type="primary"),
-    austraits$methods %>% select(dataset_id, citation = source_secondary_citation) %>% mutate(type="secondary")
-  ) %>%
-  select(citation) %>%
-            distinct() %>% 
-            mutate(source_number=1:n())
-
-sprintf("%s. %s", sources_unique$source_number, sources_unique$citation) %>% paste(collapse = "\n\n") %>% writeLines()
-
-```
-
-#### Version 3 - via csl
-
-```{r, results='asis', eval=FALSE}
-
-austraits$sources %>% sapply(function(i) i$key) -> keys
-
-sprintf("@%s",keys) %>% paste(collapse = "; ") %>% writeLines()
-```
-
 ### References

### 2020-10-28 R/get_sites.R

@@ -1,5 +1,3 @@
-austraits <- readRDS("data/austraits_2.0.0.rds")
-library(tidyverse)
 
 # collect sites
 austraits$sites %>%
@@ -21,7 +19,7 @@ austraits$traits %>%
   ungroup() -> sp_number
 
 # get trait categories
-trait_category_lookup <- read_csv("data/traits_filled_in.csv") %>%
+trait_category_lookup <- read_csv("data/trait_categories.csv") %>%
   dplyr::select(trait_name, tissue, category)
 
 # get trait list

### 2020-10-28 R/plotting_functions.R

@@ -1,63 +1,3 @@
-#### check for missing packages and load libraries####
-list_of_packages <-
-  c(
-    "rmarkdown",
-    "knitr",
-    "rprojroot",
-    "tidyverse",
-    "RefManageR",
-    "ape",
-    "shiny",
-    "RLumShiny",
-    "ggplot2",
-    "raster",
-    "RColorBrewer",
-    "dplyr",
-    "rlang",
-    "tibble",
-    "hexbin",
-    "phytools",
-    "tidytree",
-    "ggpointdensity",
-    "viridis",
-    "scales",
-    "ggstance",
-    "stringi",
-    "cowplot"
-  )
-new_packages <-
-  list_of_packages[!(list_of_packages %in% installed.packages()[, "Package"])]
-
-if (length(new_packages))
-  install.packages(new_packages)
-
-# load libraries
-lapply(list_of_packages, library, character.only = TRUE)
-
-if(!require("V.PhyloMaker")){
-  devtools::install_github("jinyizju/V.PhyloMaker")
-  library( "V.PhyloMaker")
-}
-
-if(!require(patchwork)){
-  devtools::install_github("thomasp85/patchwork")
-  library( patchwork)
-}
-
-if(!require(plotbiomes)){
-  devtools::install_github("valentinitnelav/plotbiomes")
-  library(plotbiomes)
-}
-
-if(!require(ggtree)){
-  devtools::install_github("GuangchuangYu/ggtree")
-  library(ggtree)
-}
-if(!require(treeio)){
-  devtools::install_github("GuangchuangYu/treeio")
-  library(treeio)
-}
-
 
 ##### Functions ####
 extract_climate_data<-function(df, climstack){
@@ -71,8 +11,8 @@ extract_climate_data<-function(df, climstack){
     dplyr::select( lat, lon)%>%
     na.omit()-> coord
   
-  coordinates(coord)=~lon+lat
-  proj4string(coord)=CRS("+proj=longlat +datum=WGS84")
+  sp::coordinates(coord)=~lon+lat
+  sp::proj4string(coord)=sp::CRS("+proj=longlat +datum=WGS84")
   raster::extract(climstack,coord,na.rm=F, df=T)->df_climate
   
   df_climate

### 2020-10-28 build_figures.R

@@ -1,6 +1,9 @@
 #### Figures for the Austraits manuscript ####
 
+devtools::load_all()
+
 source("R/plotting_functions.R") # Load required packages and functions
+
 
 #### Figure 2: geographical and biome map####
 source("scripts/fig_site_locations.R") # A
@@ -14,6 +17,7 @@ austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annot
 dev.off()
 
 #### Figure 3: trait composition count ####
+
 source("scripts/fig_trait_composition.R")
 pdf("figures/austraits_fig_3_composition.pdf",
     height = 6,
@@ -38,6 +42,10 @@ dev.off()
 
 #### Figure 5: AusTraits and TRY comparison ####
 source("scripts/fig_comparison_TRY.R") 
+ 
+# aus_try %>%
+#   filter(n_try > 1, n_austraits >1, tissue=="leaf") %>%
+#   arrange(desc(n_try/n_austraits)) %>% View
 
 pdf( "figures/austraits_fig_5_austraits_try_comparison_v1.pdf",
      height = 4,
@@ -45,15 +53,8 @@ pdf( "figures/austraits_fig_5_austraits_try_comparison_v1.pdf",
 p_aus_try
 dev.off()
 
-pdf( "figures/austraits_fig_5_austraits_try_comparison_v2.pdf",
-     height = 5,
-     width = 6)
-ggExtra::ggMarginal(p_aus_try, type = "density", groupColour = T)
-dev.off()
-
 pdf("figures/austraits_fig_5_austraits_try_comparison_v3.pdf",
     height = 8,
     width = 10)
 p_aus_try + facet_wrap( ~tissue)
 dev.off()
-

### 2020-10-28 scripts/fig_climate_space_austraits.R

@@ -1,24 +1,24 @@
 ## Austraits: Precipitation–temperature space plot
 #### 01 Get climate data for Australia ####
 # Download bioclim data using library (raster)
-bioclim <- getData("worldclim", var = "bio", res = 10)
+bioclim <- raster::getData("worldclim", var = "bio", res = 10)
 # Pick BIO1 (Mean Annual Temperature; T) and BIO12 (Annual Precipitation; P)
 bioclim <- bioclim[[c(1, 12)]]
 names(bioclim) <- c("Temp", "Prec")
 
 #### 02 Get the climate data for Australia ####
 # Load Australia landmass binary map
-au_map <- raster("data/australia.tif")
+au_map <- raster::raster("data/australia.tif")
 # Clip bioclim data with the au map
 ## crop and mask
-bioclim_au <- crop(bioclim, extent(au_map)) #%>% mask(.,au_map)
+bioclim_au <- raster::crop(bioclim, raster::extent(au_map)) #%>% mask(.,au_map)
 new.bioclim <-
-  projectRaster(bioclim_au, au_map) # harmonize the spatial extent and projection
+  raster::projectRaster(bioclim_au, au_map) # harmonize the spatial extent and projection
 au_bioclim <- raster::mask(new.bioclim, au_map)
 
 # Transform raster data into a tibble
 au_bioclim_table <- au_bioclim %>%
-  as.data.frame() %>%
+  raster::as.data.frame() %>%
   na.omit() %>%
   as_tibble %>%
   mutate(region = as.factor("Australia"))

### 2020-10-28 scripts/fig_comparison_TRY.R

@@ -10,7 +10,7 @@ TRY <- bind_rows(
 # AUSTRAITS -- data over 650000 records,
 # for 339 traits, for 23,480 species. 282000 trait x species combinations, 22,000 locations
 austraits <- readRDS("data/austraits_2.0.0.rds")
-tissue_lookup <- read_csv("data/traits_filled_in.csv") %>%
+tissue_lookup <- read_csv("data/trait_categories.csv") %>%
   select(trait_name, `TRY name`, tissue)
 
 
@@ -23,14 +23,17 @@ try_by_sp <- TRY %>%
   group_by(Trait) %>%
   summarise(n_try = n_distinct(AccSpeciesName))
 
-
 aus_try <- tissue_lookup %>%
   full_join(aus_trait_by_taxon, by = "trait_name") %>%
   full_join(try_by_sp, by = c("TRY name" = "Trait")) %>%
   mutate_if(is.numeric, ~ replace_na(., 0)) %>%
-  mutate(tissue = replace_na(tissue, "uncategorised"))
-
-
+  mutate(tissue = replace_na(tissue, "uncategorised")) %>% 
+  group_by(`TRY name`) %>% 
+  summarise(
+    n_austraits = sum(n_austraits), 
+    n_try = n_try[1],
+    tissue = subset(tissue, !is.na(tissue))[1]
+  ) %>% ungroup()
 
 #==== 03 Plotting ====
 # Scatter plot

### 2020-10-28 scripts/fig_generate_phylogenies.R

@@ -4,7 +4,7 @@
 ## add austraits and read the lookup table
 austraits <- readRDS("data/austraits_2.0.0.rds")
 #tree<-readRDS("data/AusTraits_phylogenies.rds") 
-trait_category_lookup<-read_csv("data/traits_filled_in.csv") %>% 
+trait_category_lookup<-read_csv("data/trait_categories.csv") %>% 
   dplyr::select(trait_name, tissue, category)
 
 
@@ -15,11 +15,12 @@ n_species_tree <- 2000
 ## generate list of species from austraits (or sample 2000 species)
 set.seed(10)
 austraits$taxa %>% 
-  select(taxon_name, genus, family) %>% 
+  dplyr::select(taxon_name, genus, family) %>% 
   filter(!is.na(family))%>% 
   mutate(taxon_name = gsub(" ", "_", taxon_name)) %>% 
   sample_n(n_species_tree) -> austraits_sp_list_sampled
 
+## Can we use ape::drop.tip instead?
 
 # Build a tree using V.PhyloMaker (probably in a newick format)
 tree <- phylo.maker(sp.list = austraits_sp_list_sampled,

### 2020-10-28 scripts/fig_site_locations.R

@@ -9,9 +9,9 @@ if (!file.exists(filename))
   download.file("htpp:....", filename)
 
 # Read bioclim data using raster package
-au_map <- raster(filename) %>%
+au_map <- raster::raster(filename) %>%
   # aggregate(fact=6) %>%
-  as.data.frame(xy = T)
+  raster::as.data.frame(xy = T)
 
 
 #### 02 Plot AU basemap ####

### 2020-11-03 ms.Rmd
Major revision
- figure comparing to TRY
- use natbib to format references
- fix all tables
- rework text
- figure comparing to TRY

--- 2020-07-09 ms.Rmd
@@
 ---
-title: "Austraits -- a curated plant trait database for the Australian flora"
-csl: assets/nature.csl
+title: "AusTraits -- a curated plant trait database for the Australian flora"
 editor_options:
@@
     latex_engine: xelatex
-  html_document:
-    df_print: paged
-  word_document:
-    reference_docx: assets/manuscriptstyle.docx
+    citation_package: natbib
+    keep_tex: true
+    keep_md: true
+biblio-style: nature
 header-includes:
 - \usepackage{inputenc}
-- \usepackage[T1]{fontenc}
+#- \usepackage[T1]{fontenc}
 - \usepackage{float}
-bibliography: [data/references_ms.bib]
+- \usepackage{fontspec}
+- \usepackage{unicode-math}
+- \usepackage{booktabs}
+- \usepackage{longtable}
+- \usepackage{array}
+- \usepackage{multirow}
+- \usepackage{wrapfig}
+- \usepackage{colortbl}
+- \usepackage{lscape}
+- \usepackage{tabu}
+- \usepackage{threeparttable}
+- \usepackage{threeparttablex}
+- \usepackage[normalem]{ulem}
+- \usepackage{makecell}
+- \usepackage{xcolor}
+bibliography: [data/references.bib]
+
 ---
@@
 library(rmarkdown)
+source("R/ms.R")
+source("R/load_data.R")
+load_data()
 
-source("R/ms.R")
+RefManageR::WriteBib(austraits$sources, "data/references_data.bib")
 
-austraits <- readRDS("data/austraits_2.0.0.rds")
+c(
+  readLines("data/references_ms.bib"), "\n", 
+  readLines("data/references_data.bib")) %>% 
+  gsub("&", "and", ., fixed=TRUE) %>% 
+  writeLines("data/references.bib")
 ```
 
-```{r authors, results='asis', echo=FALSE}
-authors_build <- read_csv("data/lead_authors.csv", col_types = cols(.default = "c"))
+```{r authors}
+authors_lead <- read_csv("data/lead_authors.csv", col_types = cols(.default = "c"))
+
+# check list of acknowledgments and assistants
 
@@
   austraits$contributors %>% 
-    filter(
-          # old studies
-          !dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981"),
-           !role %in% c("assistant"),
+    mutate(name = gsub("Richard Duncan", "Richard P. Duncan", name, fixed=TRUE)) %>%
+    mutate(
+      acknowledge = ifelse(
+          # old studies or excluded
+          dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981") |
+           role %in% c("assistant") |
            # deceased
-           !name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper")
-           ) %>% 
-    mutate(year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
+           name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper") |
+            # no response or can't locate
+           name %in% c("Emma Laxton", "Huw Morgan", "Wendy Cooper", "Muhammad Islam", "Michael Lawes"),
+            TRUE, FALSE),
+      year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
     arrange(year) %>%
@@
     summarise(institution = last(institution),
-              dataset_ids = paste(dataset_id, collapse = ", ")) %>% 
+              dataset_ids = paste(dataset_id, collapse = ", "),
+              acknowledge = all(acknowledge)) %>% 
     ungroup() %>% 
@@
     rename(affiliation = institution) %>%
-#    bind_rows(
-#      read_csv("data/data_pending.csv")) %>%
     arrange(surname) %>% 
@@
 
-authors <- bind_rows(authors_build, 
-                     data_contributors %>% filter(! (name %in% authors_build$name))
+authors <- bind_rows(authors_lead, 
+                     data_contributors %>% filter(! (name %in% authors_lead$name), !acknowledge)
                     )
 
-#  authors %>% select(dataset_ids, everything()) %>% mutate_all(~ replace_na(.x, "")) %>% write_csv("data/author_list.csv")
+acknowledge <- data_contributors %>% filter(! (name %in% authors$name), acknowledge)
+
+authors %>% select(dataset_ids, everything()) %>% mutate_all(~ replace_na(.x, "")) %>% write_csv("data/author_list.csv")
 
@@
 
-### Abstract
+## Abstract
 
@@
 
-## Todo
-
-Refs
-
-* Australian Plant Census, Centre of Australian National Biodiversity Research, Council of Heads of Australasian Herbaria, (14/05/2020) https://id.biodiversity.org.au/tree/51354547
-* Australian Plant Name Index (continuously updated), Centre of Australian National Biodiversity Research, www.biodiversity.org.au/nsl/services/apni (14/05/2020)
-
 ## Background and Summary
 
-Species traits are essential metrics for comparing ecological strategies in plants arrayed across environmental space or evolutionary lineages [@diaz2016global; @kunstler2016plant; @zanne2014three; @cornwell2014functional].
-Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function [@chapin1993evolution; @adler2014functional; @diaz1998plant].
-Traits thereby provide useful indicators of species behaviours in communities and ecosystems, regardless of their taxonomy [@violle2007let; @funk2017revisiting].
-Through targeted global initiatives the volume of available trait information for plants has grown rapidly in the last two decades.
+Species traits are essential metrics for comparing ecological strategies in plants arrayed across environmental space or evolutionary lineages \citep{diaz2016global, kunstler2016plant, zanne2014three, cornwell2014functional}.
+Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function \citep{chapin1993evolution, adler2014functional, diaz1998plant,violle2007let}.
+Traits thereby provide useful indicators of species behaviours in communities and ecosystems, regardless of their taxonomy \citep{violle2007let, funk2017revisiting}.
+Through global initiatives the volume of available trait information for plants has grown rapidly in the last two decades \citep{Kattge2011try, Kattge-2020}.
 However, the geographic coverage of trait observations across the globe is patchy, limiting detailed analyses of trait variation and diversity in some regions.
+
 One such region is Australia; a continent with a flora of c.
-24,000 native higher-plant species [@council2010australian].
+28,000 native higher-plant species \citep{council2010australian}.
 While significant investment has been made in curating and digitising herbarium collections in Australia over the last two decades (e.g.
 The Australian Virtual Herbarium houses ~7 million specimen occurrence records; https://avh.ala.org.au), no complementary resource yet exists for consolidating information on plant traits.
+Moreover, relatively few species are represented in the leading global databases.
+For example, the TRY international database has only 3830 species represented in total across all traits measured.
+Such low coverage greatly limits the use of traits in understanding and managing our vegetation.
+While initiatives, such as the TRY database \citep{Kattge-2020} and Open Traits Network \citep{Gallagher-2020}, have been working towards sharing of data about plant traits, though most Australian plant taxa remain unrepresented.
 
-Here we introduce the AusTraits database (hereafter simply AusTraits) - a compilation of plant traits for the Australian flora.
-Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources – including published scientific papers, floras, and taxonomic descriptions– and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa (Table 1; available online only).
+Here we introduce the AusTraits database (hereafter simply AusTraits),  a compilation of plant traits for the Australian flora.
+Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources -– including published scientific papers, floras, and taxonomic descriptions –- and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
+To assemble AusTraits from diverse primary sources and make data available for reuse, we needed to overcome three main types of challenges (Figure 1): 1) Accessing data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) Harmonising these diverse sources into a federated resource, with common units, trait names, and data formats; and 3) Distributing versions of the data under suitable license.
+To meet this challenge, we developed a workflow for harmonising and distributing data, drawing from emerging community standards and our collective experience building traits databases.
+
 By providing a harmonised and curated dataset on `r austraits$traits$trait_name %>% n_distinct()` plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
-Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives (but see @guerin2017opportunities).
-Global initiatives, such as the TRY database [@Kattge-2020] and Open Traits Network [@Gallagher-2019], have been working towards sharing of data about plant traits though large numbers of Australian plant taxa remain unrepresented.
+Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives \citep[but see][]{guerin2017opportunities}.
+We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to accelerate our collective understanding the Australian flora.
+Open access to a comprehensive resource like this will generate significant new knowledge about the Australian flora across multiple scales of interest, as well as reduce duplication of effort in the compilation of plant trait data, particularly for research students and government agencies seeking to access information on traits.
 
-To assemble AusTraits from diverse primary sources and make it available for reuse, we needed to overcome three main types of challenges (Figure 1).
-These included: 1) **Accessing** data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) **Harmonising** these diverse datasets into a federated resource, with common units, trait names, and data formats; and 3) **Distributing** versions of the data under suitable license.
-Successive versions of AusTraits will iterate through these to incorporate new data and correct identified errors.
-We developed a workflow for harmonising and distributing data, drawing from emerging community standards and our collective experience building traits databases.
-Combined, the tools applied ensure current best-practices in the construction and distribution of data resources, by achieving the following outcomes:
+## Methods
 
-  - A reproducible and extensible compilation process from raw data to assembled dataset;
-  - Transparent mapping of original values (taxon names, trait names, units) into the harmonised dataset;
-  - High quality data for end users; 
-  - Easy addition of new data and correction of community-reported errors; 
-  - Satisfying the FAIR principles [@wilkinson2016fair], i.e. standards needed to make data resources Findable, Accessible, Interoperable, and Reusable; and
-  - Recognition of contributors through citation of this resource and primary sources.
+### Primary sources
 
-We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to accelerate our collective understanding the Australian flora.
-It is anticipated that open access to a comprehensive resource like AusTraits will generate significant new knowledge about the Australian flora across multiple scales of interest.
-AusTraits will likely reduce duplication of effort in the compilation of plant trait data, particularly for research students and government agencies seeking to create trait-based answers to their questions.
+AusTraits Version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments (see Online Table 1).
+Initially we identified a list of candidate traits of interest then identified primary sources containing measurements for these trait, before contacting authors for access.
+As the compilation grew, we expanded the list of traits considered to include any measureable quantity that had been quantified for a moderate number of taxa ( n > 20).
 
-### Methodology
+### Trait definitions
 
-#### Primary sources
+A full list of traits and their sources appears in Table 10 (available online only).
+This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics \citep{Garnier-2017}.
+We categorised traits based on the tissue where it is measured (`r trait_categories$tissue %>% unique() %>% sort() %>% paste(collapse=", ")`) and the type of measurement (`r trait_categories$category %>% unique() %>% sort() %>% paste(collapse=", ")`).
+Version `r austraits$build_info$version` of AusTraits includes  `r n_records$Type %>% nrow` of type: `r n_records$Type %>% table() %>% paste(., names(.))`.
 
-AusTraits was assembled from `r austraits$sources %>% length()` distinct sources, including published literature, field campaigns, floras, and taxonomic treatments.
-We identified a list of candidate traits of interest, including growth form, plant height, leaf size, specific leaf area and seed mass, though several other traits were also collected opportunistically.
-A full list of traits and their sources appears in Table 1 (available online only).
-We then identified primary sources containing measurements for these traits and sought access to this data, including both data and metadata.
-Each original source was designated a `dataset_id`.
-Ordinarily the id is the surname of first author and year of publication, for the source’s primary citation (e.g.
-Falster_2005, for Dataset citation 40).
-Details from each source were saved with minimal modification into two plain text files:
+### Database schema
 
-1.	`data.csv`: a table containing the actual trait data in comma-separated values format;
-2.	`metadata.yml`: contains relevant metadata for the study, as well as options for mapping trait names and units onto standard types, and any substitutions applied to the data in processing.
-
-These two files provide all the information needed to compile each study into a standardised format.
-
-#### Schema
-
-The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology [@madin2007ontology] in that, where available, trait data are connected to contextual information about the collection site (e.g.
+The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection site (e.g.
 location co-ordinates, soil conditions, number of replicates) and information about the methods used to derive measurements (e.g.
 number of replicates, equipment used where available).
-The target format contains eight core objects, as described in Table 2.
-The contents of several of these objects are described further in Tables 3-6.
-This format was developed to include sufficient information about the data and methods used to collect trait measurements in each source.
-Where possible, trait measurement was recorded as being one of several possible `value_type` (Table 7), reflecting the type of measurement recorded.
+The finished database contains 11 elements, as described in Table 1.
+This format was developed to include information about the trait measurements, taxa sampled, the methods used, sites, contextual information, the peopled involved, and citation sources.
 
-#### Harmonisation
+For storage efficiency, the main traits table contains relatively little information (Table 2), but can be cross linked against other tables (Tables 3-8) using several identifiers for dataset, site, context, observation and taxon (Table 1).
+The `dataset_id` is ordinarily the surname of first author and year of publication, for the source’s primary citation (e.g.
+`Blackman_2014`).
+Trait measurement were also recorded as being one of several possible `value_type` (Table 9), reflecting the type of measurement recorded.
 
-To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R [@R-2020] and using custom code and drawing on functions from the packages `tidyverse` [@wickham2016tidyverse], `stringr` [@wickham2017stringr], `yaml` [@stephens2014yaml], `remake` [@fitzjohn2016remake],  `knitr` [@xie1general] and `rmarkdown` [@allaire2015rmarkdown].
-In this workflow, we performed a series of operations, including reformatting data into standardising format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorial variables, encoding suitable metadata, and flagging data that did not pass quality checks.
+### Harmonisation
 
-To enhance data quality, we produced detailed reports for each variable and source.
-These reports were reviewed both by AusTraits compilers as well as data contributors.
-Whilst all care is taken to curate datasets in AusTraits this is not intended to absolve individual researchers from ensuring the appropriate use of the data provided.
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R \citep{R-2020} and using custom code and drawing on functions from the packages `tidyverse` \citep{wickham2016tidyverse}, `stringr` \citep{wickham2017stringr}, `yaml` \citep{stephens2014yaml}, `remake` \citep{fitzjohn2016remake},  `knitr` \citep{xie1general} and `rmarkdown` \citep{allaire2015rmarkdown}.
+In this workflow, we performed a series of operations, including reformatting data into a standardised format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
+Successive versions of AusTraits iterate through the steps in Figure 1, to incorporate new data and correct identified errors, leading to high-quality harmonised dataset.
 
-#### Code Availability
+Details from each primary source were saved with minimal modification into two plain text files.
+The first file, `data.csv`, contains the actual trait data in comma-separated values format.
+The second file, `metadata.yml`, contains relevant metadata for the study, as well as options for mapping trait names and units onto standard types, and any substitutions applied to the data in processing.
+These two files provide all the information needed to compile each study into a standardised AusTraits format.
 
-All code, raw and compiled data are hosted within a GitHub repository under the Trait Ecology and Evolution organisation (see https://github.com/traitecoevo/austraits.build, https://github.com/traitecoevo/austraits).
-Sequential versions of both source and compiled dataset are also being archived within Zenodo [@falster_2019].
-The archived material includes all data sources and code for rebuilding the compiled dataset.
+### Taxonomy
 
-#### Data Records
+We developed a custom workflow to clean and standardise taxonomic names using the latest and most comprehensive taxonomic resources for the Australian flora: the Australian Plant Census (APC) \citep{apc-2020} and the Australian Plant Names Index (APNI) \citep{apni-2020}.
+While several automated tools exist, such as `taxize` \citep{Chamberlain-2013}, these do not currently include up to date information for Australian taxa.
+Updates were completed in two steps.
+In the first step, we used both direct and then fuzzy matching (with up to 2 characters difference) to search for an alignment between reported names and those in three name sets: 1) All accepted taxa in the APC, 2) All known names in the APC, 3) All names in the APNI.
+Names were aligned without name authorities, as we found this information was rarely reported in the raw datasets provided to us.
+Second, we used the aligned name to update any outdated names to their current accepted name, using the information provided in the APC.
+If a name was recorded as being both an accepted name and an alternative (e.g.
+synonym) we preferred the accepted name, but also noting the alternative records.
+When a suitable match could not be found, we manually reviewed near matches and web portals such as the Atlas of Living Australia to find a suitable match.
+The final resource reports both the original and the updated taxon name alongside trait records (Table 2), as well as additional tables summarising all taxonomic names changes (Table 6) and further information from the APC and APNI on all taxa included (Table 7).
 
-A static version of the data contained in this descriptor has been released as version`r austraits$build_info$version` of AusTraits, available on the project website (Data Citation 1) and Zenodo (Data Citation 2).
-Details and references for all traits are summarised in Table 1 (available online only).
-The latest data can be downloaded directly from the project website https://github.com/traitecoevo/austraits.
-However, as validation (see Technical Validation, below) and data entry is ongoing, users are recommended to pull data from the static releases, to ensure results in their downstream analyses remain consistent as the database is updated.
+## Data records
 
-We envision AusTraits as a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
+Static versions of the AusTraits, including version`r austraits$build_info$version` used in this descriptor, are available on the project website (https://github.com/traitecoevo/austraits) and Zenodo (Data Citation 1).
+The latest data can also be downloaded directly from the project website.
+As validation (see Technical Validation, below) and data entry is ongoing, users are recommended to pull data from the static releases, to ensure results in their downstream analyses remain consistent as the database is updated.
+
+### Data coverage
+
+```{r}
+#As of May 2020, the APC included 28,981 accepted species, subspecies, varieties and forma.  
+```
+
+The number of vascular plants (species, subspecies, varieties and forma) in Australia is ~28000 \citep{apc-2020}.
+Version`r austraits$build_info$version` of AusTraits included at least one record for `r austraits$taxa$taxon_name %>% n_distinct()`, or about 80% of taxa.
+Five traits (leaf_length, leaf_width, plant_height, life_history, plant_growth_form) have records for more than 50% of taxa.
+Across all traits, the median number of taxa with records is 62.
+Table 10 shows the number of studies, taxa, and families recording data in AusTraits, as well as the number of geo-referenced records, for each trait.
+
+There were substantial differences in coverage among different tissues and trait types, also with respect to number of geo-referenced points (Figure 2).The most common traits are not geo-referenced recorded from floras.
+Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 2a).
+
+We found that trait records were spread across the climate space of Australia (Figure 3a), as well as geographic locations (Figure 3b).
+As with most data, in Australia, the density of records was somewhat concentrated around  cities, but particularly for leaf traits.
+
+Figure 4 shows that overall coverage across the phylogenetic tree is relatively unbiased, though there are some notable exceptions.
+One exception is for root traits, where the taxa within the family Poaceae have considerably more information available than other taxa.
+A cluster of taxa within the family Myrtaceceae have little leaf information available, while reproductive information is limited for species near the based ot the tree.
+
+Comparing coverage in AusTraits to the global database TRY, there were `r aus_try %>% filter(n_austraits >0, n_try >0)%>% nrow()` traits overlapping.
+Of these, AusTraits tended to contain records for more taxa, but not always (Figure 5).
+Multiple traits had more than 10 times the number of taxa represented in AusTraits.
+However, there were more records in TRY for `r aus_try %>% filter(n_try > n_austraits, n_austraits >0) %>% nrow()` traits, in particular quite physiological leaf traits.
+Many traits were not overlapping between the two databases (Figure 5).
+We noted that AusTraits includes more seed and fruit nutrient data; possibly reflecting the interest in Australia in understanding how fruit and seeds are provisioned in nutrient depauperate environments.
+AusTraits includes more categorical values, especially variables documenting different components of species’ fire response strategies, reflecting the importance of fire in shaping Australian communities and the research to document different strategies species have evolved to succeed in fire-prone environments.
+
+## Technical Validation
+
+We implemented several strategies to maintain data quality.
+First, we conducted detailed review of each source based on bespoke reports.
+A report is generated for each source, showing all data and metadata, and is checked by both an AusTraits curator and, where possible, the original contributor for accuracy.
+In these reports, observations for each trait are plotted against all other values for the trait in AusTraits.
+This allows AusTraits curators and contributors to quickly visually check the data to ensure it plots against other trait values as expected.
+Corrections suggested by contributors are combined back into AusTraits and made available with the next release.
+
+Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
+Data that does not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use if desired.
+
+Third, we provide a pathway for user feedback.
+As AusTraits is a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
 For instance, we welcome the reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
 
-#### Technical Validation
-
-AusTraits is curated on a voluntary basis. Quality control of data and editorial procedures include:
-
-  - **Automated tests** which confirm that the taxa names provided can be mapped onto accepted names on The Plant List, that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
-Data that does not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use if desired.
-
-  - **Detailed review of each source based on bespoke reports**: A report is generated for each source, showing all data and metadata, and is checked by both an AusTraits curator and, where possible, the original contributor for accuracy.
-In these reports, observations for each trait are plotted against all other values for the trait in AusTraits.
-This allows AusTraits curators and contributors to quickly visually check the data to ensure it plots against other trait values as expected.
-Corrections suggested by contributors are combined back into AusTraits and made available with the next static release.
-
-  - **User feedback**: Data issues can be reported for any observation on the project website.
-
-#### Usage Notes
+## Usage Notes
 
@@
 
-#### Acknowledgement
+Matched against location records.
 
-DSF was supported by Australian Research Council Future Fellowship (FT160100113), RVG was supported by Australian Research Council DECRA Fellowship (DE170100208).
+Phylogenetic tree constructed by pruning  \citep{Smith-2018} using package `V.PhyloMaker` \citep{Jin-2020} and visualising using the package `ggtree` \citep{Yu-2017}
 
-We achknowelde XXXXXXXXXXx institutions, who
+## Code Availability
 
-#### Author contributions
+All code, raw and compiled data are hosted within a GitHub repository under the Trait Ecology and Evolution organisation (see https://github.com/traitecoevo/austraits.build, https://github.com/traitecoevo/austraits).
+Sequential versionscompiled dataset are also being archived within Zenodo \citep{falster_2019}.
+The archived material includes all data sources and code for rebuilding the compiled dataset.
 
-RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, CB, JL, SA error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
-All other authors contributed the raw data and metadata undeprinning the resource and reviewed the harmonised data for errors.
+## Acknowledgements
 
-#### Competing interests
+This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FTXXX), and funding from the Australian Research Data Commons under their "Transformation data collections" program.
+We gratefully acknowledge input from the following persons who contributed to data collection `r acknowledge$name %>% unique() %>% paste( collapse= ", ")`.
+We acknowledge XXXXXXXXXXx institutions, who XXXXX
+
+## Author contributions
+
+RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW,
+SA, CB, JL error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
+All other authors contributed the raw data and metadata underpinning the resource, reviewed the harmonised data for errors, and reviewed the final paper for publication.
+
+## Competing interests
 
@@
 
-```{r austraits}
-## Assumes to items exist in global name space
-austraits <- readRDS("data/austraits_2.0.0.rds")
-
-definitions <- austraits$definitions
-
-tmp <- RefManageR::WriteBib(austraits$sources, "data/references_data.bib")
-
-# check non ASCII file
-#tools::showNonASCIIfile("references.bib")
-
-sources_all <- 
-  bind_rows(
-    austraits$methods %>% select(dataset_id, citation = source_primary_key) %>% 
-            distinct() %>% mutate(type="primary"),
-    austraits$methods %>% select(dataset_id, citation = source_secondary_key) %>% 
-            distinct() %>% mutate(type="secondary")
-  ) %>% 
-  na.omit() %>% 
-  arrange(dataset_id) %>% 
-  group_by(dataset_id) %>% 
-  summarise(refs = sprintf("@%s", citation) %>% 
-  paste(collapse = "; "))
-
-```
-
-### Overview
+## Overview
 
@@
 
-### Data structures
+\clearpage
+
+## Figures & Tables
+
+![The data curation pathway used to assemble the AusTraits database.
+Trait observations are accessed from original data sources, including published floras and field campaigns.
+Features such as variables names, units and taxonomy are harmonised to a common standard.
+Versioned releases are distributed to users, allowing the dataset to be used and re-used in a reproducible way.
+](figures/Pipeline.png)
 
@@
 
-## Figures
-
-![**Figure 1.
-The data curation pathway used to assemble the AusTraits database.** Trait observations are accessed from original data sources, including published floras and field campaigns.
-Factors such as units and taxonomy are harmonised to a common standard.
-Versioned releases are distributed to users, allowing the dataset to be used and re-used in a reproducible way.
-Errors are corrected and new datasets added via the blue path for sequential releases of the database.](figures/Pipeline.png)
-
-![**Figure 2.** (a.) Locations where the trait data were collected and the corresponding number of taxa.
-(b.) AusTraits’ sites (grey) within Australia’s precipitation-temperature space (black) superimposed upon Whittaker’s major biomes diagram.
-Data from WorldClim (worldclim.org).
-(c.) Overview of the number of taxa with which traits are measured within each major plant organ and trait category.](figures/austraits_fig2_blue.png)
-
-![**Figure 3.** The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
-The heatmap colour intensities denote the number of traits measured within each major plant organ.
-The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_phlogenetic_coverage.png)
+![Number of taxa with trait records by plant tissue and trait category, for data that is  (A.) georeferenced and (B.) not georeferenced. ](figures/austraits_fig_composition.pdf)
 
@@
 
-<!-- Cite: -->
-  <!-- G Yu, DK Smith, H Zhu, Y Guan, TTY Lam*.
-ggtree: an R package for visualization and annotation of phylogenetic trees with their covariates and other associated data.
-Methods in Ecology and Evolution.
-2017, 8(1):28-36.
-doi: 10.1111/2041-210X.12628.
--->
-  <!-- G Yu*, TTY Lam, H Zhu, Y Guan*.
-Two methods for mapping and visualizing associated data on phylogeny using ggtree.
-Molecular Biology and Evolution, 2018, 35(2):3041-3043.
-doi: 10.1093/molbev/msy194.
--->
+![Coverage of georeferenced trait records acorss Australian climate space and gerography for traits in different categories.
+(a.) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s major biomes diagram.
+Climate data were extracted from Worldclim (www.worldclim.org).(B.) Locations of georeferenced-sites grouped by plant tissues.
+](figures/austraits_fig_biome_map.pdf)
 
-## Tables
+\clearpage
 
-Table 1:  Main elements of Austraits
+![The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
+The heatmap colour intensity denotes the number of traits measured within family for each plant tissue.
+The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_fig_austraits_phylogenetic_coverage.pdf)
+
+\clearpage
+
+![The number of taxa with trait records in AusTraits and global TRY database (asccessed 28 May 2020). Each point shows a seperate trait. ](figures/austraits_fig_austraits_try_comparison.pdf)
+
+\clearpage
+
+Table 1:  Main elements of the harmonised AusTraits database. See Tables 2-8 for details on each component.
 
@@
   Element = names(austraits), 
-  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description") 
+  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description") %>% gsub(".", "", ., fixed=TRUE)
   ) %>% 
-  kable()
+  knitr::kable("latex",  booktabs = TRUE) %>%
+  kableExtra::kable_styling(latex_options = c("striped")) %>%
+  kableExtra::column_spec(1, width = "15em") %>%
+  kableExtra::column_spec(2, width = "30em") %>% 
+  str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
+   writeLines()
 ```
 
-Table 2: Traits represented in database
-```{r}
-sites <- austraits$sites %>% 
-  filter(site_property %in%  c("longitude (deg)","latitude (deg)")) %>% 
-  spread(site_property, value)
+```{r, results="asis"}
+i <- 2
 
-trait_categories <- read_csv("data/trait_categories.csv", col_types = cols(.default = "c"))
+for(v in names(austraits)[c(1:4, 6:7, 9)]) {
+  elements <- austraits$definitions$austraits$elements[[v]]
 
-data_geo <- 
-  left_join(by=c("dataset_id", "site_name"),
-      select(austraits$traits, -original_name), sites) %>%
-  left_join(by="taxon_name",
-      select(austraits$taxa, taxon_name, family)) %>% 
-  left_join(sources_all, by=c("dataset_id"))
+  if((elements$type == "table" || v == "taxa"))
 
-n_records <- 
-  data_geo %>% 
-  group_by(trait_name) %>%
-  summarise(
-            Description = "",
-            Type= "",
-            `Records (all)` = n(),
-            `Records (geo.)` = sum(!is.na(`longitude (deg)`) & !is.na(`latitude (deg)`)),
-            Studies = n_distinct(dataset_id),
-            Taxon = n_distinct(taxon_name),
-            Families = n_distinct(family),
-            `Data Citations` = refs %>% unique() %>% paste(collapse = "; ")
-            ) %>% 
-  left_join(trait_categories, by=c("trait_name")) %>% 
-  select(Tissue = tissue, Category = category, Trait = trait_name, everything()) %>% 
-  arrange(Tissue, Category, Trait)
+    writeLines("\\clearpage")
 
-n_records$Type <- sapply(austraits$definitions$traits$elements, "[[","type")[n_records$Trait]
-n_records$Description <- sapply(austraits$definitions$traits$elements, "[[","description")[n_records$Trait]
-
-n_records %>%
-  select(-Tissue, Description) %>%
-  kable()
-```
-
-```{r, results="asis", eval=FALSE}
-print_defintions_element <- function(elements) {
-  if(elements$type == "character") {
-    sprintf("**Content:** %s\n", elements$value) %>% 
-      writeLines()
-  }
-
-  if(elements$type == "table") {
-
-    sprintf("**Content:** \n") %>% 
-      writeLines()
+    writeLines(sprintf("Table %s: Structure of the `%s` table, %s \n\n", i, v, gsub("A table", "", elements$description)))
 
@@
     list1_to_df() %>%
-      kable() %>% 
-      writeLines()
-  }
-}
-
-for(v in names(austraits)) {
-  elements <- austraits$definitions$austraits$elements[[v]]
-
-  sprintf("### %s {#%s}\n\n**Description:** %s\n", v, v, elements$description) %>% 
+    knitr::kable("latex",  booktabs = TRUE) %>% 
+    kableExtra::kable_styling(latex_options = c("striped")) %>%
+    kableExtra::column_spec(1, width = "15em") %>%
+    kableExtra::column_spec(2, width = "30em") %>%
+    str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
     writeLines()
 
-  elements %>%
-    print_defintions_element()
-
-  writeLines(c(""))
+    i <- i+1
 }
@@
 
-Value types
+\clearpage
+
+Table 9: Possible value types of trait records.
 
@@
   list1_to_df()  %>%
-  kable() %>% 
+  knitr::kable("latex",  booktabs = TRUE) %>%
+  kableExtra::kable_styling(latex_options = c("striped")) %>%
+  kableExtra::column_spec(1, width = "15em") %>%
+  kableExtra::column_spec(2, width = "30em") %>%
+  str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
   writeLines()
@@
 
-\footnotesize
+\newpage
 
-### Data sources:
+Table 10: Details on all traits represented in Version `r austraits$build_info$version` of AusTraits.
 
-#### Version 1 - via bib.print
+```{r}
 
-```{r, results='asis', eval=TRUE}
-sources_unique <-
-  bind_rows(
-    austraits$methods %>% select(citation = source_primary_citation)  %>% mutate(type="primary"),
-    austraits$methods %>% select(dataset_id, citation = source_secondary_citation) %>% mutate(type="secondary")
+## Add `\citet{}` for references. 
+## Separate into groups of max size n. 
+## This is needed for latex table, overwise citations
+## spill off the page
+format_cites_n <- function(x, n = 6){
+
+  # Make sorted vector of all unique refs
+  x2 <- str_split(x, ", ") %>% unlist() %>% sort()
+  # Make a vector with group ids for references
+  groups <- (rep(1:ceiling(length(x2)/n), n) %>% sort() )[seq_len(length(x2))]
+
+  ## wrap given txt with citation call  
+  format_cites_sub <- function(x) {x %>% paste(collapse = ", ") %>% paste0("\\citep{", ., "}" )}
+
+  ## split into groups of n citations, wrap each with citation call, the combine
+  split(x2, groups) %>% lapply(format_cites_sub) %>% paste(collapse = " \\newline ")
+  }
+
+## format table
+make_latex_table <- function(data) {
+  data %>%
+  # nasty hacks - replace `_` with `TTTTT` then return below, prevents unhelpful handling of _ in creation of latex table
+  mutate(refs = str_replace_all(refs, "_", "TTTT")) %>%  
+  knitr::kable("latex",  booktabs = TRUE, longtable = TRUE ) %>%
+  kableExtra::kable_styling(font_size = 7, latex_options = c("repeat_header", "striped")) %>%
+  kableExtra::column_spec(1, width = "3cm") %>%
+  kableExtra::column_spec(2, width = "4cm") %>%
+  kableExtra::column_spec(9, width = "3cm") %>%
+  kableExtra::add_header_above(c(" ", " ", " ", "Number of records" = 5)) %>%
+  # nasty hacks - clean up backslashes and references. (why oh why is this needed?)
+  str_replace_all(fixed("textbackslash{}"), fixed("")) %>%
+  str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
+  str_replace_all(fixed("\\{"), fixed("{")) %>%
+  str_replace_all(fixed("\\}"), fixed("}"))  %>% 
+  str_replace_all(fixed("TTTT"), fixed("_")) %>%
+  str_split("\\n") %>% unlist()
+}
+
+# wraps text in table cell to max number of characters per line -- needed for long words with no spaces (i.e. trait names)
+txt_wrap <- function(txt, n=22) {
+
+  f <- function(txt, end=-1L) {
+      ifelse(str_length(txt) > 0, paste("\\newline", str_sub(txt, 0, end)), txt)
+  }
+
+  g <- function(txt) {
+      ifelse(str_length(txt) < n, -1L, 
+              str_locate_all(txt,"_") %>% purrr::map_dbl(~.x %>% .[,"start"] %>% subset(., .<= n) %>% max(., -1L)) 
+            )
+  }
+
+  i <- g(txt)
+  txt2 <- ifelse(str_length(txt) > n , str_sub(txt, i+1), "")
+  ii <- g(txt2)
+  txt3 <- ifelse(str_length(txt2) > n , str_sub(txt2, ii+1), "")
+
+  paste(str_sub(txt, 1, i), f(txt2, ii), f(txt3))
+}
+
+n_records %>%
+  mutate(
+         Trait = txt_wrap(Trait),
+         Group = paste0(str_to_sentence(Tissue), " (", Category, ")") %>%  str_replace_all("_", " "), 
+         Type = ifelse(Type=="numeric", "num.", Type),
+         Type = ifelse(Type=="categorical", "cat.", Type),
+         Type = ifelse(Type=="character", "chr.", Type),
+         refs = purrr::map_chr(refs, format_cites_n),
          ) %>%
-  select(citation) %>%
-            distinct() %>% 
-            mutate(source_number=1:n())
+  select(-Tissue, -Category) %>%
+  select(Group, everything()) %>%
+  nest(data = -Group) %>% 
+  mutate(
+    ## turn each group into table
+    tex = purrr::map(data, make_latex_table),
+    ## drop top and bottom of each table
+    tex_sub =  purrr::map2_chr(tex, Group, ~c(sprintf("\\multicolumn{3}{l}{{\\bf %s}}\\\\", .y),.x[ seq(1 + which(grepl("endlastfoot", .x)), -1 + which(grepl("end{longtable}", .x, fixed=TRUE)))], "\\\\ \\midrule \n  \\\\ \n ") %>% paste(collapse = "\n"))
+    ) -> x
 
-sprintf("%s. %s", sources_unique$source_number, sources_unique$citation) %>% paste(collapse = "\n\n") %>% writeLines()
-
+c(x$tex[[1]][1:18], unlist(x$tex_sub), x$tex[[1]][c(-1:0+length(x$tex[[1]]))]) %>%
+  writeLines()
 ```
 
-#### Version 3 - via csl
+\clearpage
 
-```{r, results='asis', eval=FALSE}
-
-austraits$sources %>% sapply(function(i) i$key) -> keys
-
-sprintf("@%s",keys) %>% paste(collapse = "; ") %>% writeLines()
-```
-
-### References
+\footnotesize

### 2020-11-03 R/load_data.R

@@ -1,0 +1,98 @@
+
+load_data <- function() {
+  
+  if(!exists("austraits", envir = .GlobalEnv )) {
+    austraits <- readRDS("data/austraits_2.0.0.rds")
+    assign("austraits", austraits, envir = .GlobalEnv)
+  }
+  
+  assign("trait_category_lookup", read_csv("data/trait_categories.csv", col_types = cols(.default = "c")), envir = .GlobalEnv)
+  
+  aus_trait_by_taxon <- 
+    austraits$traits %>%
+    group_by(trait_name) %>%
+    summarise(n_austraits = n_distinct(taxon_name))
+  
+  # TRY - data for 3820 species, 46,328 trait x species combinations, 1448 locations
+  TRY <- 
+    suppressWarnings(
+    bind_rows(
+      read_tsv("data/2020.05.28-TRY/tde202052811159.txt", skip = 3),
+      read_tsv("data/2020.05.28-TRY/tde2020528113232.txt", skip = 3)
+      )) %>% 
+    distinct() %>%
+    left_join(by= "Trait", read_csv("data/2020.05.28-TRY/try_categories.csv"))
+  
+  
+  try_by_sp <- 
+    TRY %>%
+    group_by(Trait, tissue) %>%
+    summarise(
+      n_try = n_distinct(AccSpeciesName)) %>% 
+    rename("TRY tissue" = "tissue")
+  
+  aus_try <- 
+    trait_category_lookup %>% 
+    select(trait_name, `TRY name`, tissue) %>%
+    full_join(aus_trait_by_taxon, by = "trait_name") %>%
+    full_join(try_by_sp, by = c("TRY name" = "Trait")) %>%
+    mutate_if(is.numeric, ~ replace_na(., 0)) %>%
+    mutate(tissue = ifelse(is.na(trait_name),`TRY tissue`, tissue)) %>%
+    group_by(`TRY name`) %>% 
+    summarise(
+      trait_name = trait_name[1],
+      n_austraits = sum(n_austraits), 
+      n_try = n_try[1],
+      tissue = subset(tissue, !is.na(tissue))[1]
+    ) %>% ungroup()
+  
+  assign("aus_try", aus_try, envir = .GlobalEnv)
+  
+  
+  sources_all <- 
+    bind_rows(
+      austraits$methods %>% select(dataset_id, citation = source_primary_key),
[…]

### 2020-11-03 build_figures.R

@@ -10,7 +10,7 @@ source("scripts/fig_site_locations.R") # A
 source("scripts/fig_climate_space_austraits.R") # B
 
 
-pdf("figures/austraits_fig_2_biome_map.pdf",
+pdf("figures/austraits_fig_biome_map.pdf",
     height = 12,
     width = 8)
 austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A')
@@ -19,7 +19,7 @@ dev.off()
 #### Figure 3: trait composition count ####
 
 source("scripts/fig_trait_composition.R")
-pdf("figures/austraits_fig_3_composition.pdf",
+pdf("figures/austraits_fig_composition.pdf",
     height = 6,
     width = 8)
 austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'A')
@@ -32,7 +32,7 @@ source("scripts/fig_link_tree_and_trait_count_matrix.R") # C
 
 
 pdf(
-  "figures/austraits_fig_4_austraits_phlogenetic_coverage.pdf",
+  "figures/austraits_fig_austraits_phylogenetic_coverage.pdf",
   height = 12,
   width = 12
 )
@@ -43,18 +43,9 @@ dev.off()
 #### Figure 5: AusTraits and TRY comparison ####
 source("scripts/fig_comparison_TRY.R") 
  
-# aus_try %>%
-#   filter(n_try > 1, n_austraits >1, tissue=="leaf") %>%
-#   arrange(desc(n_try/n_austraits)) %>% View
-
-pdf( "figures/austraits_fig_5_austraits_try_comparison_v1.pdf",
-     height = 4,
+pdf( "figures/austraits_fig_austraits_try_comparison.pdf",
+     height = 6,
      width = 5)
 p_aus_try
 dev.off()
 
-pdf("figures/austraits_fig_5_austraits_try_comparison_v3.pdf",
-    height = 8,
-    width = 10)
-p_aus_try + facet_wrap( ~tissue)
-dev.off()

### 2020-11-03 scripts/fig_comparison_TRY.R

@@ -1,39 +1,3 @@
-library(tidyverse)
-
-#==== 01 Load data ====
-# TRY - data for 3820 species, 46,328 trait x species combinations, 1448 locations
-TRY <- bind_rows(
-  read_tsv("data/2020.05.28-TRY/tde202052811159.txt", skip = 3),
-  read_tsv("data/2020.05.28-TRY/tde2020528113232.txt", skip = 3)) %>% 
-  distinct()
-
-# AUSTRAITS -- data over 650000 records,
-# for 339 traits, for 23,480 species. 282000 trait x species combinations, 22,000 locations
-austraits <- readRDS("data/austraits_2.0.0.rds")
-tissue_lookup <- read_csv("data/trait_categories.csv") %>%
-  select(trait_name, `TRY name`, tissue)
-
-
-#==== 02 Data wrangling ====
-aus_trait_by_taxon <- austraits$traits %>%
-  group_by(trait_name) %>%
-  summarise(n_austraits = n_distinct(taxon_name))
-
-try_by_sp <- TRY %>%
-  group_by(Trait) %>%
-  summarise(n_try = n_distinct(AccSpeciesName))
-
-aus_try <- tissue_lookup %>%
-  full_join(aus_trait_by_taxon, by = "trait_name") %>%
-  full_join(try_by_sp, by = c("TRY name" = "Trait")) %>%
-  mutate_if(is.numeric, ~ replace_na(., 0)) %>%
-  mutate(tissue = replace_na(tissue, "uncategorised")) %>% 
-  group_by(`TRY name`) %>% 
-  summarise(
-    n_austraits = sum(n_austraits), 
-    n_try = n_try[1],
-    tissue = subset(tissue, !is.na(tissue))[1]
-  ) %>% ungroup()
 
 #==== 03 Plotting ====
 # Scatter plot
@@ -41,14 +5,24 @@ aus_try <- tissue_lookup %>%
 #  y = number of records of trait x in TRY
 # color by tissue
 p_aus_try <-
-  ggplot(data = aus_try, aes(x = n_austraits, y = n_try, color = tissue)) +
-  geom_jitter(alpha = 0.5, size = 2) +
-  geom_abline(alpha = 0.1) +
-  theme_minimal() +
-  scale_x_log10() +
-  scale_y_log10() +
-  xlab("Taxon in Austraits") +
-  ylab("Species in TRY") +
+  aus_try %>% 
[…]

### 2020-11-03 scripts/fig_trait_composition.R

@@ -4,7 +4,8 @@ positions <-
   c("leaf", "stem", "bark", "root", "reproductive", "whole_plant")
 
 # Georeferenced trait data
-austraits_composition_georef <- trait_composition_list %>%
+austraits_composition_georef <- 
+  trait_composition_list %>%
   filter(!is.na(`latitude (deg)`),!is.na(`longitude (deg)`)) %>%
   group_by(tissue, category) %>%
   summarise(n = n_distinct(taxon_name)) %>%
@@ -16,6 +17,7 @@ austraits_composition_georef <- trait_composition_list %>%
   scale_y_log10(
     breaks = trans_breaks("log10", function(x)
       10 ^ x),
+    limits = c(1, 20000),
     labels = trans_format("log10", math_format(10 ^ .x))
   ) +
   xlab("Major plant part") + ylab("Species (n)") +
@@ -26,7 +28,8 @@ austraits_composition_georef <- trait_composition_list %>%
   ggtitle("Georeferenced")
 
 # Non-georeferenced trait data
-austraits_composition_non_georef <- trait_composition_list %>%
+austraits_composition_non_georef <- 
+  trait_composition_list %>%
   filter(is.na(`latitude (deg)`), is.na(`longitude (deg)`)) %>%
   group_by(tissue, category) %>%
   summarise(n = n_distinct(taxon_name)) %>%
@@ -38,11 +41,12 @@ austraits_composition_non_georef <- trait_composition_list %>%
   scale_y_log10(
     breaks = trans_breaks("log10", function(x)
       10 ^ x),
+    limits = c(1, 20000),
     labels = trans_format("log10", math_format(10 ^ .x))
   ) +
   xlab("Major plant part") + ylab("Species (n)") +
   scale_x_discrete(limits = positions) +
   theme(legend.position = "right") +
-  scale_fill_viridis_d(option = "plasma") +
+  scale_fill_viridis_d() +
   annotation_logticks(sides = "l") +
   ggtitle("Non-Georeferenced") 

### 2020-11-03 ms.Rmd
Consolidate dataset constructionm package loading

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
-library(knitr)
-library(tidyverse)
-library(rmarkdown)
+devtools::load_all()
 source("R/ms.R")
@@
 ```{r}
-
-## Add `\citet{}` for references. 
-## Separate into groups of max size n. 
-## This is needed for latex table, overwise citations
-## spill off the page
-format_cites_n <- function(x, n = 6){
-
-  # Make sorted vector of all unique refs
-  x2 <- str_split(x, ", ") %>% unlist() %>% sort()
-  # Make a vector with group ids for references
-  groups <- (rep(1:ceiling(length(x2)/n), n) %>% sort() )[seq_len(length(x2))]
-
-  ## wrap given txt with citation call  
-  format_cites_sub <- function(x) {x %>% paste(collapse = ", ") %>% paste0("\\citep{", ., "}" )}
-
-  ## split into groups of n citations, wrap each with citation call, the combine
-  split(x2, groups) %>% lapply(format_cites_sub) %>% paste(collapse = " \\newline ")
-  }
-
-## format table
-make_latex_table <- function(data) {
-  data %>%
-  # nasty hacks - replace `_` with `TTTTT` then return below, prevents unhelpful handling of _ in creation of latex table
-  mutate(refs = str_replace_all(refs, "_", "TTTT")) %>%  
-  knitr::kable("latex",  booktabs = TRUE, longtable = TRUE ) %>%
-  kableExtra::kable_styling(font_size = 7, latex_options = c("repeat_header", "striped")) %>%
-  kableExtra::column_spec(1, width = "3cm") %>%
-  kableExtra::column_spec(2, width = "4cm") %>%
-  kableExtra::column_spec(9, width = "3cm") %>%
-  kableExtra::add_header_above(c(" ", " ", " ", "Number of records" = 5)) %>%
-  # nasty hacks - clean up backslashes and references. (why oh why is this needed?)
-  str_replace_all(fixed("textbackslash{}"), fixed("")) %>%
-  str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
-  str_replace_all(fixed("\\{"), fixed("{")) %>%
-  str_replace_all(fixed("\\}"), fixed("}"))  %>% 
-  str_replace_all(fixed("TTTT"), fixed("_")) %>%
-  str_split("\\n") %>% unlist()
-}
-
-# wraps text in table cell to max number of characters per line -- needed for long words with no spaces (i.e. trait names)
-txt_wrap <- function(txt, n=22) {
-
-  f <- function(txt, end=-1L) {
-      ifelse(str_length(txt) > 0, paste("\\newline", str_sub(txt, 0, end)), txt)
-  }
-
-  g <- function(txt) {
-      ifelse(str_length(txt) < n, -1L, 
-              str_locate_all(txt,"_") %>% purrr::map_dbl(~.x %>% .[,"start"] %>% subset(., .<= n) %>% max(., -1L)) 
-            )
-  }
-
-  i <- g(txt)
-  txt2 <- ifelse(str_length(txt) > n , str_sub(txt, i+1), "")
-  ii <- g(txt2)
-  txt3 <- ifelse(str_length(txt2) > n , str_sub(txt2, ii+1), "")
-
-  paste(str_sub(txt, 1, i), f(txt2, ii), f(txt3))
-}
-
  n_records %>%

### 2020-11-03 R/get_sites.R

@@ -1,56 +1,0 @@
-
-# collect sites
-austraits$sites %>%
-  as_tibble() %>%
-  filter(site_property %in% c("latitude (deg)", "longitude (deg)")) %>%
-  pivot_wider(names_from = site_property, values_from = value) %>%
-  mutate(
-    `latitude (deg)` = as.numeric(`latitude (deg)`),
-    `longitude (deg)` = as.numeric(`longitude (deg)`)
-  ) -> sites
-
-# calculate species number
-austraits$traits %>%
-  as_tibble() %>%
-  dplyr::select(dataset_id, taxon_name, site_name) %>%
-  dplyr::group_by(dataset_id, site_name) %>%
-  distinct(taxon_name) %>%
-  count(dataset_id, sort = TRUE) %>%
-  ungroup() -> sp_number
-
-# get trait categories
-trait_category_lookup <- read_csv("data/trait_categories.csv") %>%
-  dplyr::select(trait_name, tissue, category)
-
-# get trait list
-trait_list <- austraits$traits %>%
-  as_tibble() %>%
-  dplyr::select(dataset_id, taxon_name, site_name, trait_name)
-
-trait_composition_list <- trait_list %>%
-  full_join(trait_category_lookup, by = "trait_name") %>%
-  full_join(sites, by = c("dataset_id", "site_name"))
-
-# uncategorised/ unclassified traits
-unclassified_traits <- austraits$traits %>%
-  dplyr::select(dataset_id, site_name, trait_name) %>%
-  left_join(trait_category_lookup, by = "trait_name") %>%
-  dplyr::filter(is.na(category)) %>%
-  dplyr::select(trait_name, tissue, category) %>%
-  distinct()
-
-# extract trait categories by dataset_id and site_name
-trait_categories <- austraits$traits %>%
-  as_tibble()  %>%
-  dplyr::select(dataset_id, site_name, trait_name) %>%
-  left_join(trait_category_lookup, by = "trait_name")  %>%
-  mutate(category = str_replace(category, "physiology", "physiological")) %>%
-  dplyr::select(-trait_name) %>%
-  distinct()
-
-sites %<>%
-  left_join(sp_number, by = c("dataset_id", "site_name")) %>%
-  left_join(trait_categories, by = c("dataset_id", "site_name"))
-
-
-rm(sp_number, trait_categories)

### 2020-11-03 R/load_data.R

@@ -6,49 +6,11 @@ load_data <- function() {
     assign("austraits", austraits, envir = .GlobalEnv)
   }
   
+  ## Trait lookup table
+
   assign("trait_category_lookup", read_csv("data/trait_categories.csv", col_types = cols(.default = "c")), envir = .GlobalEnv)
   
-  aus_trait_by_taxon <- 
-    austraits$traits %>%
-    group_by(trait_name) %>%
-    summarise(n_austraits = n_distinct(taxon_name))
-  
-  # TRY - data for 3820 species, 46,328 trait x species combinations, 1448 locations
-  TRY <- 
-    suppressWarnings(
-    bind_rows(
-      read_tsv("data/2020.05.28-TRY/tde202052811159.txt", skip = 3),
-      read_tsv("data/2020.05.28-TRY/tde2020528113232.txt", skip = 3)
-      )) %>% 
-    distinct() %>%
-    left_join(by= "Trait", read_csv("data/2020.05.28-TRY/try_categories.csv"))
-  
-  
-  try_by_sp <- 
-    TRY %>%
-    group_by(Trait, tissue) %>%
-    summarise(
-      n_try = n_distinct(AccSpeciesName)) %>% 
-    rename("TRY tissue" = "tissue")
-  
-  aus_try <- 
-    trait_category_lookup %>% 
-    select(trait_name, `TRY name`, tissue) %>%
-    full_join(aus_trait_by_taxon, by = "trait_name") %>%
-    full_join(try_by_sp, by = c("TRY name" = "Trait")) %>%
-    mutate_if(is.numeric, ~ replace_na(., 0)) %>%
-    mutate(tissue = ifelse(is.na(trait_name),`TRY tissue`, tissue)) %>%
-    group_by(`TRY name`) %>% 
-    summarise(
-      trait_name = trait_name[1],
-      n_austraits = sum(n_austraits), 
-      n_try = n_try[1],
-      tissue = subset(tissue, !is.na(tissue))[1]
-    ) %>% ungroup()
-  
-  assign("aus_try", aus_try, envir = .GlobalEnv)
-  
-  
+  ### 
   sources_all <- 
     bind_rows(
       austraits$methods %>% select(dataset_id, citation = source_primary_key),
@@ -85,14 +47,95 @@ load_data <- function() {
       families = n_distinct(family),
       refs =  refs %>% unique() %>% paste(collapse = ", ")
     ) %>% 
-    left_join(trait_categories %>% select(trait_name, tissue, category), by=c("trait_name")) %>% 
[…]

### 2020-11-03 R/ms.R

@@ -9,3 +9,65 @@ list1_to_df <- function(my_list) {
   tibble(key = names(my_list), value = unlist(my_list))
 }
 
+
+
+## Add `\citet{}` for references. 
+## Separate into groups of max size n. 
+## This is needed for latex table, overwise citations
+## spill off the page
+format_cites_n <- function(x, n = 6){
+  
+  # Make sorted vector of all unique refs
+  x2 <- str_split(x, ", ") %>% unlist() %>% sort()
+  # Make a vector with group ids for references
+  groups <- (rep(1:ceiling(length(x2)/n), n) %>% sort() )[seq_len(length(x2))]
+  
+  ## wrap given txt with citation call  
+  format_cites_sub <- function(x) {x %>% paste(collapse = ", ") %>% paste0("\\citep{", ., "}" )}
+  
+  ## split into groups of n citations, wrap each with citation call, the combine
+  split(x2, groups) %>% lapply(format_cites_sub) %>% paste(collapse = " \\newline ")
+}
+
+
+## format table
+make_latex_table <- function(data) {
+  data %>%
+    # nasty hacks - replace `_` with `TTTTT` then return below, prevents unhelpful handling of _ in creation of latex table
+    mutate(refs = str_replace_all(refs, "_", "TTTT")) %>%  
+    knitr::kable("latex",  booktabs = TRUE, longtable = TRUE ) %>%
+    kableExtra::kable_styling(font_size = 7, latex_options = c("repeat_header", "striped")) %>%
+    kableExtra::column_spec(1, width = "3cm") %>%
+    kableExtra::column_spec(2, width = "4cm") %>%
+    kableExtra::column_spec(9, width = "3cm") %>%
+    kableExtra::add_header_above(c(" ", " ", " ", "Number of records" = 5)) %>%
+    # nasty hacks - clean up backslashes and references. (why oh why is this needed?)
+    str_replace_all(fixed("textbackslash{}"), fixed("")) %>%
+    str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
+    str_replace_all(fixed("\\{"), fixed("{")) %>%
+    str_replace_all(fixed("\\}"), fixed("}"))  %>% 
+    str_replace_all(fixed("TTTT"), fixed("_")) %>%
+    str_split("\\n") %>% unlist()
+}
+
[…]

### 2020-11-03 build_figures.R

@@ -2,13 +2,14 @@
 
 devtools::load_all()
 
+load_data()
+
 source("R/plotting_functions.R") # Load required packages and functions
 
 
 #### Figure 2: geographical and biome map####
 source("scripts/fig_site_locations.R") # A
 source("scripts/fig_climate_space_austraits.R") # B
-
 
 pdf("figures/austraits_fig_biome_map.pdf",
     height = 12,

### 2020-11-03 scripts/fig_comparison_TRY.R

@@ -5,7 +5,7 @@
 #  y = number of records of trait x in TRY
 # color by tissue
 p_aus_try <-
-  aus_try %>% 
+  aus_try_comparison %>% 
   mutate_at(c("n_austraits", "n_try"), ~.x/28000*100) %>% 
   ggplot(aes(y = n_austraits, x = n_try, color = tissue)) +
   geom_point(alpha = 0.5, size = 2) +

### 2020-11-03 scripts/fig_generate_phylogenies.R

@@ -1,11 +1,11 @@
 ## 1 Build a phylogenetic tree from the list of species in AusTraits
 
-##### 1 Read data #####
-## add austraits and read the lookup table
-austraits <- readRDS("data/austraits_2.0.0.rds")
-#tree<-readRDS("data/AusTraits_phylogenies.rds") 
-trait_category_lookup<-read_csv("data/trait_categories.csv") %>% 
-  dplyr::select(trait_name, tissue, category)
+# ##### 1 Read data #####
+# ## add austraits and read the lookup table
+# austraits <- readRDS("data/austraits_2.0.0.rds")
+# #tree<-readRDS("data/AusTraits_phylogenies.rds") 
+# trait_category_lookup<-read_csv("data/trait_categories.csv") %>% 
+#   dplyr::select(trait_name, tissue, category)
 
 
 ##### 2 Build a phylogenetic tree ####
@@ -58,4 +58,3 @@ colnames(austraits_summary)<-c( "Root", # Root Traits
                                     "Leaf" # Leaf Traits
                                    ) 
 
-rm(austraits)

### 2020-11-03 scripts/fig_site_locations.R

@@ -5,8 +5,6 @@
 
 # one strategy for retrieving data available on web
 filename <- "data/australia.tif"
-if (!file.exists(filename))
-  download.file("htpp:....", filename)
 
 # Read bioclim data using raster package
 au_map <- raster::raster(filename) %>%
@@ -24,12 +22,11 @@ ggplot() +
   myTheme -> au_basemap
 
 
-
 #### 03 Overlay site locations  ####
-#sites <- read_csv("data/austraits_sites.csv")
-source("R/get_sites.R")
-## Filter sites
-sites2 <- sites %>%
+
+austraits_site_locations_by_tissue_fig_2 <- 
+  sites %>%
+  select(site_name, `latitude (deg)`, `longitude (deg)`, `tissue`) %>%
   filter(
     `latitude (deg)` > (-45),
     `latitude (deg)` < (-9.5),
@@ -43,27 +40,14 @@ sites2 <- sites %>%
       "site_at_-16.9333_degS_and_149.1833_degE",
       "site_at_-16.9833_degS_and_149.8833_degE"
     )
-  ) 
-
-
-
-#### 03 Overlay site locations  ####
-# austraits_site_locations <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = NA )
-# austraits_site_locations_by_tissue <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "tissue" )
-# austraits_site_locations_by_category <- produce_site_map(sites2, "latitude (deg)", "longitude (deg)", feature = "category" )
-
-austraits_site_locations_by_tissue_fig_2 <- sites2 %>%
+  ) %>%
   drop_na() %>%
   produce_site_map("latitude (deg)", "longitude (deg)", feature = "tissue") +
   theme(
-    #legend.justification = c(-.05, -.05),
     legend.position = "bottom",
-    # legend.direction  = "horizontal",
     strip.background = element_blank(),
     strip.text.x = element_text(size = 12),
     legend.key.height = unit(0.5, "cm"),
     legend.key.width = unit(2, "cm"),
     legend.box = "horizontal" #,
-    # legend.spacing  = unit(0.1, "cm")
-    # legend.text = element_text(size=2)
   )

### 2020-11-03 ms.Rmd
spelling, grammar, usage

@@
 Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives \citep[but see][]{guerin2017opportunities}.
-We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to accelerate our collective understanding the Australian flora.
+We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to accelerate our collective understanding of the Australian flora.
 Open access to a comprehensive resource like this will generate significant new knowledge about the Australian flora across multiple scales of interest, as well as reduce duplication of effort in the compilation of plant trait data, particularly for research students and government agencies seeking to access information on traits.
@@
 The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection site (e.g.
-location co-ordinates, soil conditions, number of replicates) and information about the methods used to derive measurements (e.g.
+location coordinates, soil conditions, number of replicates) and information about the methods used to derive measurements (e.g.
 number of replicates, equipment used where available).
 The finished database contains 11 elements, as described in Table 1.
-This format was developed to include information about the trait measurements, taxa sampled, the methods used, sites, contextual information, the peopled involved, and citation sources.
+This format was developed to include information about the trait measurements, taxa sampled, the methods used, sites, contextual information, the people involved, and citation sources.
 
@@
 
+### Access
+
+As an evolving data product, successive versions of AusTraits are being released, containing updates and corrections.
+Versions are labeled using semantic versioning to indicate the change between versions \citep{Falster-2019}.
 Static versions of the AusTraits, including version`r austraits$build_info$version` used in this descriptor, are available on the project website (https://github.com/traitecoevo/austraits) and Zenodo (Data Citation 1).
@@
 
-Comparing coverage in AusTraits to the global database TRY, there were `r aus_try %>% filter(n_austraits >0, n_try >0)%>% nrow()` traits overlapping.
+Comparing coverage in AusTraits to the global database TRY, there were `r aus_try_comparison %>% filter(n_austraits >0, n_try >0)%>% nrow()` traits overlapping.
 Of these, AusTraits tended to contain records for more taxa, but not always (Figure 5).
 Multiple traits had more than 10 times the number of taxa represented in AusTraits.
-However, there were more records in TRY for `r aus_try %>% filter(n_try > n_austraits, n_austraits >0) %>% nrow()` traits, in particular quite physiological leaf traits.
+However, there were more records in TRY for `r aus_try_comparison %>% filter(n_try > n_austraits, n_austraits >0) %>% nrow()` traits, in particular quite physiological leaf traits.
 Many traits were not overlapping between the two databases (Figure 5).
-We noted that AusTraits includes more seed and fruit nutrient data; possibly reflecting the interest in Australia in understanding how fruit and seeds are provisioned in nutrient depauperate environments.
+We noted that AusTraits includes more seed and fruit nutrient data; possibly reflecting the interest in Australia in understanding how fruit and seeds are provisioned in nutrient-depauperate environments.
 AusTraits includes more categorical values, especially variables documenting different components of species’ fire response strategies, reflecting the importance of fire in shaping Australian communities and the research to document different strategies species have evolved to succeed in fire-prone environments.
@@
 
-We implemented several strategies to maintain data quality.
-First, we conducted detailed review of each source based on bespoke reports.
-A report is generated for each source, showing all data and metadata, and is checked by both an AusTraits curator and, where possible, the original contributor for accuracy.
-In these reports, observations for each trait are plotted against all other values for the trait in AusTraits.
-This allows AusTraits curators and contributors to quickly visually check the data to ensure it plots against other trait values as expected.
+We implemented three strategies to maintain data quality.
+First, we conducted a detailed review of each source based on bespoke report, showing all data and metadata, by both an AusTraits curator and the original contributor (where possible).
+Observations for each trait are plotted against all other values for the trait in AusTraits, allowing quick identification of outliers.
 Corrections suggested by contributors are combined back into AusTraits and made available with the next release.
@@
 
-Each data release is a compressed folder containing separate files for each of the major components of AusTraits (Table 2). Detailed examples for analysis are available on the project website.
+Each data release is available in multiple formats.
+First, as a compressed folder containing text files for each of the main components.
+Second, as a compressed R object, enabling easy loading into R for those using that platform.
 
-Matched against location records.
+Data is released under a CC-BY license enabling reuse with attribution -- being citation of this descriptor and, where possible, original sources.
 
-Phylogenetic tree constructed by pruning  \citep{Smith-2018} using package `V.PhyloMaker` \citep{Jin-2020} and visualising using the package `ggtree` \citep{Yu-2017}
+Using the taxon names aligned with the APC, data can be eaily queried against location data from the Atlas of Living Australia.
+To create the phylogenetic tree in Figure 5, we pruned a master tree for all higher plants \citep{Smith-2018} using the package `V.PhyloMaker` \citep{Jin-2020} and visualising via `ggtree` \citep{Yu-2017}.
 
@@
 
-All code, raw and compiled data are hosted within a GitHub repository under the Trait Ecology and Evolution organisation (see https://github.com/traitecoevo/austraits.build, https://github.com/traitecoevo/austraits).
-Sequential versionscompiled dataset are also being archived within Zenodo \citep{falster_2019}.
+All code, raw and compiled data are hosted within GitHub repositories under the Trait Ecology and Evolution organisation (see https://github.com/traitecoevo/austraits.build, https://github.com/traitecoevo/austraits).
 The archived material includes all data sources and code for rebuilding the compiled dataset.
+The code used to produce this paper is available at http://github.com/traitecoevo/austraits_ms.
 
@@
 ![The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
-The heatmap colour intensity denotes the number of traits measured within family for each plant tissue.
+The heatmap colour intensity denotes the number of traits measured within a family for each plant tissue.
 The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_fig_austraits_phylogenetic_coverage.pdf)
@@
 
-![The number of taxa with trait records in AusTraits and global TRY database (asccessed 28 May 2020). Each point shows a seperate trait. ](figures/austraits_fig_austraits_try_comparison.pdf)
+![The number of taxa with trait records in AusTraits and global TRY database (accessed 28 May 2020). Each point shows a separate trait. ](figures/austraits_fig_austraits_try_comparison.pdf)
 

### 2020-11-09 ms.Rmd
Feedback from coauthors throughout

@@
     keep_md: true
-biblio-style: nature
+biblio-style: naturemag
 header-includes:
@@
 devtools::load_all()
-source("R/ms.R")
-source("R/load_data.R")
 load_data()
@@
 properties.
-This data descriptor provides information on Version `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxa combinations.
+This data descriptor provides information on version  `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxa combinations.
 We envision AusTraits as an on-
@@
 One such region is Australia; a continent with a flora of c.
-28,000 native higher-plant species \citep{council2010australian}.
-While significant investment has been made in curating and digitising herbarium collections in Australia over the last two decades (e.g.
+26,000 native higher-plant species \citep{council2010australian}.
+While significant investment has been made in curating and digitising herbarium collections and observation records in Australia over the last two decades (e.g.
 The Australian Virtual Herbarium houses ~7 million specimen occurrence records; https://avh.ala.org.au), no complementary resource yet exists for consolidating information on plant traits.
-Moreover, relatively few species are represented in the leading global databases.
-For example, the TRY international database has only 3830 species represented in total across all traits measured.
-Such low coverage greatly limits the use of traits in understanding and managing our vegetation.
-While initiatives, such as the TRY database \citep{Kattge-2020} and Open Traits Network \citep{Gallagher-2020}, have been working towards sharing of data about plant traits, though most Australian plant taxa remain unrepresented.
+Moreover, relatively few Australian species are represented in the leading global databases.
+For example, the international TRY database \citep{Kattge-2020} has observations for only 3830 Australian species across all collated traits.
+This level of species coverage limits our ability to use traits to understand and ultimately manage Australian vegetation \citep{Kissling_2018}.
+While initiatives such as TRY \citep{Kattge-2020} and the Open Traits Network \citep{Gallagher-2020} are working towards global synthesis of trait data a stronger representation of Australian plant taxa in these efforts is essential given the high richness and endemicity of this continental flora.
 
 Here we introduce the AusTraits database (hereafter simply AusTraits),  a compilation of plant traits for the Australian flora.
-Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources -– including published scientific papers, floras, and taxonomic descriptions –- and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
+Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
 To assemble AusTraits from diverse primary sources and make data available for reuse, we needed to overcome three main types of challenges (Figure 1): 1) Accessing data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) Harmonising these diverse sources into a federated resource, with common units, trait names, and data formats; and 3) Distributing versions of the data under suitable license.
-To meet this challenge, we developed a workflow for harmonising and distributing data, drawing from emerging community standards and our collective experience building traits databases.
+To meet this challenge, we developed a workflow which draws on emerging community standards and our collective experience building trait databases.
 
 By providing a harmonised and curated dataset on `r austraits$traits$trait_name %>% n_distinct()` plant traits, AusTraits contributes substantially to filling the gap in Australian and global biodiversity resources.
-Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives \citep[but see][]{guerin2017opportunities}.
-We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data to accelerate our collective understanding of the Australian flora.
+Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives.
+We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data about the Australian flora.
 Open access to a comprehensive resource like this will generate significant new knowledge about the Australian flora across multiple scales of interest, as well as reduce duplication of effort in the compilation of plant trait data, particularly for research students and government agencies seeking to access information on traits.
@@
 
-AusTraits Version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments (see Online Table 1).
-Initially we identified a list of candidate traits of interest then identified primary sources containing measurements for these trait, before contacting authors for access.
-As the compilation grew, we expanded the list of traits considered to include any measureable quantity that had been quantified for a moderate number of taxa ( n > 20).
+AusTraits version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments (see Online Table 1).
+Initially we identified a list of candidate traits of interest then identified primary sources containing measurements for these traits, before contacting authors for access.
+As the compilation grew, we expanded the list of traits considered to include any measurable quantity that had been quantified for a moderate number of taxa (n > 20).
 
@@
 This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics \citep{Garnier-2017}.
-We categorised traits based on the tissue where it is measured (`r trait_categories$tissue %>% unique() %>% sort() %>% paste(collapse=", ")`) and the type of measurement (`r trait_categories$category %>% unique() %>% sort() %>% paste(collapse=", ")`).
-Version `r austraits$build_info$version` of AusTraits includes  `r n_records$Type %>% nrow` of type: `r n_records$Type %>% table() %>% paste(., names(.))`.
+We categorised traits based on the tissue where it is measured (`r trait_categories$tissue %>% unique() %>% sort() %>% paste(collapse=", ")`) and the type of measurement (`r trait_categories$category %>% unique() %>% sort() %>% gsub("_", " ", .) %>% paste(collapse=", ")`).
+Version `r austraits$build_info$version` of AusTraits includes  `r n_records$Type %>% table() %>% sort(decreasing = T) %>% paste(c( "", "", "and"), .,  names(.), collapse= ", ")` traits.
 
@@
 
-The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection site (e.g.
-location coordinates, soil conditions, number of replicates) and information about the methods used to derive measurements (e.g.
-number of replicates, equipment used where available).
-The finished database contains 11 elements, as described in Table 1.
+The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection (e.g.
+location coordinates, light levels s) and information about the methods used to derive measurements (e.g.
+number of replicates, equipment used).
+The database contains 11 elements, as described in Table 1.
 This format was developed to include information about the trait measurements, taxa sampled, the methods used, sites, contextual information, the people involved, and citation sources.
 
-For storage efficiency, the main traits table contains relatively little information (Table 2), but can be cross linked against other tables (Tables 3-8) using several identifiers for dataset, site, context, observation and taxon (Table 1).
-The `dataset_id` is ordinarily the surname of first author and year of publication, for the source’s primary citation (e.g.
+For storage efficiency, the main table of traits contains relatively little information (Table 2), but can be cross linked against other tables (Tables 3-8) using identifiers for dataset, site, context, observation and taxon (Table 1).
+The `dataset_id` is ordinarily the surname of the first author and year of publication associated with the source’s primary citation (e.g.
 `Blackman_2014`).
-Trait measurement were also recorded as being one of several possible `value_type` (Table 9), reflecting the type of measurement recorded.
+Trait measurements were also recorded as being one of several possible `value_type` (Table 9), reflecting the type of measurement recorded.
 
@@
 If a name was recorded as being both an accepted name and an alternative (e.g.
-synonym) we preferred the accepted name, but also noting the alternative records.
+synonym) we preferred the accepted name, but also noted the alternative records.
 When a suitable match could not be found, we manually reviewed near matches and web portals such as the Atlas of Living Australia to find a suitable match.
-The final resource reports both the original and the updated taxon name alongside trait records (Table 2), as well as additional tables summarising all taxonomic names changes (Table 6) and further information from the APC and APNI on all taxa included (Table 7).
+The final resource reports both the original and the updated taxon name alongside each trait record (Table 2), as well an additional table summarising all taxonomic names changes (Table 6) and further information from the APC and APNI on all taxa included (Table 7).
 
@@
 
-The number of vascular plants (species, subspecies, varieties and forma) in Australia is ~28000 \citep{apc-2020}.
-Version`r austraits$build_info$version` of AusTraits included at least one record for `r austraits$taxa$taxon_name %>% n_distinct()`, or about 80% of taxa.
+The number of vascular plants (species, subspecies, varieties and forma) in Australia is  ca.
+26,000 \citep{apc-2020}.
+Version `r austraits$build_info$version` of AusTraits included at least one record for `r austraits$taxa$taxon_name %>% n_distinct()`, or about 80% of taxa.
 Five traits (leaf_length, leaf_width, plant_height, life_history, plant_growth_form) have records for more than 50% of taxa.
@@
 
-There were substantial differences in coverage among different tissues and trait types, also with respect to number of geo-referenced points (Figure 2).The most common traits are not geo-referenced recorded from floras.
+There were substantial differences in coverage among different tissues and trait types, also with respect to number of georeferenced points (Figure 2).The most common traits are nongeoreferenced records from floras.
 Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 2a).
@@
 We found that trait records were spread across the climate space of Australia (Figure 3a), as well as geographic locations (Figure 3b).
-As with most data, in Australia, the density of records was somewhat concentrated around  cities, but particularly for leaf traits.
+As with most data, in Australia, the density of records was somewhat concentrated around cities or roads in remote regions, particularly for leaf traits.
 
-Figure 4 shows that overall coverage across the phylogenetic tree is relatively unbiased, though there are some notable exceptions.
-One exception is for root traits, where the taxa within the family Poaceae have considerably more information available than other taxa.
-A cluster of taxa within the family Myrtaceceae have little leaf information available, while reproductive information is limited for species near the based ot the tree.
+Figure 4 shows that overall coverage across a phylogenetic tree of Australian plant species is relatively unbiased, though there are some notable exceptions.
+One exception is for root traits, where taxa within Poaceae have large amounts of information available relative to other plant families.
+A cluster of taxa within the family Myrtaceceae have little leaf information available, while reproductive information is limited for species near the base of the tree.
 
@@
 Multiple traits had more than 10 times the number of taxa represented in AusTraits.
-However, there were more records in TRY for `r aus_try_comparison %>% filter(n_try > n_austraits, n_austraits >0) %>% nrow()` traits, in particular quite physiological leaf traits.
+However, there were more records in TRY for `r aus_try_comparison %>% filter(n_try > n_austraits, n_austraits >0) %>% nrow()` traits, in particular physiological leaf traits.
 Many traits were not overlapping between the two databases (Figure 5).
@@
 We implemented three strategies to maintain data quality.
-First, we conducted a detailed review of each source based on bespoke report, showing all data and metadata, by both an AusTraits curator and the original contributor (where possible).
-Observations for each trait are plotted against all other values for the trait in AusTraits, allowing quick identification of outliers.
-Corrections suggested by contributors are combined back into AusTraits and made available with the next release.
+First, we conducted a detailed review of each source based on a bespoke report, showing all data and metadata, by both an AusTraits curator and the original contributor (where possible).
+Observations for each trait were plotted against all other values for the trait in AusTraits, allowing quick identification of outliers.
+Corrections suggested by contributors were combined back into AusTraits and made available with the next release.
 
 Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
-Data that does not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use if desired.
+Data that did not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use and reviewed.
 
-Third, we provide a pathway for user feedback.
+Third, are providing a pathway for user feedback.
 As AusTraits is a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
-For instance, we welcome the reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
+As such, we welcome reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
 
@@
 
-Data is released under a CC-BY license enabling reuse with attribution -- being citation of this descriptor and, where possible, original sources.
+Data is released under a CC-BY license enabling reuse with attribution -- being a citation of this descriptor and, where possible, original sources.
 
-Using the taxon names aligned with the APC, data can be eaily queried against location data from the Atlas of Living Australia.
+Using the taxon names aligned with the APC, data can be queried against location data from the Atlas of Living Australia.
 To create the phylogenetic tree in Figure 5, we pruned a master tree for all higher plants \citep{Smith-2018} using the package `V.PhyloMaker` \citep{Jin-2020} and visualising via `ggtree` \citep{Yu-2017}.
+To create Figure 3A, we visualised the Whittaker's biome using the package `plotbiomes` \citep{Stefan-2020}.
 
@@
 
-This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FTXXX), and funding from the Australian Research Data Commons under their "Transformation data collections" program.
+This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910) and funding from Macquarie University to Gallagher, and the Australian Research Data Commons via their "Transformation data collections" program.
 We gratefully acknowledge input from the following persons who contributed to data collection `r acknowledge$name %>% unique() %>% paste( collapse= ", ")`.
-We acknowledge XXXXXXXXXXx institutions, who XXXXX
+We acknowledge the work of all Australian taxonomists and their supporting institutions, whose long-term work on describing the flora has provided a rich source of data for AusTraits.
 
@@
 
-RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW,
-SA, CB, JL error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
+RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SA, CB, JL error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
 All other authors contributed the raw data and metadata underpinning the resource, reviewed the harmonised data for errors, and reviewed the final paper for publication.
@@
 
-This document provides information on the structure of AusTraits and corresponds to Version `r austraits$definitions$austraits$elements$build_info$elements$version$value` of the dataset.
+This document provides information on the structure of AusTraits and corresponds to version `r austraits$definitions$austraits$elements$build_info$elements$version$value` of the dataset.
 
@@
 
-![Number of taxa with trait records by plant tissue and trait category, for data that is  (A.) georeferenced and (B.) not georeferenced. ](figures/austraits_fig_composition.pdf)
+![Number of taxa with trait records by plant tissue and trait category, for data that is  (A.) geo-referenced, and (B.) not geo-referenced. ](figures/austraits_fig_composition.pdf)
 
@@
 
-![Coverage of georeferenced trait records acorss Australian climate space and gerography for traits in different categories.
+![Coverage of georeferenced trait records across Australian climatic and geographic space for traits in different categories.
 (a.) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s major biomes diagram.
-Climate data were extracted from Worldclim (www.worldclim.org).(B.) Locations of georeferenced-sites grouped by plant tissues.
+Climate data were extracted from Worldclim \citep{Fick-2017}.(B.) Locations of geo-referenced records for different plant tissues.
 ](figures/austraits_fig_biome_map.pdf)
@@
 
-![The phylogenetic representation of traits in the AusTraits database for the subset of 2000 randomly sampled taxa.
+![Phylogenetic representation of traits in AusTraits for the subset of 2000 randomly sampled taxa.
 The heatmap colour intensity denotes the number of traits measured within a family for each plant tissue.
-The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.](figures/austraits_fig_austraits_phylogenetic_coverage.pdf)
+The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.
+](figures/austraits_fig_austraits_phylogenetic_coverage.pdf)
 
@@
 tibble(
-  Element = names(austraits), 
-  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description") %>% gsub(".", "", ., fixed=TRUE)
+  Element = names(austraits$definitions$austraits$elements), 
+  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description") %>% gsub("`", "’", ., fixed=TRUE)
   ) %>% 
@@
 
-Table 10: Details on all traits represented in Version `r austraits$build_info$version` of AusTraits.
+Table 10: Details on all traits represented in version `r austraits$build_info$version` of AusTraits.
+Note the count of studies is less than the number of references when studies are linked to multiple references.
 
@@
          Trait = txt_wrap(Trait),
-         Group = paste0(str_to_sentence(Tissue), " (", Category, ")") %>%  str_replace_all("_", " "), 
+         Group = paste0(str_to_sentence(Tissue), " (", Category, ")"), 
          Type = ifelse(Type=="numeric", "num.", Type),

### 2020-11-09 R/ms.R

@@ -59,14 +59,14 @@ txt_wrap <- function(txt, n=22) {
   
   g <- function(txt) {
     ifelse(str_length(txt) < n, -1L, 
-           str_locate_all(txt,"_") %>% purrr::map_dbl(~.x %>% .[,"start"] %>% subset(., .<= n) %>% max(., -1L)) 
+           str_locate_all(txt,"_") %>% purrr::map_dbl(~.x %>% .[,"start"] %>% subset(., .< n) %>% max(., -1L)) 
     )
   }
   
   i <- g(txt)
-  txt2 <- ifelse(str_length(txt) > n , str_sub(txt, i+1), "")
+  txt2 <- ifelse(str_length(txt) >= n , str_sub(txt, i+1), "")
   ii <- g(txt2)
-  txt3 <- ifelse(str_length(txt2) > n , str_sub(txt2, ii+1), "")
+  txt3 <- ifelse(str_length(txt2) >= n , str_sub(txt2, ii+1), "")
   
   paste(str_sub(txt, 1, i), f(txt2, ii), f(txt3))
 }

### 2020-11-09 scripts/fig_climate_space_austraits.R

@@ -24,7 +24,6 @@ au_bioclim_table <- au_bioclim %>%
   mutate(region = as.factor("Australia"))
 
 #### 03 Extract geographic coordinates of AusTraits' sites#
-source("R/get_sites.R")
 
 # Get climate values for each site
 au_sites_clim <- sites %>% rowid_to_column("ID") %>%
@@ -76,7 +75,7 @@ austraits_climate_space <- ggplot() +
   theme_classic() +
   guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
   xlab(expression(Temperature (degree * C))) +
-  ylab(" Precipitation (cm)") +
+  ylab(" Precipitation (cm/yr)") +
   theme(text = element_text(size = 12))  +
   theme(
     legend.justification = c(-0.1, 0),

### 2020-11-09 scripts/fig_comparison_TRY.R

@@ -6,7 +6,7 @@
 # color by tissue
 p_aus_try <-
   aus_try_comparison %>% 
-  mutate_at(c("n_austraits", "n_try"), ~.x/28000*100) %>% 
+  mutate_at(c("n_austraits", "n_try"), ~.x/26000*100) %>% 
   ggplot(aes(y = n_austraits, x = n_try, color = tissue)) +
   geom_point(alpha = 0.5, size = 2) +
   geom_abline(intercept = 4, slope =1, alpha = 0.2) +

### 2020-11-09 scripts/fig_generate_phylogenies.R

@@ -43,7 +43,7 @@ austraits$traits %>%
   mutate(taxon_name = gsub(" ", "_", taxon_name)) %>% 
   filter(taxon_name %in% tree$species.list$species) %>% 
   spread(key=tissue,value=n) %>%
-  select(taxon_name, root, whole_plant, stem, reproductive, leaf) %>% 
+  select(taxon_name, root, `whole plant`, stem, reproductive, leaf) %>% 
   #mutate_if(is.integer, as.logical)%>% 
   #mutate_if(as.integer) %>% 
   as.data.frame() -> austraits_summary
@@ -52,7 +52,7 @@ austraits$traits %>%
 row.names(austraits_summary)<-austraits_summary$taxon_name
 austraits_summary$taxon_name<-NULL
 colnames(austraits_summary)<-c( "Root", # Root Traits
-                                    "Plant", # whole_plant traits
+                                    "Plant", # whole plant traits
                                     "Stem", # Stem Traits
                                     "Reprod.", # Reproductive traits
                                     "Leaf" # Leaf Traits

### 2020-11-09 scripts/fig_site_locations.R

@@ -45,6 +45,7 @@ austraits_site_locations_by_tissue_fig_2 <-
   produce_site_map("latitude (deg)", "longitude (deg)", feature = "tissue") +
   theme(
     legend.position = "bottom",
+    legend.title = element_blank(),
     strip.background = element_blank(),
     strip.text.x = element_text(size = 12),
     legend.key.height = unit(0.5, "cm"),

### 2020-11-09 scripts/fig_trait_composition.R

@@ -1,14 +1,14 @@
 ## Austraits: trait composition plot
 ## Rearrange order of tissue
 positions <-
-  c("leaf", "stem", "bark", "root", "reproductive", "whole_plant")
+  c("leaf", "stem", "bark", "root", "reproductive", "whole plant")
 
 # Georeferenced trait data
 austraits_composition_georef <- 
   trait_composition_list %>%
   filter(!is.na(`latitude (deg)`),!is.na(`longitude (deg)`)) %>%
   group_by(tissue, category) %>%
-  summarise(n = n_distinct(taxon_name)) %>%
+  summarise(n = n_distinct(taxon_name) %>% replace_na(0)) %>%
   rename(Category = category) %>%
   drop_na() %>%
   ggplot(aes(x = tissue, y = n, fill = Category)) +
@@ -20,19 +20,19 @@ austraits_composition_georef <-
     limits = c(1, 20000),
     labels = trans_format("log10", math_format(10 ^ .x))
   ) +
-  xlab("Major plant part") + ylab("Species (n)") +
+  xlab("") + ylab("Species (n)") +
   scale_x_discrete(limits = positions) +
-  theme(legend.position = "right") +
+  theme(legend.position = "none") +
   scale_fill_viridis_d() +
   annotation_logticks(sides = "l") +
-  ggtitle("Georeferenced")
+  ggtitle("Geo-referenced")
 
 # Non-georeferenced trait data
 austraits_composition_non_georef <- 
   trait_composition_list %>%
   filter(is.na(`latitude (deg)`), is.na(`longitude (deg)`)) %>%
   group_by(tissue, category) %>%
-  summarise(n = n_distinct(taxon_name)) %>%
+  summarise(n = n_distinct(taxon_name)%>% replace_na(0)) %>%
   rename(Category = category) %>%
   drop_na() %>%
   ggplot(aes(x = tissue, y = n, fill = Category)) +
@@ -49,4 +49,4 @@ austraits_composition_non_georef <-
   theme(legend.position = "right") +
   scale_fill_viridis_d() +
   annotation_logticks(sides = "l") +
-  ggtitle("Non-Georeferenced") 
+  ggtitle("Not geo-referenced") 

### 2020-11-10 ms.Rmd
Better citation formatting

@@
 - \usepackage{inputenc}
-#- \usepackage[T1]{fontenc}
 - \usepackage{float}
@@
 - \usepackage{xcolor}
+- \usepackage[square,sort,comma,numbers]{natbib}
 bibliography: [data/references.bib]

### 2020-11-10 ms.Rmd
Update data to 2.1.0

@@
 ```{r authors}
-authors_lead <- read_csv("data/lead_authors.csv", col_types = cols(.default = "c"))
-
-# check list of acknowledgments and assistants
-
-data_contributors <- 
-  austraits$contributors %>% 
-    mutate(name = gsub("Richard Duncan", "Richard P. Duncan", name, fixed=TRUE)) %>%
-    mutate(
-      acknowledge = ifelse(
-          # old studies or excluded
-          dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981") |
-           role %in% c("assistant") |
-           # deceased
-           name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper") |
-            # no response or can't locate
-           name %in% c("Emma Laxton", "Huw Morgan", "Wendy Cooper", "Muhammad Islam", "Michael Lawes"),
-            TRUE, FALSE),
-      year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
-    arrange(year) %>%
-    # take most recent address for each author
-    group_by(name) %>% 
-    summarise(institution = last(institution),
-              dataset_ids = paste(dataset_id, collapse = ", "),
-              acknowledge = all(acknowledge)) %>% 
-    ungroup() %>% 
-    distinct() %>% 
-    mutate(surname = str_split(name, " ") %>% map_chr(last)) %>% 
-    rename(affiliation = institution) %>%
-    arrange(surname) %>% 
-    select(-surname)
-
-authors <- bind_rows(authors_lead, 
-                     data_contributors %>% filter(! (name %in% authors_lead$name), !acknowledge)
-                    )
-
-acknowledge <- data_contributors %>% filter(! (name %in% authors$name), acknowledge)
-
-authors %>% select(dataset_ids, everything()) %>% mutate_all(~ replace_na(.x, "")) %>% write_csv("data/author_list.csv")
-
-# Update author info from google sheets
+data_contributors <- read_csv("data/author_list.csv")
+authors <- data_contributors %>% filter(!acknowledge)
+acknowledge <- data_contributors %>% filter(acknowledge)
 
@@
 Climate data were extracted from Worldclim \citep{Fick-2017}.(B.) Locations of geo-referenced records for different plant tissues.
-](figures/austraits_fig_biome_map.pdf)
+](figures/austraits_fig_biome_map.png)
 

### 2020-11-10 R/load_data.R

@@ -2,7 +2,7 @@
 load_data <- function() {
   
   if(!exists("austraits", envir = .GlobalEnv )) {
-    austraits <- readRDS("data/austraits_2.0.0.rds")
+    austraits <- readRDS("data/austraits_2.1.0.rds")
     assign("austraits", austraits, envir = .GlobalEnv)
   }
   
@@ -120,22 +120,16 @@ load_data <- function() {
   aus_try_comparison <- 
     trait_category_lookup %>% 
     select(trait_name, `TRY name`, tissue) %>%
+    mutate(`TRY name` = ifelse(duplicated(`TRY name`), NA, `TRY name`)) %>%
     full_join(by = "trait_name",
       austraits$traits %>%
       group_by(trait_name) %>%
-      summarise(n_austraits = n_distinct(taxon_name))
+      summarise(n_austraits = n_distinct(taxon_name)) 
               ) %>%
     full_join(try_by_sp, by = c("TRY name" = "Trait")) %>%
     mutate_if(is.numeric, ~ replace_na(., 0)) %>%
-    mutate(tissue = ifelse(is.na(trait_name),`TRY tissue`, tissue)) %>%
-    group_by(`TRY name`) %>% 
-    summarise(
-      trait_name = trait_name[1],
-      n_austraits = sum(n_austraits), 
-      n_try = n_try[1],
-      tissue = subset(tissue, !is.na(tissue))[1]
-    ) %>% ungroup()
-  
+    mutate(tissue = ifelse(is.na(trait_name),`TRY tissue`, tissue))
+
   assign("aus_try_comparison", aus_try_comparison, envir = .GlobalEnv)
     
 }

### 2020-11-10 R/ms.R

@@ -15,7 +15,7 @@ list1_to_df <- function(my_list) {
 ## Separate into groups of max size n. 
 ## This is needed for latex table, overwise citations
 ## spill off the page
-format_cites_n <- function(x, n = 6){
+format_cites_n <- function(x, n = 5){
   
   # Make sorted vector of all unique refs
   x2 <- str_split(x, ", ") %>% unlist() %>% sort()

### 2020-11-10 build_figures.R

@@ -17,6 +17,12 @@ pdf("figures/austraits_fig_biome_map.pdf",
 austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A')
 dev.off()
 
+png("figures/austraits_fig_biome_map.png",
+    height = 12,
+    width = 8, units="in", res=300)
+austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A')
+dev.off()
+
 #### Figure 3: trait composition count ####
 
 source("scripts/fig_trait_composition.R")
@@ -33,7 +39,7 @@ source("scripts/fig_link_tree_and_trait_count_matrix.R") # C
 
 
 pdf(
-  "figures/austraits_fig_austraits_phylogenetic_coverage.pdf",
+  "figures/austraits_fig_austraits_phylogenetic_coverage1.pdf",
   height = 12,
   width = 12
 )
@@ -45,8 +51,8 @@ dev.off()
 source("scripts/fig_comparison_TRY.R") 
  
 pdf( "figures/austraits_fig_austraits_try_comparison.pdf",
-     height = 6,
-     width = 5)
+     height = 4,
+     width = 4)
 p_aus_try
 dev.off()
 

### 2020-11-10 scripts/fig_comparison_TRY.R

@@ -8,7 +8,7 @@ p_aus_try <-
   aus_try_comparison %>% 
   mutate_at(c("n_austraits", "n_try"), ~.x/26000*100) %>% 
   ggplot(aes(y = n_austraits, x = n_try, color = tissue)) +
-  geom_point(alpha = 0.5, size = 2) +
+  geom_point(alpha = 1, size = 2) +
   geom_abline(intercept = 4, slope =1, alpha = 0.2) +
   geom_abline(intercept = 3, slope =1, alpha = 0.2) +
   geom_abline(intercept = 2, slope = 1, alpha = 0.2) +
@@ -19,8 +19,8 @@ p_aus_try <-
   geom_abline(intercept = -3, slope = 1, alpha = 0.2) +
   geom_abline(intercept = -4, slope = 1, alpha = 0.2) +
   theme_classic() +
-  scale_x_log10(limits =c(1E-3, 100), breaks = c(0.001, 0.01, 0.1,1, 10, 100), labels = function(.x) scales::comma(.x, accuracy = 0.01), expand= expansion(0)) +
-  scale_y_log10(limits =c(1E-3, 100), breaks = c(0.001, 0.01, 0.1,1, 10, 100), labels = function(.x) scales::comma(.x, accuracy = 0.01), expand= expansion(0)) +
+  scale_x_log10(limits =c(1E-3, 120), breaks = c(0.001, 0.01, 0.1,1, 10, 100), labels = function(.x) scales::comma(.x, accuracy = 0.01), expand= expansion(0)) +
+  scale_y_log10(limits =c(1E-3, 120), breaks = c(0.001, 0.01, 0.1,1, 10, 100), labels = function(.x) scales::comma(.x, accuracy = 0.01), expand= expansion(0)) +
   ylab("Taxa in AusTraits (% flora)") +
   xlab("Species in TRY (% flora)") +
   theme(legend.position = "bottom") +

### 2020-11-10 scripts/fig_trait_composition.R

@@ -17,7 +17,7 @@ austraits_composition_georef <-
   scale_y_log10(
     breaks = trans_breaks("log10", function(x)
       10 ^ x),
-    limits = c(1, 20000),
+    limits = c(1, 100000),
     labels = trans_format("log10", math_format(10 ^ .x))
   ) +
   xlab("") + ylab("Species (n)") +
@@ -41,7 +41,7 @@ austraits_composition_non_georef <-
   scale_y_log10(
     breaks = trans_breaks("log10", function(x)
       10 ^ x),
-    limits = c(1, 20000),
+    limits = c(1, 100000),
     labels = trans_format("log10", math_format(10 ^ .x))
   ) +
   xlab("Major plant part") + ylab("Species (n)") +

### 2020-11-10 scripts/update_author_list.R
Update author lists

@@ -1,0 +1,46 @@
+
+
+# check list of acknowledgements and assistants
+
+data_contributors <- 
+  austraits$contributors %>% 
+  mutate(
+    acknowledge = ifelse(
+      # old studies or excluded
+      dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981", "Clarke_2015") |
+        role %in% c("assistant") |
+        # deceased
+        name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper", "J. Bastow Wilson") |
+        # no response or can't locate
+        name %in% c("Emma Laxton", "Huw Morgan", "Wendy Cooper", "Muhammad Islam", "Kate McClenahan") |
+        # can't locate
+        name %in% c("Ian Davidson", "Kate Hughes", "Kirsten Knox", "Paula Peeters", "Burak Pekin", "Anne Sjostrom")
+        ,
+      TRUE, FALSE),
+    year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
+  arrange(year) %>%
+  # take most recent address for each author
+  group_by(name) %>% 
+  summarise(institution = last(institution),
+            dataset_ids = paste(dataset_id, collapse = ", "),
+            acknowledge = all(acknowledge)) %>% 
+  ungroup() %>% 
+  distinct() %>% 
+  mutate(surname = str_split(name, " ") %>% map_chr(last)) %>% 
+  rename(affiliation = institution) %>%
+  arrange(surname) %>% 
+  select(-surname)
+
+affiliations_sheet <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1nYE5V7utbExcfGmsfILjoXdVsEMDi5CLF_H-uE57EOo/edit#gid=280623389") %>% 
+  filter(!is.na(name))
+
+authors_lead <- read_csv("data/lead_authors.csv")
+
+authors <- bind_rows(authors_lead %>% mutate(acknowledge=FALSE), 
+                     data_contributors %>% filter(! (name %in% authors_lead$name))
+  ) %>% 
+  rename(affiliation_dataset = affiliation) %>%  
+  left_join(by="name", affiliations_sheet %>% select(name, affiliation))
+
+write_csv(authors, "data/author_list.csv")
+

### 2020-11-10 ms.Rmd
Update author lists
fix minor grammar, quotes in tables

@@
 
-We introduce the AusTraits database - a compilation of measurements of plant traits for taxa in the Australian Flora (hereafter AusTraits).
+We introduce the AusTraits database - a compilation of measurements of plant traits for taxa in the Australian flora (hereafter AusTraits).
 AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$taxon_name %>% n_distinct()` taxa from field campaigns, published literature, taxonomic monographs, and individual taxa descriptions.
@@
 Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function \citep{chapin1993evolution, adler2014functional, diaz1998plant,violle2007let}.
-Traits thereby provide useful indicators of species behaviours in communities and ecosystems, regardless of their taxonomy \citep{violle2007let, funk2017revisiting}.
+Traits thereby provide useful indicators of species' behaviours in communities and ecosystems, regardless of their taxonomy \citep{violle2007let, funk2017revisiting}.
 Through global initiatives the volume of available trait information for plants has grown rapidly in the last two decades \citep{Kattge2011try, Kattge-2020}.
@@
 One such region is Australia; a continent with a flora of c.
-26,000 native higher-plant species \citep{council2010australian}.
+26,000 native higher-plant species \citep{apc-2020}.
 While significant investment has been made in curating and digitising herbarium collections and observation records in Australia over the last two decades (e.g.
@@
 Versions are labeled using semantic versioning to indicate the change between versions \citep{Falster-2019}.
-Static versions of the AusTraits, including version`r austraits$build_info$version` used in this descriptor, are available on the project website (https://github.com/traitecoevo/austraits) and Zenodo (Data Citation 1).
+Static versions of the AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available on the project website (https://github.com/traitecoevo/austraits) and Zenodo \citep{falster_2019}.
 The latest data can also be downloaded directly from the project website.
@@
 
-There were substantial differences in coverage among different tissues and trait types, also with respect to number of georeferenced points (Figure 2).The most common traits are nongeoreferenced records from floras.
+There were substantial differences in coverage among different tissues and trait types, also with respect to number of geo-referenced points (Figure 2).The most common traits are non geo-referenced records from floras.
 Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 2a).
@@
 
-Third, are providing a pathway for user feedback.
-As AusTraits is a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
+Third, ww provide a pathway for user feedback.
+AusTraits is a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
 As such, we welcome reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
+Feedback can be posted as an issue directly at the project website: https://github.com/traitecoevo/austraits.build/issues/new/.
 
@@
 
-This document provides information on the structure of AusTraits and corresponds to version `r austraits$definitions$austraits$elements$build_info$elements$version$value` of the dataset.
+This document provides information on the structure of AusTraits and corresponds to version `r austraits$build_info$version` of the dataset.
 
@@
 
-![Number of taxa with trait records by plant tissue and trait category, for data that is  (A.) geo-referenced, and (B.) not geo-referenced. ](figures/austraits_fig_composition.pdf)
+![Number of taxa with trait records by plant tissue and trait category, for data that is  (A.) geo-referenced, and (B.) not geo-referenced.
+Many records without a geo-reference come from botanical collections, such as floras.
+](figures/austraits_fig_composition.pdf)
 
@@
 
-![Coverage of georeferenced trait records across Australian climatic and geographic space for traits in different categories.
+![Coverage of geo-referenced trait records across Australian climatic and geographic space for traits in different categories.
 (a.) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s major biomes diagram.
@@
 ```{r}
+clean_quotes <- function(txt) {
+  str_replace_all(txt, "`([a-zA-Z_ 0-9()\\.\\\\_]+)`", " texttt{\\1}") %>% 
+  str_replace_all(fixed(" texttt"), fixed("\\texttt")) 
+}
+
 tibble(
   Element = names(austraits$definitions$austraits$elements), 
-  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description") %>% gsub("`", "’", ., fixed=TRUE)
+  Contents = sapply(austraits$definitions$austraits$elements, "[[", "description")
   ) %>% 
@@
   str_replace_all(fixed("gray!6"), fixed("gray!20"))  %>%
+    clean_quotes() %>%
   writeLines()
@@
     list1_to_df() %>%
+#    mutate(value = clean_quotes(value)) %>%
     knitr::kable("latex",  booktabs = TRUE) %>% 
@@
     str_replace_all(fixed("gray!6"), fixed("gray!20"))  %>%
+    clean_quotes() %>%
     writeLines()
@@
   str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
+  clean_quotes() %>%
   writeLines()

### 2020-12-03 ms.Rmd
Update author details, references, minor corrections

@@
 
+**Journal:** For submission as a 'Data Descriptor' to the journal [Scientific Data](https://www.nature.com/sdata/).
+
 ```{r global_options, include=FALSE}
@@
 ```{r authors}
-data_contributors <- read_csv("data/author_list.csv")
-authors <- data_contributors %>% filter(!acknowledge)
-acknowledge <- data_contributors %>% filter(acknowledge)
+authors <- read_csv("data/author_list.csv")
+acknowledge <- read_csv("data/acknowledge_list.csv")
 
@@
 
-writeLines("^\\*^ contributed equally\n\n")
+writeLines("^\\*^ contributed equally")
+
+writeLines("\n\n \\footnotesize \n")
 
@@
 ```
+
+\normalsize
 
@@
 This data descriptor provides information on version  `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxa combinations.
-We envision AusTraits as an on-
-going collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
+We envision AusTraits as an ongoing collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
 
@@
 This level of species coverage limits our ability to use traits to understand and ultimately manage Australian vegetation \citep{Kissling_2018}.
-While initiatives such as TRY \citep{Kattge-2020} and the Open Traits Network \citep{Gallagher-2020} are working towards global synthesis of trait data a stronger representation of Australian plant taxa in these efforts is essential given the high richness and endemicity of this continental flora.
+While initiatives such as TRY \citep{Kattge-2020} and the Open Traits Network \citep{Gallagher-2020} are working towards global synthesis of trait data, a stronger representation of Australian plant taxa in these efforts is essential given the high richness and endemicity of this continental flora.
 
-Here we introduce the AusTraits database (hereafter simply AusTraits),  a compilation of plant traits for the Australian flora.
+Here we introduce the AusTraits database (hereafter AusTraits), a compilation of plant traits for the Australian flora.
 Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
@@
 
-AusTraits version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments (see Online Table 1).
-Initially we identified a list of candidate traits of interest then identified primary sources containing measurements for these traits, before contacting authors for access.
+AusTraits version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments (Table 10).
+Initially we identified a list of candidate traits of interest, then identified primary sources containing measurements for these traits, before contacting authors for access.
 As the compilation grew, we expanded the list of traits considered to include any measurable quantity that had been quantified for a moderate number of taxa (n > 20).
@@
 
-A full list of traits and their sources appears in Table 10 (available online only).
+A full list of traits and their sources appears in Table 10 (available online).
 This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics \citep{Garnier-2017}.
@@
 The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection (e.g.
-location coordinates, light levels s) and information about the methods used to derive measurements (e.g.
+location coordinates, light levels) and information about the methods used to derive measurements (e.g.
 number of replicates, equipment used).
@@
 
-To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R \citep{R-2020} and using custom code and drawing on functions from the packages `tidyverse` \citep{wickham2016tidyverse}, `stringr` \citep{wickham2017stringr}, `yaml` \citep{stephens2014yaml}, `remake` \citep{fitzjohn2016remake},  `knitr` \citep{xie1general} and `rmarkdown` \citep{allaire2015rmarkdown}.
-In this workflow, we performed a series of operations, including reformatting data into a standardised format, generating observation id’s for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
-Successive versions of AusTraits iterate through the steps in Figure 1, to incorporate new data and correct identified errors, leading to high-quality harmonised dataset.
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R \citep{R-2020}, using custom code, and the packages `tidyverse` \citep{wickham2016tidyverse}, `stringr` \citep{wickham2017stringr}, `yaml` \citep{stephens2014yaml}, `remake` \citep{fitzjohn2016remake},  `knitr` \citep{xie1general}, and `rmarkdown` \citep{allaire2015rmarkdown}.
+In this workflow, we performed a series of operations, including reformatting data into a standardised format, generating observation ids for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
+Successive versions of AusTraits iterate through the steps in Figure 1, to incorporate new data and correct identified errors, leading to a high-quality, harmonised dataset.
 
@@
 Versions are labeled using semantic versioning to indicate the change between versions \citep{Falster-2019}.
-Static versions of the AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available on the project website (https://github.com/traitecoevo/austraits) and Zenodo \citep{falster_2019}.
+Static versions of the AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available on the project website (http://traitecoevo.github.io/austraits.build/) and Zenodo \citep{falster_2019}.
 The latest data can also be downloaded directly from the project website.
 As validation (see Technical Validation, below) and data entry is ongoing, users are recommended to pull data from the static releases, to ensure results in their downstream analyses remain consistent as the database is updated.
+
+Data is released under a CC-BY license enabling reuse with attribution -- being a citation of this descriptor and, where possible, original sources.
 
@@
 
-The number of vascular plants (species, subspecies, varieties and forma) in Australia is  ca.
-26,000 \citep{apc-2020}.
-Version `r austraits$build_info$version` of AusTraits included at least one record for `r austraits$taxa$taxon_name %>% n_distinct()`, or about 80% of taxa.
+The number of accepted vascular plant species in the APC (as of May 2020) is around 24,750 \citep{apc-2020}.
+Version `r austraits$build_info$version` of AusTraits includes at least one record for `r austraits$taxa %>% filter(taxonRank == "Species") %>% nrow() %>% format(big.mark=",")`, or about 97% of taxa.
 Five traits (leaf_length, leaf_width, plant_height, life_history, plant_growth_form) have records for more than 50% of taxa.
@@
 
-There were substantial differences in coverage among different tissues and trait types, also with respect to number of geo-referenced points (Figure 2).The most common traits are non geo-referenced records from floras.
+There were substantial differences in coverage among different tissues and trait types, also with respect to number of geo-referenced points (Figure 2).
+The most common traits are non geo-referenced records from floras.
 Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 2a).
@@
 One exception is for root traits, where taxa within Poaceae have large amounts of information available relative to other plant families.
-A cluster of taxa within the family Myrtaceceae have little leaf information available, while reproductive information is limited for species near the base of the tree.
+A cluster of taxa within the family Myrtaceae have little leaf information available, while reproductive information is limited for species near the base of the tree.
 
@@
 Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
-Data that did not pass these tests is moved to a separate spreadsheet (“excluded_data”) that is also made available for use and reviewed.
+Data that did not pass these tests were moved to a separate spreadsheet (“excluded_data”) that is also made available for use and review.
 
-Third, ww provide a pathway for user feedback.
+Third, we provide a pathway for user feedback.
 AusTraits is a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
 As such, we welcome reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
-Feedback can be posted as an issue directly at the project website: https://github.com/traitecoevo/austraits.build/issues/new/.
+Feedback can be posted as an issue directly at the project.
 
@@
 
-Each data release is available in multiple formats.
-First, as a compressed folder containing text files for each of the main components.
-Second, as a compressed R object, enabling easy loading into R for those using that platform.
-
-Data is released under a CC-BY license enabling reuse with attribution -- being a citation of this descriptor and, where possible, original sources.
+Each data release is available in multiple formats: first, as a compressed folder containing text files for each of the main components, second, as a compressed R object, enabling easy loading into R for those using that platform.
 
@@
 To create the phylogenetic tree in Figure 5, we pruned a master tree for all higher plants \citep{Smith-2018} using the package `V.PhyloMaker` \citep{Jin-2020} and visualising via `ggtree` \citep{Yu-2017}.
-To create Figure 3A, we visualised the Whittaker's biome using the package `plotbiomes` \citep{Stefan-2020}.
+To create Figure 3A, we used the package `plotbiomes` \citep{Stefan-2020} to create the baseline plot of biomes.
 
@@
 
-All code, raw and compiled data are hosted within GitHub repositories under the Trait Ecology and Evolution organisation (see https://github.com/traitecoevo/austraits.build, https://github.com/traitecoevo/austraits).
+All code, raw and compiled data are hosted within GitHub repositories under the Trait Ecology and Evolution organisation (http://traitecoevo.github.io/austraits.build/).
 The archived material includes all data sources and code for rebuilding the compiled dataset.
 The code used to produce this paper is available at http://github.com/traitecoevo/austraits_ms.
+(All code will be made available prior to final publication.)
 
@@
 
-RVG, IJW conceived the original idea; RVG, EHW, CB collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SA, CB, JL error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
+RVG, IJW conceived the original idea; RVG, EHW, CB, SA collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SCA, CB, JL error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
 All other authors contributed the raw data and metadata underpinning the resource, reviewed the harmonised data for errors, and reviewed the final paper for publication.
@@
 
-AusTraits harmonises data on `r austraits$traits$trait_name %>% unique() %>% length()` traits from `r austraits$methods$dataset_id %>% unique() %>% length()` different sources, including field campaigns, published literature, taxonomic monographs, and individual taxa descriptions.
+AusTraits harmonises data on `r austraits$traits$trait_name %>% unique() %>% length()` traits from `r austraits$methods$dataset_id %>% unique() %>% length()` different sources, including field campaigns, published literature, taxonomic monographs, and individual taxon descriptions.
 
@@
 Trait observations are accessed from original data sources, including published floras and field campaigns.
-Features such as variables names, units and taxonomy are harmonised to a common standard.
+Features such as variable names, units and taxonomy are harmonised to a common standard.
 Versioned releases are distributed to users, allowing the dataset to be used and re-used in a reproducible way.
@@
 
-![Number of taxa with trait records by plant tissue and trait category, for data that is  (A.) geo-referenced, and (B.) not geo-referenced.
+![Number of taxa with trait records by plant tissue and trait category, for data that are  (A) Geo-referenced, and (B) Not geo-referenced.
 Many records without a geo-reference come from botanical collections, such as floras.
@@
 ![Coverage of geo-referenced trait records across Australian climatic and geographic space for traits in different categories.
-(a.) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s major biomes diagram.
-Climate data were extracted from Worldclim \citep{Fick-2017}.(B.) Locations of geo-referenced records for different plant tissues.
+(A) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s classifctaion of majore biomes by climate \citep{Whittaker-1975}.
+Climate data were extracted at 10" resolution from WorldClim \citep{Fick-2017}.(B) Locations of geo-referenced records for different plant tissues.
 ](figures/austraits_fig_biome_map.png)
@@
 
-![Phylogenetic representation of traits in AusTraits for the subset of 2000 randomly sampled taxa.
+![Phylogenetic distribution of trait data in AusTraits for a subset of 2000 randomly sampled taxa.
 The heatmap colour intensity denotes the number of traits measured within a family for each plant tissue.
@@
 clean_quotes <- function(txt) {
-  str_replace_all(txt, "`([a-zA-Z_ 0-9()\\.\\\\_]+)`", " texttt{\\1}") %>% 
+  str_replace_all(txt, "`([a-zA-Z_ \\-0-9()\\.\\\\_]+)`", " texttt{\\1}") %>% 
   str_replace_all(fixed(" texttt"), fixed("\\texttt")) 

### 2020-12-03 scripts/update_author_list.Rmd
Update author details, references, minor corrections

# check list of acknowledgments and assistants

```{r}
authors_lead <- read_csv("data/lead_authors.csv")

data_contributors <- 
  austraits$contributors %>% 
  mutate(
    acknowledge = ifelse(
      # old studies or excluded
      dataset_id %in% c("Westman_1977", "Hall_1981", "Hocking_1982", "Kuo_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981") |
        role %in% c("assistant")
        , TRUE, FALSE),
    year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
  arrange(year) %>%
  # take most recent address for each author
  group_by(name) %>% 
  summarise(institution = last(institution),
            dataset_ids = paste(dataset_id, collapse = ", "),
            acknowledge = all(acknowledge)) %>% 
  ungroup() %>% 
  distinct() %>% 
  mutate(surname = str_split(name, " ") %>% map_chr(last)) %>% 
  rename(affiliation = institution) %>%
  arrange(surname) %>% 
  select(-surname)

authors <- bind_rows(authors_lead %>% mutate(acknowledge=FALSE), 
                     data_contributors %>% filter(! (name %in% authors_lead$name))
  ) %>% 
  rename(affiliation_dataset = affiliation)
```

```{r}
affiliations_sheet <- 
  googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1nYE5V7utbExcfGmsfILjoXdVsEMDi5CLF_H-uE57EOo/edit#gid=280623389") %>% 
  filter(!is.na(name))

responses_sheet <- 
  googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1UJKSb5YMzIMTmEd-PIqgHwxpSkCEHMJN_gf0r1vTmxU/edit#gid=499607997") %>% 
  rename(name = `Name as shown on paper`)

```

Mark acknowledgments

```{r}
authors2 <- 
  authors %>% 
  mutate(
    #known 
    acknowledge = ifelse(name %in% (affiliations_sheet %>% filter(status != "author") %>% pull(name)), TRUE, acknowledge)
  )

# Check who hasn't responded
authors2 %>% filter(!acknowledge, !(name %in% responses_sheet$name)) %>% pull(name)

authors3 <- 
  authors2 %>% 
  mutate(
    # no response
    acknowledge = ifelse(!(name %in% responses_sheet$name), TRUE, acknowledge)
  )

# Save acknowledgments
authors3 %>% filter(acknowledge) %>% select(name, dataset_ids) %>% write_csv("data/acknowledge_list.csv")

authors4 <- 
  authors3 %>% 
  filter(!acknowledge) %>%
  left_join(by="name", affiliations_sheet %>% select(name, email, affiliation))  %>%
  left_join(by="name",
    responses_sheet %>% 
    select(name, 
            corrected_name = `If you answered no to above, please provide a corrected name`,
            corrected_affiliation = `If you answered no to above, please provide a corrected affiliation`,
            preferred_email = `Preferred email address you would like us to use?  (one only please)`,
            orcid = `Please provide your ORCID id, if you have one.` 
            ) %>%
    mutate(orcid = orcid %>% str_remove_all("[https]+://")  %>% str_remove_all("orcid.org/") )
    ) %>%
  mutate(
    name = ifelse(!is.na(corrected_name), corrected_name, name),
    affiliation = ifelse(!is.na(corrected_affiliation), corrected_affiliation, affiliation),
    email = ifelse(!is.na(preferred_email), preferred_email, email)
  )

authors4 %>% select (name, affiliation, orcid) %>% 
write_csv("data/author_list.csv")

authors4 %>% select (name, affiliation, email, orcid) %>% 
write_csv("data/author_list_email.csv")

```

```{r}
authors4 %>% pull(email) %>% paste(collapse= ", ") %>% clipr::write_clip()
authors %>% pull(name)
```

### 2020-12-03 scripts/update_author_list.R

@@ -1,46 +1,0 @@
-
-
-# check list of acknowledgements and assistants
-
-data_contributors <- 
-  austraits$contributors %>% 
-  mutate(
-    acknowledge = ifelse(
-      # old studies or excluded
-      dataset_id %in% c("Lawes_2012", "Westman_1977", "Hall_1981", "Hocking_1982", "Hocking_1986", "Toelken_1996", "Brock_1993", "Forster_1992", "Forster_1995", "Goble_1981", "Clarke_2015") |
-        role %in% c("assistant") |
-        # deceased
-        name %in% c("Peter Clarke", "Barbara Rice", "Peter Myerscough", "Lyn Craven", "Harold Trevor Clifford", "William Cooper", "J. Bastow Wilson") |
-        # no response or can't locate
-        name %in% c("Emma Laxton", "Huw Morgan", "Wendy Cooper", "Muhammad Islam", "Kate McClenahan") |
-        # can't locate
-        name %in% c("Ian Davidson", "Kate Hughes", "Kirsten Knox", "Paula Peeters", "Burak Pekin", "Anne Sjostrom")
-        ,
-      TRUE, FALSE),
-    year = dataset_id %>% str_split_fixed("_", 4) %>% .[,2]) %>% 
-  arrange(year) %>%
-  # take most recent address for each author
-  group_by(name) %>% 
-  summarise(institution = last(institution),
-            dataset_ids = paste(dataset_id, collapse = ", "),
-            acknowledge = all(acknowledge)) %>% 
-  ungroup() %>% 
-  distinct() %>% 
-  mutate(surname = str_split(name, " ") %>% map_chr(last)) %>% 
-  rename(affiliation = institution) %>%
-  arrange(surname) %>% 
-  select(-surname)
-
-affiliations_sheet <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1nYE5V7utbExcfGmsfILjoXdVsEMDi5CLF_H-uE57EOo/edit#gid=280623389") %>% 
-  filter(!is.na(name))
-
-authors_lead <- read_csv("data/lead_authors.csv")
-
-authors <- bind_rows(authors_lead %>% mutate(acknowledge=FALSE), 
-                     data_contributors %>% filter(! (name %in% authors_lead$name))
-  ) %>% 
-  rename(affiliation_dataset = affiliation) %>%  
-  left_join(by="name", affiliations_sheet %>% select(name, affiliation))
-
-write_csv(authors, "data/author_list.csv")
-

### 2021-04-28 ms.Rmd
Update data to v3.0.0, address reviews

@@
 ---
-title: "AusTraits -- a curated plant trait database for the Australian flora"
+title: "AusTraits: A curated plant trait database for the Australian flora"
 editor_options:
@@
 
+refs <- 
   c(
@@
   readLines("data/references_data.bib")) %>% 
-  gsub("&", "and", ., fixed=TRUE) %>% 
-  writeLines("data/references.bib")
+  gsub("&", "and", ., fixed=TRUE)
+
+# Replace with journal abbreviation
+abbreviations <- read_csv("data/journal_abbrevs.csv") %>% 
+  filter(!is.na(abbrev))
+
+for(i in seq_along(abbreviations$title))
+  refs <- gsub(abbreviations$title[i], abbreviations$abbrev[i], refs, fixed=TRUE)
+
+writeLines(refs, "data/references.bib")
 ```
@@
 
-We introduce the AusTraits database - a compilation of measurements of plant traits for taxa in the Australian flora (hereafter AusTraits).
-AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$taxon_name %>% n_distinct()` taxa from field campaigns, published literature, taxonomic monographs, and individual taxa descriptions.
+We introduce the AusTraits database - a compilation of values of plant traits for taxa in the Australian flora (hereafter AusTraits).
+AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$taxon_name %>% n_distinct()` taxa from field campaigns, published literature, taxonomic monographs, and individual taxon descriptions.
 Traits
 vary in scope from physiological measures of performance (e.g. photosynthetic gas exchange, water-use efficiency) to morphological
-parameters (e.g. leaf area, seed mass, plant height) which link to aspects of ecological variation. AusTraits contains curated and
-harmonised individual-, species- and genus-level observations coupled to, where available, contextual information on site
-properties.
-This data descriptor provides information on version  `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxa combinations.
-We envision AusTraits as an ongoing collaborative initiative for easily archiving and sharing trait data to increase our collective understanding of the Australian flora.
+attributes (e.g. leaf area, seed mass, plant height) which link to aspects of ecological variation. AusTraits contains curated and
+harmonised individual- and species-level measurements coupled to, where available, contextual information on site properties  and experimental conditions.
+This article provides information on version  `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxon combinations.
+We envision AusTraits as an ongoing collaborative initiative for easily archiving and sharing trait data and also provides a template for other national or regional initiatives globally to fill persistent gaps in trait knowledge.
 
@@
 
-Species traits are essential metrics for comparing ecological strategies in plants arrayed across environmental space or evolutionary lineages \citep{diaz2016global, kunstler2016plant, zanne2014three, cornwell2014functional}.
+Species traits are essential for comparing ecological strategies among plants, both within any given vegetation, and also arrayed across environmental space and evolutionary lineages \citep{diaz2016global, kunstler2016plant, zanne2014three, cornwell2014functional}.
 Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function \citep{chapin1993evolution, adler2014functional, diaz1998plant,violle2007let}.
-Traits thereby provide useful indicators of species' behaviours in communities and ecosystems, regardless of their taxonomy \citep{violle2007let, funk2017revisiting}.
+Traits thereby provide useful indicators of species' behaviours in communities and ecosystems, regardless of their taxonomy \citep{Westoby-1998, violle2007let, funk2017revisiting}.
 Through global initiatives the volume of available trait information for plants has grown rapidly in the last two decades \citep{Kattge2011try, Kattge-2020}.
-However, the geographic coverage of trait observations across the globe is patchy, limiting detailed analyses of trait variation and diversity in some regions.
+However, the geographic coverage of trait measurements across the globe is patchy, limiting detailed analyses of trait variation and diversity in some regions, and, more generally, development of theory accounting for the diversity of plant strategies.
 
 One such region is Australia; a continent with a flora of c.
-26,000 native higher-plant species \citep{apc-2020}.
+28,900 native vascular plant taxa \citep{apc-2020} (including species, subspecies, varietas and forma).
 While significant investment has been made in curating and digitising herbarium collections and observation records in Australia over the last two decades (e.g.
@@
 Moreover, relatively few Australian species are represented in the leading global databases.
-For example, the international TRY database \citep{Kattge-2020} has observations for only 3830 Australian species across all collated traits.
+For example, the international TRY database \citep{Kattge-2020} has measurements for only 3830 Australian species across all collated traits.
 This level of species coverage limits our ability to use traits to understand and ultimately manage Australian vegetation \citep{Kissling_2018}.
-While initiatives such as TRY \citep{Kattge-2020} and the Open Traits Network \citep{Gallagher-2020} are working towards global synthesis of trait data, a stronger representation of Australian plant taxa in these efforts is essential given the high richness and endemicity of this continental flora.
+While initiatives such as TRY \citep{Kattge-2020} and the Open Traits Network \citep{Gallagher-2020} are working towards global synthesis of trait data, a stronger representation of Australian plant taxa in these efforts is essential, especially given the high richness and endemicity of this continental flora, and the unique contribution this makes to global floral diversity \citep{chapman2009numbers, Hopper-2004}.
 
@@
 Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
-To assemble AusTraits from diverse primary sources and make data available for reuse, we needed to overcome three main types of challenges (Figure 1): 1) Accessing data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) Harmonising these diverse sources into a federated resource, with common units, trait names, and data formats; and 3) Distributing versions of the data under suitable license.
+To assemble AusTraits from diverse primary sources and make data available for reuse, we needed to overcome three main types of challenges (Figure 1): 1) Accessing data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) Harmonising these diverse sources into a federated resource, with common taxon names, units, trait names, and data formats; and 3) Distributing versions of the data under suitable license.
 To meet this challenge, we developed a workflow which draws on emerging community standards and our collective experience building trait databases.
@@
 Prior to the development of AusTraits, data on Australian plant traits existed largely as a series of disconnected datasets collected by individual laboratories or initiatives.
+
+AusTraits has been developed as a standalone database, rather than as part of the existing global database TRY \citep{Kattge-2020}, for three reasons.
+First, we sought to establish an engaged and localised community, actively collaborating to enhance coverage of plant trait data within Australia.
+It was our opinion that a community would form more readily to fill gaps in national knowledge of traits with local ownership of the resource.
+While we will never have a counterfactual, a vibrant community excited to be part of this initiative has indeed been established and coverage is much higher for Australian species than has been achieved since TRY’s inception.
+Local ownership also aligns well with funding opportunities and national research priorities, and enables database coordinators to progress at their own speed.
+Second, we wanted to implement new features not available in TRY.
+For instance, AusTraits is fully reproducible, meaning the entire database can be rebuilt from its original sources at any point.
+This has enabled us to actively curate the dataset to reduce duplications and update mapping of traits as contributors provide feedback and trait definitions and values are edited.
+Moreover, the curation is achieved on GitHub, such that the history of all changes is traceable and, where necessary, reversible.
+Third, AusTraits will be distributed via different mechanisms to TRY.
+Data from TRY is requested via a standalone portal.
+By contrast, versions of AusTraits are being regularly archived on a third party data repository (Zenodo), built for long term data archiving and an established API.
+This will allow for active connection to a variety of end users including, but not only, TRY.
+
 We envision AusTraits as an on-going collaborative initiative for easily archiving and sharing trait data about the Australian flora.
 Open access to a comprehensive resource like this will generate significant new knowledge about the Australian flora across multiple scales of interest, as well as reduce duplication of effort in the compilation of plant trait data, particularly for research students and government agencies seeking to access information on traits.
+In coming years, AusTraits will continue to be expanded, with integrations into other biodiversity platforms and expansion of coverage into historically neglected plant lineages in trait science, such as pteridophytes (lycophytes and ferns).
+Further, through international initiatives, such as the Open Traits Network and TraitNet, linkages are being forged between plant datasets and a variety of other organismal databases \citep{Naeem-2009, Gallagher-2020}.
 
@@
 
-AusTraits version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field campaigns, botanical collections, and taxonomic treatments (Table 10).
+AusTraits version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field measurements, glasshouse and field experiments, botanical collections, and taxonomic treatments.
 Initially we identified a list of candidate traits of interest, then identified primary sources containing measurements for these traits, before contacting authors for access.
-As the compilation grew, we expanded the list of traits considered to include any measurable quantity that had been quantified for a moderate number of taxa (n > 20).
+As the compilation grew, we expanded the list of traits considered to include any measurable quantity that had been quantified for at least a moderate number of taxa (n > 20).
+
+For a small subset of sources from herbaria, providing a text description of taxa, we used regular expressions in R to extract measurements of traits from the text.
+A variety of expressions were developed to extract height, leaf/seed dimensions and growth form.
+Error checking was completed on approximately 60% of mined measurements by visually inspecting the extracted values relative to the textual descriptions.
 
@@
 
-A full list of traits and their sources appears in Table 10 (available online).
+A full list of traits and their sources appears in Online-only Table 1 `r n_records$refs %>% str_split(", ") %>% unlist() %>% sort() %>% unique %>%  
+    paste(collapse = ", ") %>% paste0("\\citep{", ., "}")`.
 This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics \citep{Garnier-2017}.
 We categorised traits based on the tissue where it is measured (`r trait_categories$tissue %>% unique() %>% sort() %>% paste(collapse=", ")`) and the type of measurement (`r trait_categories$category %>% unique() %>% sort() %>% gsub("_", " ", .) %>% paste(collapse=", ")`).
-Version `r austraits$build_info$version` of AusTraits includes  `r n_records$Type %>% table() %>% sort(decreasing = T) %>% paste(c( "", "", "and"), .,  names(.), collapse= ", ")` traits.
+Version `r austraits$build_info$version` of AusTraits includes  `r n_records$Type %>% table() %>% sort(decreasing = T) %>% paste(.,  names(.), collapse= " and ")` traits.
 
-### Database schema
+### Database structure
 
 The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection (e.g.
-location coordinates, light levels) and information about the methods used to derive measurements (e.g.
+location coordinates, light levels, whether data were collected in the field or lab) and information about the methods used to derive measurements (e.g.
 number of replicates, equipment used).
 The database contains 11 elements, as described in Table 1.
-This format was developed to include information about the trait measurements, taxa sampled, the methods used, sites, contextual information, the people involved, and citation sources.
+This format was developed to include information about the trait measurements, taxon, methods, sites, contextual information, people involved, and citation sources.
 
-For storage efficiency, the main table of traits contains relatively little information (Table 2), but can be cross linked against other tables (Tables 3-8) using identifiers for dataset, site, context, observation and taxon (Table 1).
+For storage efficiency, the main table of traits contains relatively little information (Table 2), but can be cross linked against other tables (Tables 3-8) using identifiers for dataset, site, context, observation, and taxon (Table 1).
 The `dataset_id` is ordinarily the surname of the first author and year of publication associated with the source’s primary citation (e.g.
 `Blackman_2014`).
-Trait measurements were also recorded as being one of several possible `value_type` (Table 9), reflecting the type of measurement recorded.
+Trait values were also recorded as being one of several possible value types (`value_type`) (Table 9), reflecting the type of measurement submitted by the contributor, as different sources provide different levels of detail.
+Possible values include `raw_value`, `indidivdual_mean`, `site_mean`, `multi_site_mean`, `expert_mean`, `experiment_mean`.
+Further details on the methods used for collecting each trait are provided in a `methods` table (Table 5).
 
@@
 
-To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R \citep{R-2020}, using custom code, and the packages `tidyverse` \citep{wickham2016tidyverse}, `stringr` \citep{wickham2017stringr}, `yaml` \citep{stephens2014yaml}, `remake` \citep{fitzjohn2016remake},  `knitr` \citep{xie1general}, and `rmarkdown` \citep{allaire2015rmarkdown}.
-In this workflow, we performed a series of operations, including reformatting data into a standardised format, generating observation ids for each individual measured, transforming variable names into common terms, transforming data common units, standardising terms for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
-Successive versions of AusTraits iterate through the steps in Figure 1, to incorporate new data and correct identified errors, leading to a high-quality, harmonised dataset.
-
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R \citep{R-2020}, using custom code, and the packages `tidyverse` \citep{wickham-2019}, `yaml` \citep{stephens2014yaml}, `remake` \citep{fitzjohn2016remake},  `knitr` \citep{xie1general}, and `rmarkdown` \citep{allaire2015rmarkdown}.
+In this workflow, we performed a series of operations, including reformatting data into a standardised format, generating observation ids for each set of linked measurements, transforming variable names into common terms, transforming data into common units, standardising terms (trait values) for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
 Details from each primary source were saved with minimal modification into two plain text files.
@@
 These two files provide all the information needed to compile each study into a standardised AusTraits format.
+Successive versions of AusTraits iterate through the steps in Figure 1, to incorporate new data and correct identified errors, leading to a high-quality, harmonised dataset.
+
+After importing a study, we generated a detailed report which summarised the study’s metadata and compared the study’s data values to those collected by other studies for the same traits (See Supplementary materials for an example).
+Data for continuous and categorical variables are presented in scatter plots and tables respectively.
+These reports allow first the AusTraits data curator, followed by the data contributor, to rapidly scan the metadata to confirm it has been entered correctly and the trait data to ensure it has been assigned the correct units and their categorical traits values are properly aligned with AusTraits trait values.
 
@@
 We developed a custom workflow to clean and standardise taxonomic names using the latest and most comprehensive taxonomic resources for the Australian flora: the Australian Plant Census (APC) \citep{apc-2020} and the Australian Plant Names Index (APNI) \citep{apni-2020}.
-While several automated tools exist, such as `taxize` \citep{Chamberlain-2013}, these do not currently include up to date information for Australian taxa.
+These resources document all known taxonomic names for Australian plants, including currently accepted names and synonyms.
+While several automated tools exist for updating taxonomy, such as `taxize` \citep{Chamberlain-2013}, these do not currently include up to date information for Australian taxa.
 Updates were completed in two steps.
@@
 synonym) we preferred the accepted name, but also noted the alternative records.
-When a suitable match could not be found, we manually reviewed near matches and web portals such as the Atlas of Living Australia to find a suitable match.
+For phrase names, when a suitable match could not be found, we manually reviewed near matches via web portals such as the Atlas of Living Australia to find a suitable match.
 The final resource reports both the original and the updated taxon name alongside each trait record (Table 2), as well an additional table summarising all taxonomic names changes (Table 6) and further information from the APC and APNI on all taxa included (Table 7).
+Any changes in taxonomy are exposed within the compiled dataset, enabling researchers to review these as needed.
 
@@
 
+Static versions of AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available via Zenodo \citep{falster_2019}.
+Data is released under a CC-BY license enabling reuse with attribution -- being a citation of this descriptor and, where possible, original sources.
+Deposition within Zenodo helps makes the dataset consistent with FAIR principles \citep{wilkinson2016fair}.
 As an evolving data product, successive versions of AusTraits are being released, containing updates and corrections.
 Versions are labeled using semantic versioning to indicate the change between versions \citep{Falster-2019}.
-Static versions of the AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available on the project website (http://traitecoevo.github.io/austraits.build/) and Zenodo \citep{falster_2019}.
-The latest data can also be downloaded directly from the project website.
-As validation (see Technical Validation, below) and data entry is ongoing, users are recommended to pull data from the static releases, to ensure results in their downstream analyses remain consistent as the database is updated.
+As validation (see Technical Validation, below) and data entry are ongoing, users are recommended to pull data from the particular release, to ensure results in their downstream analyses remain consistent as the database is updated.
 
-Data is released under a CC-BY license enabling reuse with attribution -- being a citation of this descriptor and, where possible, original sources.
+An R package is being developed (see [github.com/traitecoevo/austraits](http://github.com/traitecoevo/austraits)) to provide easy access and guidance on usage for those using this platform.
 
@@
 
-The number of accepted vascular plant species in the APC (as of May 2020) is around 24,750 \citep{apc-2020}.
-Version `r austraits$build_info$version` of AusTraits includes at least one record for `r austraits$taxa %>% filter(taxonRank == "Species") %>% nrow() %>% format(big.mark=",")`, or about 97% of taxa.
-Five traits (leaf_length, leaf_width, plant_height, life_history, plant_growth_form) have records for more than 50% of taxa.
+The number of accepted vascular plant taxa in the APC (as of May 2020) is around 28,981  \citep{apc-2020}.
+Version `r austraits$build_info$version` of AusTraits includes at least one record for `r austraits$taxa %>% filter(taxonRank %in% c("Species", "Subspecies", "Varietas", "Forma")) %>% nrow() %>% format(big.mark=",")` taxa (~ 93% of known taxa).
+Five traits (leaf_length, leaf_width, plant_height, life_history, plant_growth_form) have records for more than 50% of known species (Figure 2).
 Across all traits, the median number of taxa with records is 62.
-Table 10 shows the number of studies, taxa, and families recording data in AusTraits, as well as the number of geo-referenced records, for each trait.
+Online-only Table 1 shows the number of studies, taxa, and families recording data in AusTraits, as well as the number of geo-referenced records, for each trait.
+Looking across traits and tissue categories, coverage declined gradually, with moderate coverage(> 20%) for more than 50 traits (Figure 2).
+Coverage for root, stem and bark traits declined much faster than trait measurements for other plant tissues (Figure 2b).
 
-There were substantial differences in coverage among different tissues and trait types, also with respect to number of geo-referenced points (Figure 2).
-The most common traits are non geo-referenced records from floras.
-Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 2a).
+The most common traits are non geo-referenced records from floras; these are trait values representing a continental or region mean (or spread) and hence are not linked to a location.
+Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 3a).
+Coverage is notably higher for geo-referenced measurements of some tissues and trait types - such as bark stems and roots - relative to non-georeferenced measurements (Figure 3).
 
-We found that trait records were spread across the climate space of Australia (Figure 3a), as well as geographic locations (Figure 3b).
-As with most data, in Australia, the density of records was somewhat concentrated around cities or roads in remote regions, particularly for leaf traits.
+We found that trait records were spread across the climate space of Australia (Figure 4a), as well as geographic locations (Figure 4b).
+As with most data in Australia, the density of records was somewhat concentrated around cities or roads in remote regions, particularly for leaf traits.
 
-Figure 4 shows that overall coverage across a phylogenetic tree of Australian plant species is relatively unbiased, though there are some notable exceptions.
+Overall trait coverage across an estimated phylogenetic tree of Australian plant species is relatively unbiased (Figure 5), though there are some notable exceptions.
 One exception is for root traits, where taxa within Poaceae have large amounts of information available relative to other plant families.
-A cluster of taxa within the family Myrtaceae have little leaf information available, while reproductive information is limited for species near the base of the tree.
+A cluster of taxa within the family Myrtaceae which are largely from Western Australia have little leaf information available.
 
 Comparing coverage in AusTraits to the global database TRY, there were `r aus_try_comparison %>% filter(n_austraits >0, n_try >0)%>% nrow()` traits overlapping.
-Of these, AusTraits tended to contain records for more taxa, but not always (Figure 5).
-Multiple traits had more than 10 times the number of taxa represented in AusTraits.
+Of these, AusTraits tended to contain records for more taxa, but not always; multiple traits had more than 10 times the number of taxa represented in AusTraits (Figure 6).
 However, there were more records in TRY for `r aus_try_comparison %>% filter(n_try > n_austraits, n_austraits >0) %>% nrow()` traits, in particular physiological leaf traits.
-Many traits were not overlapping between the two databases (Figure 5).
+Many traits were not overlapping between the two databases (Figure 6).
 We noted that AusTraits includes more seed and fruit nutrient data; possibly reflecting the interest in Australia in understanding how fruit and seeds are provisioned in nutrient-depauperate environments.
@@
 First, we conducted a detailed review of each source based on a bespoke report, showing all data and metadata, by both an AusTraits curator and the original contributor (where possible).
-Observations for each trait were plotted against all other values for the trait in AusTraits, allowing quick identification of outliers.
+Measurements for each trait were plotted against all other values for the trait in AusTraits, allowing quick identification of outliers.
 Corrections suggested by contributors were combined back into AusTraits and made available with the next release.
+Version `r austraits$build_info$version` of AusTraits, described here, is the sixth release.
 
-Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the creators.
+Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the curators.
 Data that did not pass these tests were moved to a separate spreadsheet (“excluded_data”) that is also made available for use and review.
@@
 As such, we welcome reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
-Feedback can be posted as an issue directly at the project.
+Feedback can be posted as an issue directly at the projects GitHub page (http://traitecoevo.github.io/austraits.build).
 
@@
 Using the taxon names aligned with the APC, data can be queried against location data from the Atlas of Living Australia.
-To create the phylogenetic tree in Figure 5, we pruned a master tree for all higher plants \citep{Smith-2018} using the package `V.PhyloMaker` \citep{Jin-2020} and visualising via `ggtree` \citep{Yu-2017}.
+To create the phylogenetic tree in Figure 6, we pruned a master tree for all higher plants \citep{Smith-2018} using the package `V.PhyloMaker` \citep{Jin-2020} and visualising via `ggtree` \citep{Yu-2017}.
 To create Figure 3A, we used the package `plotbiomes` \citep{Stefan-2020} to create the baseline plot of biomes.
@@
 
-This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910) and funding from Macquarie University to Gallagher, and the Australian Research Data Commons via their "Transformation data collections" program.
+We acknowledge the work of all Australian taxonomists and their supporting institutions, whose long-term work on describing the flora has provided a rich source of data for AusTraits, including: Australian National Botanic Gardens; Australian National Herbarium; Biodiversity Science, Parks Australia; Centre for Australian National Biodiversity Research; Department of Biodiversity, Conservation and Attractions, Western Australia; Department of Environment, Land, Water and Planning, Victoria; Flora of Australia; Kew; National Herbarium of NSW; National Herbarium of Victoria; Northern Territory Herbarium; NSW Department of Planning, Industry, and Environment; Queensland Herbarium; State Herbarium of South Australia; Tasmanian Herbarium; and the Western Australian Herbarium.
 We gratefully acknowledge input from the following persons who contributed to data collection `r acknowledge$name %>% unique() %>% paste( collapse= ", ")`.
-We acknowledge the work of all Australian taxonomists and their supporting institutions, whose long-term work on describing the flora has provided a rich source of data for AusTraits.
+We thank H Cornelissen, H Poorter, SC McColl-Gausden, and one anonymous reviewer for feedback on an earlier draft, and K Levett for advice on data structures.
+This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910), a grant from Macquarie University to Gallagher, and investment from the Australian Research Data Commons (ARDC), via their "Transformation data collections" ([doi:
+10.47486/TD044](https://doi.org/10.47486/TD044)) and "Data Partnerships" ([doi: 10.47486/DP720](https://doi.org/10.47486/DP720)) programs.
+The ARDC is enabled by National Collaborative Research Investment Strategy (NCRIS).
 
@@
 
-RVG, IJW conceived the original idea; RVG, EHW, CB, SA collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SCA, CB, JL error-checked trait observations; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
+RVG, IJW conceived the original idea; RVG, EHW, CB, SA collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SCA, CB, JL error-checked trait measurements; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
 All other authors contributed the raw data and metadata underpinning the resource, reviewed the harmonised data for errors, and reviewed the final paper for publication.
@@
 
-## Overview
-
-AusTraits harmonises data on `r austraits$traits$trait_name %>% unique() %>% length()` traits from `r austraits$methods$dataset_id %>% unique() %>% length()` different sources, including field campaigns, published literature, taxonomic monographs, and individual taxon descriptions.
-
-This document provides information on the structure of AusTraits and corresponds to version `r austraits$build_info$version` of the dataset.
-
 \clearpage
@@
 ![The data curation pathway used to assemble the AusTraits database.
-Trait observations are accessed from original data sources, including published floras and field campaigns.
+Trait measurements are accessed from original data sources, including published floras and field campaigns.
 Features such as variable names, units and taxonomy are harmonised to a common standard.
@@
 
-![Number of taxa with trait records by plant tissue and trait category, for data that are  (A) Geo-referenced, and (B) Not geo-referenced.
+![Coverage of traits by taxa.
+(a) Matrix showing the coverage of taxa for each trait, with yellow indicating presence of data.
+The figure was generated with a subset of 500 randomly selected taxa.
+(b) Number of taxa with data for first 100 traits for all traits and separated by tissue.
+](figures/austraits_coverage_by_trait.png)
+
+\clearpage
+
+![Number of taxa with trait records by plant tissue and trait category, for data that are  (a) Geo-referenced, and (B) Not geo-referenced.
 Many records without a geo-reference come from botanical collections, such as floras.
@@
 ![Coverage of geo-referenced trait records across Australian climatic and geographic space for traits in different categories.
-(A) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s classifctaion of majore biomes by climate \citep{Whittaker-1975}.
+(A) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s classification of major biomes by climate \citep{Whittaker-1975}.
 Climate data were extracted at 10" resolution from WorldClim \citep{Fick-2017}.(B) Locations of geo-referenced records for different plant tissues.
@@
 The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.
-](figures/austraits_fig_austraits_phylogenetic_coverage.pdf)
+](figures/austraits_fig_austraits_phylogenetic_coverage.png)
 
@@
 
-Table 10: Details on all traits represented in version `r austraits$build_info$version` of AusTraits.
+Online-only Table 1: Details on all traits represented in version `r austraits$build_info$version` of AusTraits.
 Note the count of studies is less than the number of references when studies are linked to multiple references.

### 2021-04-28 scripts/update_author_list.Rmd
Update data to v3.0.0, address reviews

@@
 ```{r}
+load_data()
+
 authors_lead <- read_csv("data/lead_authors.csv")
@@
   rename(affiliation_dataset = affiliation)
-```
 
-```{r}
 affiliations_sheet <- 
@@
 # Check who hasn't responded
-authors2 %>% filter(!acknowledge, !(name %in% responses_sheet$name)) %>% pull(name)
+authors2 %>% filter(!acknowledge, !(name %in% (responses_sheet$name))) %>% pull(name)
 
@@
     # no response
-    acknowledge = ifelse(!(name %in% responses_sheet$name), TRUE, acknowledge)
+    acknowledge = ifelse(!(name %in% (responses_sheet$name)), TRUE, acknowledge)
   )

### 2021-04-28 R/load_data.R

@@ -2,13 +2,13 @@
 load_data <- function() {
   
   if(!exists("austraits", envir = .GlobalEnv )) {
-    austraits <- readRDS("data/austraits_2.1.0.rds")
+    austraits <- readRDS("data/austraits_3.0.0.rds")
     assign("austraits", austraits, envir = .GlobalEnv)
   }
   
   ## Trait lookup table
 
-  assign("trait_category_lookup", read_csv("data/trait_categories.csv", col_types = cols(.default = "c")), envir = .GlobalEnv)
+  assign("trait_category_lookup", read_csv("data/traits_names_list_20210401v2 - traits_names_list_20210401v2.csv", col_types = cols(.default = "c")), envir = .GlobalEnv)
   
   ### 
   sources_all <- 
@@ -42,7 +42,7 @@ load_data <- function() {
       Type= "",
       `all` = n(),
       `geo.` = sum(!is.na(`longitude (deg)`) & !is.na(`latitude (deg)`)),
-      studies = n_distinct(dataset_id),
+      datasets = n_distinct(dataset_id),
       taxa = n_distinct(taxon_name),
       families = n_distinct(family),
       refs =  refs %>% unique() %>% paste(collapse = ", ")

### 2021-04-28 build_figures.R

@@ -1,5 +1,6 @@
 #### Figures for the Austraits manuscript ####
 
+#devtools::install_deps()
 devtools::load_all()
 
 load_data()
@@ -14,13 +15,13 @@ source("scripts/fig_climate_space_austraits.R") # B
 pdf("figures/austraits_fig_biome_map.pdf",
     height = 12,
     width = 8)
-austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A')
+austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'a')
 dev.off()
 
 png("figures/austraits_fig_biome_map.png",
     height = 12,
     width = 8, units="in", res=300)
-austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'A')
+austraits_climate_space / austraits_site_locations_by_tissue_fig_2 +  plot_annotation(tag_levels = 'a')
 dev.off()
 
 #### Figure 3: trait composition count ####
@@ -29,7 +30,7 @@ source("scripts/fig_trait_composition.R")
 pdf("figures/austraits_fig_composition.pdf",
     height = 6,
     width = 8)
-austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'A')
+austraits_composition_georef / austraits_composition_non_georef +  plot_annotation(tag_levels = 'a')
 dev.off()
 
 #### Figure 4: Phylogenetic representation of AusTraits  ####
@@ -39,7 +40,7 @@ source("scripts/fig_link_tree_and_trait_count_matrix.R") # C
 
 
 pdf(
-  "figures/austraits_fig_austraits_phylogenetic_coverage1.pdf",
+  "figures/austraits_fig_austraits_phylogenetic_coverage.pdf",
   height = 12,
   width = 12
 )
@@ -56,3 +57,112 @@ pdf( "figures/austraits_fig_austraits_try_comparison.pdf",
 p_aus_try
 dev.off()
 
+
+#### Figure - coverage
+
+trait_species <- 
+  austraits$traits %>% group_by(taxon_name, trait_name) %>% summarise(n=n()) %>% ungroup
+
+trait_species_wide <- 
[…]

### 2021-05-16 ms.Rmd
requested changes from journal

@@
 ---
-title: "AusTraits: A curated plant trait database for the Australian flora"
+title: "AusTraits, a curated plant trait database for the Australian flora"
 editor_options:
@@
 
-A full list of traits and their sources appears in Online-only Table 1 `r n_records$refs %>% str_split(", ") %>% unlist() %>% sort() %>% unique %>%  
+A full list of traits and their sources appears in Supplementary Table 1 `r n_records$refs %>% str_split(", ") %>% unlist() %>% sort() %>% unique %>%  
     paste(collapse = ", ") %>% paste0("\\citep{", ., "}")`.
@@
 Across all traits, the median number of taxa with records is 62.
-Online-only Table 1 shows the number of studies, taxa, and families recording data in AusTraits, as well as the number of geo-referenced records, for each trait.
+Supplementary Table 1 shows the number of studies, taxa, and families recording data in AusTraits, as well as the number of geo-referenced records, for each trait.
 Looking across traits and tissue categories, coverage declined gradually, with moderate coverage(> 20%) for more than 50 traits (Figure 2).
@@
 
-\newpage
-
-Online-only Table 1: Details on all traits represented in version `r austraits$build_info$version` of AusTraits.
+Supplementary Table 1: Details on all traits represented in version `r austraits$build_info$version` of AusTraits.
 Note the count of studies is less than the number of references when studies are linked to multiple references.
@@
 ```{r}
+# Supplementary Table 1
  n_records %>%

### 2021-07-14 ms.Rmd
Convert to word doc

@@
 output:
-  pdf_document:
-    latex_engine: xelatex
-    citation_package: natbib
-    keep_tex: true
-    keep_md: true
-biblio-style: naturemag
-header-includes:
-- \usepackage{inputenc}
-- \usepackage{float}
-- \usepackage{fontspec}
-- \usepackage{unicode-math}
-- \usepackage{booktabs}
-- \usepackage{longtable}
-- \usepackage{array}
-- \usepackage{multirow}
-- \usepackage{wrapfig}
-- \usepackage{colortbl}
-- \usepackage{lscape}
-- \usepackage{tabu}
-- \usepackage{threeparttable}
-- \usepackage{threeparttablex}
-- \usepackage[normalem]{ulem}
-- \usepackage{makecell}
-- \usepackage{xcolor}
-- \usepackage[square,sort,comma,numbers]{natbib}
-bibliography: [data/references.bib]
-
+  word_document: default
+csl: nature.csl
+bibliography: data/references.bib
 ---
@@
 
-Species traits are essential for comparing ecological strategies among plants, both within any given vegetation, and also arrayed across environmental space and evolutionary lineages \citep{diaz2016global, kunstler2016plant, zanne2014three, cornwell2014functional}.
-Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function \citep{chapin1993evolution, adler2014functional, diaz1998plant,violle2007let}.
-Traits thereby provide useful indicators of species' behaviours in communities and ecosystems, regardless of their taxonomy \citep{Westoby-1998, violle2007let, funk2017revisiting}.
-Through global initiatives the volume of available trait information for plants has grown rapidly in the last two decades \citep{Kattge2011try, Kattge-2020}.
+Species traits are essential for comparing ecological strategies among plants, both within any given vegetation, and also arrayed across environmental space and evolutionary lineages [@diaz2016global; @kunstler2016plant; @zanne2014three; @cornwell2014functional].
+Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function [@chapin1993evolution; @adler2014functional; @diaz1998plant,violle2007let].
+Traits thereby provide useful indicators of species' behaviours in communities and ecosystems, regardless of their taxonomy [@Westoby-1998; @violle2007let; @funk2017revisiting].
+Through global initiatives the volume of available trait information for plants has grown rapidly in the last two decades [@Kattge2011try; @Kattge-2020].
 However, the geographic coverage of trait measurements across the globe is patchy, limiting detailed analyses of trait variation and diversity in some regions, and, more generally, development of theory accounting for the diversity of plant strategies.
@@
 One such region is Australia; a continent with a flora of c.
-28,900 native vascular plant taxa \citep{apc-2020} (including species, subspecies, varietas and forma).
+28,900 native vascular plant taxa [@apc-2020] (including species, subspecies, varietas and forma).
 While significant investment has been made in curating and digitising herbarium collections and observation records in Australia over the last two decades (e.g.
@@
 Moreover, relatively few Australian species are represented in the leading global databases.
-For example, the international TRY database \citep{Kattge-2020} has measurements for only 3830 Australian species across all collated traits.
-This level of species coverage limits our ability to use traits to understand and ultimately manage Australian vegetation \citep{Kissling_2018}.
-While initiatives such as TRY \citep{Kattge-2020} and the Open Traits Network \citep{Gallagher-2020} are working towards global synthesis of trait data, a stronger representation of Australian plant taxa in these efforts is essential, especially given the high richness and endemicity of this continental flora, and the unique contribution this makes to global floral diversity \citep{chapman2009numbers, Hopper-2004}.
+For example, the international TRY database [@Kattge-2020] has measurements for only 3830 Australian species across all collated traits.
+This level of species coverage limits our ability to use traits to understand and ultimately manage Australian vegetation [@Kissling_2018].
+While initiatives such as TRY [@Kattge-2020] and the Open Traits Network [@Gallagher-2020] are working towards global synthesis of trait data, a stronger representation of Australian plant taxa in these efforts is essential, especially given the high richness and endemicity of this continental flora, and the unique contribution this makes to global floral diversity [@chapman2009numbers; @Hopper-2004].
 
@@
 
-AusTraits has been developed as a standalone database, rather than as part of the existing global database TRY \citep{Kattge-2020}, for three reasons.
+AusTraits has been developed as a standalone database, rather than as part of the existing global database TRY [@Kattge-2020], for three reasons.
 First, we sought to establish an engaged and localised community, actively collaborating to enhance coverage of plant trait data within Australia.
@@
 In coming years, AusTraits will continue to be expanded, with integrations into other biodiversity platforms and expansion of coverage into historically neglected plant lineages in trait science, such as pteridophytes (lycophytes and ferns).
-Further, through international initiatives, such as the Open Traits Network and TraitNet, linkages are being forged between plant datasets and a variety of other organismal databases \citep{Naeem-2009, Gallagher-2020}.
+Further, through international initiatives, such as the Open Traits Network and TraitNet, linkages are being forged between plant datasets and a variety of other organismal databases [@Naeem-2009; @Gallagher-2020].
 
@@
 
-A full list of traits and their sources appears in Supplementary Table 1 `r n_records$refs %>% str_split(", ") %>% unlist() %>% sort() %>% unique %>%  
-    paste(collapse = ", ") %>% paste0("\\citep{", ., "}")`.
-This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics \citep{Garnier-2017}.
+A full list of traits and their sources appears in Supplementary Table 1 `r n_records$refs %>% str_split(", ") %>% unlist() %>% sort() %>% unique %>% paste(collapse = "; @") %>% paste0("[@", ., "]")`.
+This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics [@Garnier-2017].
 We categorised traits based on the tissue where it is measured (`r trait_categories$tissue %>% unique() %>% sort() %>% paste(collapse=", ")`) and the type of measurement (`r trait_categories$category %>% unique() %>% sort() %>% gsub("_", " ", .) %>% paste(collapse=", ")`).
@@
 
-The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology \citep{madin2007ontology} in that, where available, trait data are connected to contextual information about the collection (e.g.
+The schema of AusTraits broadly follows the principles of the established Observation and Measurement Ontology [@madin2007ontology] in that, where available, trait data are connected to contextual information about the collection (e.g.
 location coordinates, light levels, whether data were collected in the field or lab) and information about the methods used to derive measurements (e.g.
@@
 
-To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R \citep{R-2020}, using custom code, and the packages `tidyverse` \citep{wickham-2019}, `yaml` \citep{stephens2014yaml}, `remake` \citep{fitzjohn2016remake},  `knitr` \citep{xie1general}, and `rmarkdown` \citep{allaire2015rmarkdown}.
+To harmonise each source into the common AusTraits format we applied a reproducible and transparent workflow (Figure 1), written in R [@R-2020], using custom code, and the packages `tidyverse` [@Wickham-2019], `yaml` [@stephens2014yaml], `remake` [@fitzjohn2016remake],  `knitr` [@xie1general], and `rmarkdown` [@allaire2015rmarkdown].
 In this workflow, we performed a series of operations, including reformatting data into a standardised format, generating observation ids for each set of linked measurements, transforming variable names into common terms, transforming data into common units, standardising terms (trait values) for categorical variables, encoding suitable metadata, and flagging data that did not pass quality checks.
@@
 
-We developed a custom workflow to clean and standardise taxonomic names using the latest and most comprehensive taxonomic resources for the Australian flora: the Australian Plant Census (APC) \citep{apc-2020} and the Australian Plant Names Index (APNI) \citep{apni-2020}.
+We developed a custom workflow to clean and standardise taxonomic names using the latest and most comprehensive taxonomic resources for the Australian flora: the Australian Plant Census (APC) [@apc-2020] and the Australian Plant Names Index (APNI) [@apni-2020].
 These resources document all known taxonomic names for Australian plants, including currently accepted names and synonyms.
-While several automated tools exist for updating taxonomy, such as `taxize` \citep{Chamberlain-2013}, these do not currently include up to date information for Australian taxa.
+While several automated tools exist for updating taxonomy, such as `taxize` [@Chamberlain-2013], these do not currently include up to date information for Australian taxa.
 Updates were completed in two steps.
@@
 
-Static versions of AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available via Zenodo \citep{falster_2019}.
+Static versions of AusTraits, including version `r austraits$build_info$version` used in this descriptor, are available via Zenodo [@falster_2019].
 Data is released under a CC-BY license enabling reuse with attribution -- being a citation of this descriptor and, where possible, original sources.
-Deposition within Zenodo helps makes the dataset consistent with FAIR principles \citep{wilkinson2016fair}.
+Deposition within Zenodo helps makes the dataset consistent with FAIR principles [@wilkinson2016fair].
 As an evolving data product, successive versions of AusTraits are being released, containing updates and corrections.
-Versions are labeled using semantic versioning to indicate the change between versions \citep{Falster-2019}.
+Versions are labeled using semantic versioning to indicate the change between versions [@Falster-2019].
 As validation (see Technical Validation, below) and data entry are ongoing, users are recommended to pull data from the particular release, to ensure results in their downstream analyses remain consistent as the database is updated.
@@
 
-The number of accepted vascular plant taxa in the APC (as of May 2020) is around 28,981  \citep{apc-2020}.
+The number of accepted vascular plant taxa in the APC (as of May 2020) is around 28,981  [@apc-2020].
 Version `r austraits$build_info$version` of AusTraits includes at least one record for `r austraits$taxa %>% filter(taxonRank %in% c("Species", "Subspecies", "Varietas", "Forma")) %>% nrow() %>% format(big.mark=",")` taxa (~ 93% of known taxa).
@@
 Using the taxon names aligned with the APC, data can be queried against location data from the Atlas of Living Australia.
-To create the phylogenetic tree in Figure 6, we pruned a master tree for all higher plants \citep{Smith-2018} using the package `V.PhyloMaker` \citep{Jin-2020} and visualising via `ggtree` \citep{Yu-2017}.
-To create Figure 3A, we used the package `plotbiomes` \citep{Stefan-2020} to create the baseline plot of biomes.
+To create the phylogenetic tree in Figure 6, we pruned a master tree for all higher plants [@Smith-2018] using the package `V.PhyloMaker` [@Jin-2020] and visualising via `ggtree` [@Yu-2017].
+To create Figure 3A, we used the package `plotbiomes` [@Stefan-2020] to create the baseline plot of biomes.
 
@@
 We thank H Cornelissen, H Poorter, SC McColl-Gausden, and one anonymous reviewer for feedback on an earlier draft, and K Levett for advice on data structures.
-This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910), a grant from Macquarie University to Gallagher, and investment from the Australian Research Data Commons (ARDC), via their "Transformation data collections" ([doi:
-10.47486/TD044](https://doi.org/10.47486/TD044)) and "Data Partnerships" ([doi: 10.47486/DP720](https://doi.org/10.47486/DP720)) programs.
+This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910), a grant from Macquarie University to Gallagher, and grants from the Australian Research Data Commons (ARDC), via their "Transformative data collections" ([doi:10.47486/TD044](https://doi.org/10.47486/TD044)) and "Data Partnerships" ([doi: 10.47486/DP720](https://doi.org/10.47486/DP720)) programs.
 The ARDC is enabled by National Collaborative Research Investment Strategy (NCRIS).
@@
 
-RVG, IJW conceived the original idea; RVG, EHW, CB, SA collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SCA, CB, JL error-checked trait measurements; DI developed figures for the paper; DSF, RVG, DI, EHW wrote the first draft of the paper.
+RVG, IJW conceived the original idea; RVG, EHW, CB, SA collated data from primary sources; DSF developed the workflow for the harmonising of data and led all coding; EHW, DI, SCA, JL contributed to coding; EHW, SCA, CB, JL error-checked trait measurements; AM, AF assisted with workflow for updating taxonomy; DI developed figures for the paper; FK, DSF developed the R package; DSF, RVG, DI, EHW wrote the first draft of the paper.
 All other authors contributed the raw data and metadata underpinning the resource, reviewed the harmonised data for errors, and reviewed the final paper for publication.
@@
 
-![The data curation pathway used to assemble the AusTraits database.
+**Fig 1:** The data curation pathway used to assemble the AusTraits database.
 Trait measurements are accessed from original data sources, including published floras and field campaigns.
@@
 Versioned releases are distributed to users, allowing the dataset to be used and re-used in a reproducible way.
-](figures/Pipeline.png)
 
-\clearpage
-
-![Coverage of traits by taxa.
+**Fig 2:** Coverage of traits by taxa.
 (a) Matrix showing the coverage of taxa for each trait, with yellow indicating presence of data.
@@
 (b) Number of taxa with data for first 100 traits for all traits and separated by tissue.
-](figures/austraits_coverage_by_trait.png)
 
-\clearpage
+**Fig 3:** Number of taxa with trait records by plant tissue and trait category, for data that are  (a) Geo-referenced, and (B) Not geo-referenced.
+Many records without a geo-reference come from botanical collections, such as floras.
 
-![Number of taxa with trait records by plant tissue and trait category, for data that are  (a) Geo-referenced, and (B) Not geo-referenced.
-Many records without a geo-reference come from botanical collections, such as floras.
-](figures/austraits_fig_composition.pdf)
+**Fig 4:** Coverage of geo-referenced trait records across Australian climatic and geographic space for traits in different categories.
+(A) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s classification of major biomes by climate [@Whittaker-1975].
+Climate data were extracted at 10" resolution from WorldClim [@Fick-2017].(B) Locations of geo-referenced records for different plant tissues.
 
-\clearpage
-
-![Coverage of geo-referenced trait records across Australian climatic and geographic space for traits in different categories.
-(A) AusTraits’ sites (orange) within Australia’s precipitation-temperature space (dark-grey) superimposed upon Whittaker’s classification of major biomes by climate \citep{Whittaker-1975}.
-Climate data were extracted at 10" resolution from WorldClim \citep{Fick-2017}.(B) Locations of geo-referenced records for different plant tissues.
-](figures/austraits_fig_biome_map.png)
-
-\clearpage
-
-![Phylogenetic distribution of trait data in AusTraits for a subset of 2000 randomly sampled taxa.
+**Fig 5:** Phylogenetic distribution of trait data in AusTraits for a subset of 2000 randomly sampled taxa.
 The heatmap colour intensity denotes the number of traits measured within a family for each plant tissue.
 The most widespread family names (with more than ten taxa) are labelled on the edge of the tree.
-](figures/austraits_fig_austraits_phylogenetic_coverage.png)
 
-\clearpage
+**Fig 6:** The number of taxa with trait records in AusTraits and global TRY database (accessed 28 May 2020). Each point shows a separate trait.
 
-![The number of taxa with trait records in AusTraits and global TRY database (accessed 28 May 2020). Each point shows a separate trait. ](figures/austraits_fig_austraits_try_comparison.pdf)
-
-\clearpage
-
-Table 1:  Main elements of the harmonised AusTraits database. See Tables 2-8 for details on each component.
+**Table 1:**  Main elements of the harmonised AusTraits database. See Tables 2-8 for details on each component.
 
 ```{r}
-clean_quotes <- function(txt) {
-  str_replace_all(txt, "`([a-zA-Z_ \\-0-9()\\.\\\\_]+)`", " texttt{\\1}") %>% 
-  str_replace_all(fixed(" texttt"), fixed("\\texttt")) 
-}
-
 tibble(
@@
   ) %>% 
-  knitr::kable("latex",  booktabs = TRUE) %>%
-  kableExtra::kable_styling(latex_options = c("striped")) %>%
-  kableExtra::column_spec(1, width = "15em") %>%
-  kableExtra::column_spec(2, width = "30em") %>% 
-  str_replace_all(fixed("gray!6"), fixed("gray!20"))  %>%
-    clean_quotes() %>%
-  writeLines()
+  write_csv("figures/table1.csv")
 ```
@@
 
-    writeLines(sprintf("Table %s: Structure of the `%s` table, %s \n\n", i, v, gsub("A table", "", elements$description)))
+    writeLines(sprintf("**Table %s**: Structure of the `%s` table, %s \n\n", i, v, gsub("A table", "", elements$description)))
 
@@
     list1_to_df() %>%
-#    mutate(value = clean_quotes(value)) %>%
-    knitr::kable("latex",  booktabs = TRUE) %>% 
-    kableExtra::kable_styling(latex_options = c("striped")) %>%
-    kableExtra::column_spec(1, width = "15em") %>%
-    kableExtra::column_spec(2, width = "30em") %>%
-    str_replace_all(fixed("gray!6"), fixed("gray!20"))  %>%
-    clean_quotes() %>%
-    writeLines()
+    write_csv(sprintf("figures/table%d.csv",i))
 
@@
 
-\clearpage
-
-Table 9: Possible value types of trait records.
+**Table 9:** Possible value types of trait records.
 
@@
   list1_to_df()  %>%
-  knitr::kable("latex",  booktabs = TRUE) %>%
-  kableExtra::kable_styling(latex_options = c("striped")) %>%
-  kableExtra::column_spec(1, width = "15em") %>%
-  kableExtra::column_spec(2, width = "30em") %>%
-  str_replace_all(fixed("gray!6"), fixed("gray!20")) %>%
-  clean_quotes() %>%
-  writeLines()
+  write_csv("figures/table9.csv")
 ```
-
-Supplementary Table 1: Details on all traits represented in version `r austraits$build_info$version` of AusTraits.
-Note the count of studies is less than the number of references when studies are linked to multiple references.
 
@@
   select(Group, everything()) %>%
-  nest(data = -Group) %>% 
-  mutate(
-    ## turn each group into table
-    tex = purrr::map(data, make_latex_table),
-    ## drop top and bottom of each table
-    tex_sub =  purrr::map2_chr(tex, Group, ~c(sprintf("\\multicolumn{3}{l}{{\\bf %s}}\\\\", .y),.x[ seq(1 + which(grepl("endlastfoot", .x)), -1 + which(grepl("end{longtable}", .x, fixed=TRUE)))], "\\\\ \\midrule \n  \\\\ \n ") %>% paste(collapse = "\n"))
-    ) -> x
+  split(., .$Group) %>% purrr::map(~bind_rows(tibble(Group=.x$Group[1]), .x)) %>% bind_rows() %>% mutate_all(~replace_na(.x, "")) %>%
+  mutate(Group = ifelse(duplicated(Group), "", Group)) ->x
 
-c(x$tex[[1]][1:18], unlist(x$tex_sub), x$tex[[1]][c(-1:0+length(x$tex[[1]]))]) %>%
-  writeLines()
+  write_csv(x, "figures/table_online1.csv")
+
+  x %>% select(Group, Trait, refs) %>% 
+    mutate_all(~paste(.x, ";") %>% gsub("\\\\newline ", "",.)) %>%
+    knitr::kable()
 ```

### 2021-07-14 R/ms.R

@@ -15,18 +15,11 @@ list1_to_df <- function(my_list) {
 ## Separate into groups of max size n. 
 ## This is needed for latex table, overwise citations
 ## spill off the page
-format_cites_n <- function(x, n = 5){
+format_cites_n <- function(x){
   
   # Make sorted vector of all unique refs
-  x2 <- str_split(x, ", ") %>% unlist() %>% sort()
-  # Make a vector with group ids for references
-  groups <- (rep(1:ceiling(length(x2)/n), n) %>% sort() )[seq_len(length(x2))]
-  
-  ## wrap given txt with citation call  
-  format_cites_sub <- function(x) {x %>% paste(collapse = ", ") %>% paste0("\\citep{", ., "}" )}
-  
-  ## split into groups of n citations, wrap each with citation call, the combine
-  split(x2, groups) %>% lapply(format_cites_sub) %>% paste(collapse = " \\newline ")
+  str_split(x, ", ") %>% unlist() %>% sort() %>% 
+    paste(collapse = "; @") %>% paste0("[  [@", ., "]  ]" )
 }
 
 

### 2021-07-18 R/load_data.R
Update data (v3.0.2)

@@ -1,8 +1,8 @@
 
-load_data <- function() {
+load_data <- function(update=FALSE) {
   
-  if(!exists("austraits", envir = .GlobalEnv )) {
-    austraits <- readRDS("data/austraits_3.0.0.rds")
+  if(update | !exists("austraits", envir = .GlobalEnv )) {
+    austraits <- readRDS("data/austraits-3.0.2.rds")
     assign("austraits", austraits, envir = .GlobalEnv)
   }
   

### 2021-07-18 ms.Rmd
Revise paper
- updated reference list
- reformat references (don't cite unpublished, correct some citations)
- change online table to include dataset ids
- more accurate description of differences to TRY

@@
 devtools::load_all()
-load_data()
+load_data(update=TRUE)
 
@@
 
-\normalsize
-
 ## Abstract
@@
 AusTraits synthesises data on `r austraits$traits$trait_name %>% n_distinct()` traits across `r austraits$traits$taxon_name %>% n_distinct()` taxa from field campaigns, published literature, taxonomic monographs, and individual taxon descriptions.
-Traits
-vary in scope from physiological measures of performance (e.g. photosynthetic gas exchange, water-use efficiency) to morphological
-attributes (e.g. leaf area, seed mass, plant height) which link to aspects of ecological variation. AusTraits contains curated and
-harmonised individual- and species-level measurements coupled to, where available, contextual information on site properties  and experimental conditions.
+Traits vary in scope from physiological measures of performance (e.g.
+photosynthetic gas exchange, water-use efficiency) to morphological
+attributes (e.g.
+leaf area, seed mass, plant height) which link to aspects of ecological variation.
+AusTraits contains curated and harmonised individual- and species-level measurements coupled to, where available, contextual information on site properties  and experimental conditions.
 This article provides information on version  `r austraits$build_info$version` of AusTraits which contains data for `r austraits$traits %>% nrow()` trait-by-taxon combinations.
@@
 
-Species traits are essential for comparing ecological strategies among plants, both within any given vegetation, and also arrayed across environmental space and evolutionary lineages [@diaz2016global; @kunstler2016plant; @zanne2014three; @cornwell2014functional].
-Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function [@chapin1993evolution; @adler2014functional; @diaz1998plant,violle2007let].
+Species traits are essential for comparing ecological strategies among plants, both within any given vegetation and across environmental space or evolutionary lineages [@zanne2014three; @cornwell2014functional; @diaz2016global; @kunstler2016plant].
+Broadly, a trait is any measurable property of a plant capturing aspects of its structure or function [@chapin1993evolution; @adler2014functional; @diaz1998plant, @violle2007let].
 Traits thereby provide useful indicators of species' behaviours in communities and ecosystems, regardless of their taxonomy [@Westoby-1998; @violle2007let; @funk2017revisiting].
@@
 Here we introduce the AusTraits database (hereafter AusTraits), a compilation of plant traits for the Australian flora.
-Currently, AusTraits draws together `r austraits$sources %>% length()` primary sources and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
+Currently, AusTraits draws together `r austraits$traits$dataset_id %>% n_distinct()` distinct sources and contains `r austraits$traits %>% nrow()` measurements spread across `r austraits$traits$trait_name %>% n_distinct()` different traits for  `r austraits$traits$taxon_name %>% n_distinct()` taxa.
 To assemble AusTraits from diverse primary sources and make data available for reuse, we needed to overcome three main types of challenges (Figure 1): 1) Accessing data from diverse original sources, including field studies, online databases, scientific articles, and published taxonomic floras; 2) Harmonising these diverse sources into a federated resource, with common taxon names, units, trait names, and data formats; and 3) Distributing versions of the data under suitable license.
@@
 First, we sought to establish an engaged and localised community, actively collaborating to enhance coverage of plant trait data within Australia.
-It was our opinion that a community would form more readily to fill gaps in national knowledge of traits with local ownership of the resource.
+We envisioned  that a community would form more readily to fill gaps in national knowledge of traits with local ownership of the resource.
 While we will never have a counterfactual, a vibrant community excited to be part of this initiative has indeed been established and coverage is much higher for Australian species than has been achieved since TRY’s inception.
 Local ownership also aligns well with funding opportunities and national research priorities, and enables database coordinators to progress at their own speed.
-Second, we wanted to implement new features not available in TRY.
-For instance, AusTraits is fully reproducible, meaning the entire database can be rebuilt from its original sources at any point.
-This has enabled us to actively curate the dataset to reduce duplications and update mapping of traits as contributors provide feedback and trait definitions and values are edited.
-Moreover, the curation is achieved on GitHub, such that the history of all changes is traceable and, where necessary, reversible.
-Third, AusTraits will be distributed via different mechanisms to TRY.
-Data from TRY is requested via a standalone portal.
-By contrast, versions of AusTraits are being regularly archived on a third party data repository (Zenodo), built for long term data archiving and an established API.
-This will allow for active connection to a variety of end users including, but not only, TRY.
+Second, we wanted to apply an entirely open-source approach to the aggregation workflow [following @Falster-2015].
+All the code and raw files used to create the compiled database are available, and this database is freely available via a third party data repository (Zenodo) which is itself built for long term data archiving, with an established API.
+Finally, we targeted primary data sources, where possible; whereas TRY accepts aggregated datasets.
+The hope was that this would increase data quality, by removing intermediaries and easier identification of duplicates.
+
+While independent, the overall structure of AusTraits is similar to that of TRY, ensuring the two databases will be interoperable.
+Both databases are founded on similar principles and terminology [@madin2007ontology; @Garnier-2017].
+Increasingly, researchers and biodiversity portals are seeking to connect diverse datasets [@Gallagher-2020], which is possible if they share a common foundation.
 
@@
 In coming years, AusTraits will continue to be expanded, with integrations into other biodiversity platforms and expansion of coverage into historically neglected plant lineages in trait science, such as pteridophytes (lycophytes and ferns).
-Further, through international initiatives, such as the Open Traits Network and TraitNet, linkages are being forged between plant datasets and a variety of other organismal databases [@Naeem-2009; @Gallagher-2020].
+Further, through international initiatives, such as the Open Traits Network, linkages are being forged between plant datasets and a variety of other organismal databases [@Gallagher-2020].
 
@@
 
-AusTraits version `r austraits$build_info$version` was assembled from `r austraits$sources %>% length()` distinct sources, including published papers, field measurements, glasshouse and field experiments, botanical collections, and taxonomic treatments.
+AusTraits version `r austraits$build_info$version` was assembled from `r austraits$traits$dataset_id %>% n_distinct()` distinct sources, including published papers, field measurements, glasshouse and field experiments, botanical collections, and taxonomic treatments.
 Initially we identified a list of candidate traits of interest, then identified primary sources containing measurements for these traits, before contacting authors for access.
@@
 
-A full list of traits and their sources appears in Supplementary Table 1 `r n_records$refs %>% str_split(", ") %>% unlist() %>% sort() %>% unique %>% paste(collapse = "; @") %>% paste0("[@", ., "]")`.
+A full list of traits and their sources appears in Supplementary Table 1 `r c(n_records$refs_primary_pub, n_records$refs_secondary_pub) %>% collapse("; @") %>% paste0("[@", ., "]")` `r c(n_records$unpub_people) %>% collapse(", ") %>% paste0("(", ., "; unpublished data)")`.
 This list was developed gradually as new datasets were incorporated, drawing from original source publications and a published thesaurus of plant characteristics [@Garnier-2017].
@@
 
-An R package is being developed (see [github.com/traitecoevo/austraits](http://github.com/traitecoevo/austraits)) to provide easy access and guidance on usage for those using this platform.
+The R package `austraits` (https://github.com/traitecoevo/austraits) provides easy access to data and examples on manipulating data (e.g. joining tables, subsetting) for those using this platform.
 
 ### Data coverage
-
-```{r}
-#As of May 2020, the APC included 28,981 accepted species, subspecies, varieties and forma.  
-```
 
@@
 Yet, geo-referenced records were available in several traits for more than 10% of the flora (Figure 3a).
-Coverage is notably higher for geo-referenced measurements of some tissues and trait types - such as bark stems and roots - relative to non-georeferenced measurements (Figure 3).
+Coverage is notably higher for geo-referenced measurements of some tissues and trait types - such as bark stems and roots - relative to non-geo-referenced measurements (Figure 3).
 
@@
 We implemented three strategies to maintain data quality.
-First, we conducted a detailed review of each source based on a bespoke report, showing all data and metadata, by both an AusTraits curator and the original contributor (where possible).
+First, we conducted a detailed review of each source based on a bespoke report, showing all data and metadata, by both an AusTraits curator (primarily Wenk) and the original contributor (where possible).
 Measurements for each trait were plotted against all other values for the trait in AusTraits, allowing quick identification of outliers.
@@
 
-Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of accepted values maintained by the curators.
+Second, we implemented automated tests for each dataset, to confirm that values for continuous traits fall within the accepted range for the trait, and that values for categorical traits are on a list of allowed values.
 Data that did not pass these tests were moved to a separate spreadsheet (“excluded_data”) that is also made available for use and review.
@@
 Third, we provide a pathway for user feedback.
-AusTraits is a community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
+AusTraits is an open-source community resource and we encourage engagement from users on maintaining the quality and usability of the dataset.
 As such, we welcome reporting of possible errors, as well as additions and edits to the online documentation for AusTraits that make using the existing data, or adding new data, easier for the community.
-Feedback can be posted as an issue directly at the projects GitHub page (http://traitecoevo.github.io/austraits.build).
+Feedback can be posted as an issue directly at the project's GitHub page (http://traitecoevo.github.io/austraits.build).
 
@@
 To create the phylogenetic tree in Figure 6, we pruned a master tree for all higher plants [@Smith-2018] using the package `V.PhyloMaker` [@Jin-2020] and visualising via `ggtree` [@Yu-2017].
-To create Figure 3A, we used the package `plotbiomes` [@Stefan-2020] to create the baseline plot of biomes.
+To create Figure 3a, we used the package `plotbiomes` [@Stefan-2020] to create the baseline plot of biomes.
 
@@
 The code used to produce this paper is available at http://github.com/traitecoevo/austraits_ms.
-(All code will be made available prior to final publication.)
 
@@
 We thank H Cornelissen, H Poorter, SC McColl-Gausden, and one anonymous reviewer for feedback on an earlier draft, and K Levett for advice on data structures.
-This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910), a grant from Macquarie University to Gallagher, and grants from the Australian Research Data Commons (ARDC), via their "Transformative data collections" ([doi:10.47486/TD044](https://doi.org/10.47486/TD044)) and "Data Partnerships" ([doi: 10.47486/DP720](https://doi.org/10.47486/DP720)) programs.
+This work was supported by fellowship grants from Australian Research Council to Falster (FT160100113), Gallagher (DE170100208) and Wright (FT100100910), a grant from Macquarie University to Gallagher, and grants from the Australian Research Data Commons (ARDC), via their "Transformative data collections" (https://doi.org/10.47486/TD044) and "Data Partnerships" (https://doi.org/10.47486/DP720) programs.
 The ARDC is enabled by National Collaborative Research Investment Strategy (NCRIS).
@@
 The authors have no conflicts of interest to declare.
-
-\clearpage
 
@@
 ```{r}
+format_cites <- function(x) {
+  # Make sorted vector of all unique refs
+  str_split(x, ", ") %>% unlist() %>% sort() %>% 
+    paste(collapse = "; @") %>% paste0("[ [@", ., "] ]" )
+}
+
 # Supplementary Table 1
@@
          Type = ifelse(Type=="character", "chr.", Type),
-         refs = purrr::map_chr(refs, format_cites_n),
+         Sources_primary = ifelse(refs_primary_pub =="","",
+                         purrr::map_chr(refs_primary_pub, format_cites)),
+         Sources_secondary = ifelse(refs_secondary_pub =="","",
+                         purrr::map_chr(refs_secondary_pub, format_cites))
          ) %>%
-  select(-Tissue, -Category) %>%
+  select(-Tissue, -Category, -starts_with("refs"), -unpub_people) %>%
   select(Group, everything()) %>%
@@
 
-  x %>% select(Group, Trait, refs) %>% 
+  x %>% select(Group, Trait, Sources_primary, Sources_secondary) %>% mutate(Newline="XXX") %>% 
     mutate_all(~paste(.x, ";") %>% gsub("\\\\newline ", "",.)) %>%
@@
 
-\clearpage
-
-\footnotesize
+## References

### 2021-07-18 R/load_data.R

@@ -1,3 +1,7 @@
+expand <- function(x) {x %>% str_split(", ") %>% unlist() %>% unique() %>% na.omit() %>% sort()}
+collapse <- function(x, sep=", ") {x %>% expand() %>% na_if("") %>% na.omit() %>% unique() %>% paste(collapse = sep)}
+remove <- function(x, id) {expand(x) %>% subset(., !(. %in% id))  }
+keep <- function(x, id) {expand(x) %>% subset(., (. %in% id))  }
 
 load_data <- function(update=FALSE) {
   
@@ -10,18 +14,26 @@ load_data <- function(update=FALSE) {
 
   assign("trait_category_lookup", read_csv("data/traits_names_list_20210401v2 - traits_names_list_20210401v2.csv", col_types = cols(.default = "c")), envir = .GlobalEnv)
   
-  ### 
+  # identify refs with unpublished data 
+  unpub <- 
+    tibble(key = map_chr(austraits$sources, ~.x$key), title = map_chr(austraits$sources, ~.x$title), 
+           author = map_chr(austraits$sources, ~.x$author[1] %>% paste(collapse = ", "))
+    ) %>% 
+    filter(grepl("Unpublished data", title, fixed=TRUE))
+  
+  # 
   sources_all <- 
     bind_rows(
-      austraits$methods %>% select(dataset_id, citation = source_primary_key),
-      austraits$methods %>% select(dataset_id, citation = source_secondary_key)
+      austraits$methods %>% select(dataset_id, citation_p = source_primary_key),
+      austraits$methods %>% select(dataset_id, citation_s = source_secondary_key)
     ) %>% 
-    na_if("") %>%
-    na.omit() %>% 
     distinct() %>% 
     arrange(dataset_id) %>% 
     group_by(dataset_id) %>% 
-    summarise(refs = sprintf("%s", citation) %>% paste(collapse = ", ") %>% str_replace_all(";", ",") )
+    summarise(
+      refs_primary = citation_p %>% str_replace_all(";", ",") %>% collapse(),
+      refs_secondary = citation_s %>% str_replace_all(";", ",") %>% collapse()
+      )
   
   sites <- austraits$sites %>% 
     filter(site_property %in%  c("longitude (deg)","latitude (deg)")) %>% 
@@ -34,6 +46,7 @@ load_data <- function(update=FALSE) {
               select(austraits$taxa, taxon_name, family)) %>% 
     left_join(sources_all, by=c("dataset_id")) 
   
+  
[…]

### 2021-07-18 R/ms.R

@@ -10,19 +10,6 @@ list1_to_df <- function(my_list) {
 }
 
 
-
-## Add `\citet{}` for references. 
-## Separate into groups of max size n. 
-## This is needed for latex table, overwise citations
-## spill off the page
-format_cites_n <- function(x){
-  
-  # Make sorted vector of all unique refs
-  str_split(x, ", ") %>% unlist() %>% sort() %>% 
-    paste(collapse = "; @") %>% paste0("[  [@", ., "]  ]" )
-}
-
-
 ## format table
 make_latex_table <- function(data) {
   data %>%
